### 2021-01-14 docs/manuscript/Global-Water-Access-v5-wo-comments.tex
Adding manuscript and bib file.

\documentclass[10pt,twoside]{article}
\usepackage{etex, {soul}}
\newcommand{\num}{6{} }

\raggedbottom

%geometry (sets margin) and other useful packages
\usepackage{geometry}
\geometry{top=.8in, left=.8in,right=.8in,bottom=.8in}
 \usepackage{graphicx,booktabs,calc}

%=== GRAPHICS PATH ===========
%\graphicspath{{./140408-Images/}}
% Marginpar width
%Marginpar width
\newcommand{\pts}[1]{\marginpar{ \small\hspace{0pt} \textit{[#1]} } }
\setlength{\marginparwidth}{.5in}
%\reversemarginpar
%\setlength{\marginparsep}{.02in}

%% Fonts
% \usepackage{fourier}
% \usepackage[T1]{pbsi}

\usepackage{lmodern}
\usepackage[T1]{fontenc}
%\usepackage{minted}

\usepackage{rotating}
%% Cite Title
\usepackage[style=authoryear,backend=biber, natbib,maxcitenames=2,doi=false,isbn=false,url=false,eprint=false]{biblatex}
\addbibresource{../References/Water-Accessibility.bib}

%%% Counters
\usepackage{chngcntr,mathtools}
\counterwithout{figure}{section}
\counterwithout{table}{section}

\numberwithin{equation}{section}

%% Captions
\usepackage{caption}
\captionsetup{
  labelsep=quad,
  justification=raggedright,
  labelfont=sc
}

%AMS-TeX packages
\usepackage{amssymb,amsmath,amsthm}
\usepackage{bm}
\usepackage[mathscr]{eucal}
\usepackage{colortbl}
\usepackage{color}

\usepackage{epstopdf,subfigure,hyperref,enumerate,polynom,polynomial}
\usepackage{multirow,minitoc,fancybox,array,multicol}

\definecolor{slblue}{rgb}{0,.3,.62}
\hypersetup{
    colorlinks,%
    citecolor=blue,%
    filecolor=blue,%
    linkcolor=blue,
    urlcolor=slblue
}

%%%TIKZ
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{pgfgantt}
\pgfplotsset{compat=newest}

\usetikzlibrary{arrows,shapes,positioning}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{shadows,automata}
\usetikzlibrary{patterns}
%\usetikzlibrary{circuits.ee.IEC}
\usetikzlibrary{decorations.text}
% For Sagnac Picture
\usetikzlibrary{%
    decorations.pathreplacing,%
    decorations.pathmorphing%
}

%
%Redefining sections as problems
%
\makeatletter
\newenvironment{question}{\@startsection
	{section}
	{1}
	{-.2em}
	{-3.5ex plus -1ex minus -.2ex}
    	{1.3ex plus .2ex}
    	{\pagebreak[3]%forces pagebreak when space is small; use \eject for better results
	\large\bf\noindent{Question }
	}
	}
	%{\vspace{1ex}\begin{center} \rule{0.3\linewidth}{.3pt}\end{center}}
	%\begin{center}\large\bf \ldots\ldots\ldots\end{center}}
\makeatother

%
%Fancy-header package to modify header/page numbering
%
%\renewcommand{\chaptermark}[1]{ \markboth{#1}{} }
\renewcommand{\sectionmark}[1]{ \markright{#1}{} }

\usepackage{fancyhdr}
\pagestyle{fancy}
%\addtolength{\headwidth}{\marginparsep} %these change header-rule width
%\addtolength{\headwidth}{\marginparwidth}
%\fancyheadoffset{30pt}
%\fancyfootoffset{30pt}
\fancyhead[LO,RE]{\small  \it \nouppercase{\leftmark}}
\fancyhead[RO,LE]{\small Page \thepage}
\fancyfoot[RO,LE]{\small }% PR \num S-2015}
\fancyfoot[LO,RE]{\small }%\scshape MODL}
\cfoot{}
\renewcommand{\headrulewidth}{0.1pt}
\renewcommand{\footrulewidth}{0pt}
%\setlength\voffset{-0.25in}
%\setlength\textheight{648pt}

\usepackage{paralist}

%%% FORMAT PYTHON CODE
\usepackage{listings}
% Default fixed font does not support bold face
\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{8} % for bold
\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{8}  % for normal

% Custom colors
\usepackage{color}
\definecolor{deepblue}{rgb}{0,0,0.5}
\definecolor{deepred}{rgb}{0.6,0,0}
\definecolor{deepgreen}{rgb}{0,0.5,0}

%\usepackage{listings}

% % Python style for highlighting
% \newcommand\pythonstyle{\lstset{
% language=Python,
% basicstyle=\footnotesize\ttm,
% otherkeywords={self},             % Add keywords here
% keywordstyle=\footnotesize\ttb\color{deepblue},
% emph={MyClass,__init__},          % Custom highlighting
% emphstyle=\footnotesize\ttb\color{deepred},    % Custom highlighting style
% stringstyle=\color{deepgreen},
% frame=tb,                         % Any extra options here
% showstringspaces=false            %
% }}

% % Python environment
% \lstnewenvironment{python}[1][]
% {
% \pythonstyle
% \lstset{#1}
% }
% {}

% % Python for external files
% \newcommand\pythonexternal[2][]{{
% \pythonstyle
% \lstinputlisting[#1]{#2}}}

% % Python for inline
% \newcommand\pythoninline[1]{{\pythonstyle\lstinline!#1!}}

\newcommand{\osn}{\oldstylenums}
\newcommand{\dg}{^{\circ}}
\newcommand{\lt}{\left}
\newcommand{\rt}{\right}
\newcommand{\pt}{\phantom}
\newcommand{\tf}{\therefore}
\newcommand{\?}{\stackrel{?}{=}}
\newcommand{\fr}{\frac}
\newcommand{\dfr}{\dfrac}
%\newcommand{\ul}{\underline}
\newcommand{\tn}{\tabularnewline}
\newcommand{\nl}{\newline}
\newcommand\relph[1]{\mathrel{\phantom{#1}}}
\newcommand{\cm}{\checkmark}
\newcommand{\ol}{\overline}
\newcommand{\rd}{\color{red}}
\newcommand{\bl}{\color{blue}}
\newcommand{\pl}{\color{purple}}
\newcommand{\og}{\color{orange!90!black}}
\newcommand{\gr}{\color{green!40!black}}
\newcommand{\nin}{\noindent}
\newcommand{\la}{\lambda}
\renewcommand{\th}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\G}{\Gamma}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,thick,inner sep=1pt] (char) {\small #1};}}

\newcommand{\bc}{\begin{compactenum}[\quad--]}
\newcommand{\ec}{\end{compactenum}}

\newcommand{\p}{\partial}
\newcommand{\pd}[2]{\frac{\partial{#1}}{\partial{#2}}}
\newcommand{\dpd}[2]{\dfrac{\partial{#1}}{\partial{#2}}}
\newcommand{\pdd}[2]{\frac{\partial^2{#1}}{\partial{#2}^2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\vspace{-3ex}
\title{Quantifying and Modeling Global Water Accessibility including Water Typology}
\author{Hichul Chung \and Emily Kumpel \and Jimi Oke}
\date{1-8-21}
\maketitle

%\thispagestyle{empty}

%\tableofcontents

\section{Introduction}
%%(Roughly 3 paragraphs)
In 2017, 29\% of the global population (2.2 billion people) did not use a safely managed drinking - water service which is defined as water located on-premise, available, and free from contamination \citep{WHOdrinkingwater}.   If you meet about 100 people, you can imagine that 29 of the people you meet do not have access to improved drinking water. Another recognition is that many people spend a lot of time collecting water where their access to drinking water
supplies located on-premises are not common, about 26.3 percent of the world's population
\citep{cassivi2018access}. Where this raises concerns about inequality problems related to the task and time variation 
to collect water can be inferred in the urban and rural areas \citep{cassivi2018access}. Therefore, there needs to be a
great motivation to bring improved drinking water sources to all people to improve their quality of life.

%%\item Cite statistics from the WHO/UN on the water situation...
%%\item What has been the rate of improvement over the years?

Globally, access to improved drinking water sources and sanitation is increasing. From 1990 to 2012, utilization of
improved drinking water sources has increased globally from 76 percent to 89 percent, and utilization of improved
sanitation has increased from 45 percent to 64 percent \citep{fuller2016tracking}. According to the MDG categorization,
they consider improved drinking water sources as following: public tap, borehole, protected spring, rainwater
collection, bottled water source, and piped water \citep{bartram2014globala}. However, these variables do not truly
explain if the households have speedy access to retrieve the water. Therefore, we will investigate further on what
variables should be included to enhance the accessibility of improved drinking water.

% {\reversemarginpar\marginnote{\scriptsize\rd The introduction should end with a paragraph summarizing the contents and layout of the remainder of the manuscript. E.g. ``In the next section we discuss relevant prior research on water accessibility quantification. Next, we present the data and methods...''}}

In the next section we discuss relevant prior research on water accessibility quantification. We are interested in what type of modes of
transportation are utilized to provide clean and drinking water to the community. Also, we desire to learn more about
the time it takes to transport water to the community in a global modeling context. So that we can develop and quantify
water accessibility in different countries to compare different trends. Also, we will be exploring the most efficient
and effective modes of transportation of accessible water. Therefore, the new models will be able to influence and
encourage policymakers in specific countries to improve water accessibility for all. Next we present the data and methods of approaching the quantification of the water accessibility.

\section{Relevant work}
Organized by chronology and the Categorical.

\paragraph{}
\citet{onda2012global}discusses the global access to safe water by accounting for water quality and the resulting impact on MDG progress. In 2012, \citet{onda2012global} altered the current Joint monitoring programme (JMP) estimation by including for microbial water quality and sanitary risk utilizing national water quality data. \citep{onda2012global} used principal components analysis to analyze the national environment and development indicators and then created models. Overall, our analysis can highlight the potential fatigue illness caused by people transporting the water if it is done manually. How the modes of transportation can affect one's health.
\paragraph{}
\citet{sima2013water} highlights the relation of the topic where it discovered that over 76 percent of the city's water is utilized by less than ten percent of household piped water.  The study is focused in Kisumu Kenya and Sima interviewed 260 informal water business operators. This article revealed a new insight where the majority of the population in Kisumu relied on purchased water from kiosks (1.5 million $m^{3}$ per day) and used hand-drawn water-carts (0.75 million $m^{3}$ per day)\citep{sima2013water}. Also, we discovered that the water trucking industry utilized the most energy consumption and was high cost: water delivery trucks have the highest per cubic meter energy demand (35 $MJ/m^{3}$)\citep{sima2013water}. The article utilized SPSS Version 19.0 to analyze the data. 
\paragraph{}
\citet{ho2014challenge} highlights a head-to-head comparison of such indicators with other possible distance and time metrics among rural 1,103 households in Nampula province, Mozambique. The researchers found out that Euclidean line distance to be a good representative for route distance ($R^{2} = 0.98$), while self-reported travel time is a poor representative ($R^{2} = 0.12$). One of key insights was that, 15-minute decrease in one-way walk time to water source is related to a 41\% reduction in diarrhea, improved children's nutrition, and an 11\% reduction in under-five child mortality rate \citep{ho2014challenge}. This research can be a helpful tool to understand how long people are traveling in a rural area in Africa where many of our data focus. Also, understanding how people are traveling to receive the water can better understand the condition of the rural people and implement better policies that help them to thrive. 
\paragraph{}
\citet{bartram2014globala} highlights a background information about the Millennium Development Goals. How it has established global targets for drinking water and sanitation access (Bartram et al. 2014). The article indirectly gives much information about progress towards focused targets, directed by international monitoring, has reduced the global disease burden and improved the quality of life \citep{bartram2014globala}. It utilizes a variety of sources such as JMP, DHS, MICS surveys to analyze the data of many different locations in the world, therefore, it is a broad scope of study of the progress of accessible and drinking water and sanitation globally. Our research can study further what modes of transportation of clean drinking water would affect the progress of improving drinking water and sanitation in one's community. Also, understanding how long it takes to attain the water can improve the quality of the services provided for the people. 
\paragraph{}
\citet{onda2014country} discusses their new development of typology of country clusters pertain to water and sanitation sector based on congruence through multiple related indicators. \citet{onda2014country} used a hierarchical clustering method and a gap statistic analysis to cluster 156 countries which has 6.75 billion people into a relational clusters. \citet{onda2014country} suggests previous geography or income-based country clustering should be improved by using water and sanitation related indicators. Our research can improvise upon their finding and focus on identifying which water accessibility variables may improve water accessibility for people who are seeking clean drinking water. 
\paragraph{}
\citet{WHOdrinkingwater} article gives an ambitious plan to reach universal drinking water by the year of 2030 that is affordable and safe followed by the Sustainable Development Goal six goals. The WHO/UNICEF Joint Monitoring Program for Water Supply and Sanitation (JMP) are monitoring progress on drinking water and sanitation since 1990 globally. The article outlines different plans with consistent traits to determine what is deemed to be safe and affordable drinking water and good sanitation. Our study can highlight which modes of transportation of safe and affordable drinking water can develop why one's community may excel in the improvement of safe and affordable drinking water and sanitation. Drinking, sanitation, and hand washing ladder explains the qualitative determination can indirectly help what is a good source of transporting safe and drinking water. 
\paragraph{}
\citet{fuller2016tracking} highlights the global accessibility to safe drinking water and sanitation and concludes that it has been improving well during the Millennium Development Goal Period. This topic will benefit our research to understand what countries are doing for the mode of transportation of drinking water and sanitation which may have contributed to the improvements. \citet{fuller2016tracking} chose some countries and studied the overall countries progress which in his methodology revealed sigmoidal trends. Our study can reveal if certain methods of transportation of drinking water and sanitation may have influenced the effectiveness of the improved drinking water met by MDG standards.  
\paragraph{}
\citet{jia2016highresolution} highlights that there is inequality within safe sanitation. There is a great difference in wealth levels in many low-income countries. Therefore, this will hinder the improvement of MDG progress. \citet{jia2016highresolution}discusses future interventions to be engaged with different wealth categories. Our study can assist in understanding improved sanitation increase by making inferences about how they receive water may give the community focusing on sanitation. Overall, income inequality affects who will have access to proper sanitation.
\paragraph{}
\citet{cassivi2018access} highlights about how there should be more information to the effect of including a 30-minute collection time to monitor access to drinking water. The lack of access to on-premises water sources can cause the use of greater distant alternative sources. Therefore causing the quantity of retrieved water to reduce as the water plateau phenomenon.\citet{cassivi2018access} hypothesizes that the households who must travel further are using unimproved sources. For our research, we can discover what modes of transportation people are traveling to retrieve the water and in what quantity depending on their means of transportation. Therefore, learning how to decrease the travel time of people to retrieve the water because according to the article, it can be a great burden and does not increase the quality of one's life to retrieve water daily and far away from their homes. 
\paragraph{}
\citet{tortajada2018achieving} highlights that sharing the immense improvement at the 2015 convention negatively affected many people because they actually have not acquired clean drinking water. While 2.6 billion people gained access to drinking water, they did not have constructive feedback on the quality of the water. Also, our research can fill in the blank on how did these people actually retrieve the water? Our research can measure the performance of how people retrieve water which may not be in their premises.  
\paragraph{}
\citet{amit2019measuring} highlights their data analysis similar to what we will be doing and found out that for non-piped households, collection costs are 22\% of the coping costs, while collection costs for piped households are less than 2\% of the coping costs. The researchers also realize that there is economic inequality because the majority of the piped households are in households that are considered wealthy households and vice versa. Our research can pinpoint how households are traveling to collect water to improve potential policy-making for water accessibility. 
\paragraph{}
\citet{price2019difference} highlights the temporal dynamics of drinking water access and quality in urban slums which is home to about 1 billion people globally. \citet{price2019difference} discusses the temporal changes in drinking water access and quality in urban slums may influence on health risks. \citet{price2019difference} believes that monitoring temporal dynamics should be considered over three interlinked time scales:
\begin{center}
\begin{tabular}{ c c c }
 short-term & medium term & long-term \\ 
  & &  \\  
  & &     
\end{tabular}
\end{center}
\citet{price2019difference} recommends future research to explore the short-term water access and quality monitoring. Also we should learn to recognize the role of socio-cultural factors that may impact the temporal dynamics of safe water access. Our project can pinpoint additional water accessibility variables that can be used to focus to recognize the time variations as well.  
\paragraph{}
\citet{rawas2020comparing} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru. This relates and informs one mode of transportation of clean and accessible water. Our research can make inferences about if the customers receive clean water elsewhere, such as water tankers, kiosks, vendors, etc. Overall, the article found that the percent of households with IWS was a few percentage points greater than of the reported utilities. 
\paragraph{}
\citet{deshpande2020mapping} discusses their findings of improved overall water availability and piped water accessibility for the globe (40\% to 50.3\%). \citet{deshpande2020mapping} also found that it was the lowest increase in sub-Saharan Africa, where the accessibility was mostly concentrated in urban regions. \citet{deshpande2020mapping} focuses mostly on understanding trends in diarrhea burden, where the needs are greatest for the improvement of access to safe drinking water and sanitation facilities. Our study can benefit this research by implementing the best and effective modes of transportation of clean water to improve safe drinking water and sanitation. Even with the water pipe access increasing, they can't conclude if the people had clean water as they needed right away. Overall, the access to safe drinking water and sanitation improved globally between 2000 and 2017, but dis-proportionality undermines reaching the SDG goals. 
\paragraph{}
\citet{goal} explains specific targets that have goals and indicators to track global water accessibility. The article quantifies and provides Graphs: safe drinking water, safe sanitation, and hygiene, safe sanitation and hygiene, ambient water quality, water use efficiency, levels of freshwater stress, integrated water management, trans-boundary water cooperation, protect and restore water-related ecosystems, water and sanitation support, local participation in sanitation management. They have many ambitious plans to achieve by 2020 and 2030, and our research can find a correlation to why achieving this may be possible.
% citep (for citations fully enclosed in parentheses)
% cite (for citations without any parentheses)
% citet (for citations where the year only is in parentheses)
%%\textit{Discuss the who/what/when/where? (You already have this... just need to reorganize or improve structure)}
%Another example is \citet{fuller2016tracking}.
%\citet{rawas2020comparing} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru.
%\citet{cassivi2018access} discuss time access...
%\hl{Perhaps insert a table summarizing the various dimensions of water accessibility that have been studied}
%Other examples: \citep{amit2019measuring,ho2014challenge}
\subsection{Gaps and areas for improvement}
We are interested in analyzing the impact of transportation modes (usage/ownership) on water accessibility.

\eject

\printbibliography

\section*{Data}
\hl{In this section, describe the data sources and the variables, along with data preparation procedures, variable
  selection criteria, etc. You should have a table summarizing the water accessibility variables and their
  definitions. Another table can summarize the explanatory variables (e.g. transportation-related, income-related,
  etc). Also discuss the number of countries, etc. I've started a table as an example. You can add other columns if needed.}

\begin{table}[h!]
  \centering
  \begin{tabular}{c c m{7cm}}\toprule
    \bf Variable & \bf Description & \bf Definition \\\midrule
    \texttt{ptap} & Households using a public tap/standpipe
 & Percentage of households whose main source of drinking water is a public tap/standpipe \\\hline
 \texttt{bore} & Households using a tube well/borehole & Percentage of households whose main source of drinking water is a tubewell/borehole
\\\hline
\texttt{pspr} & Households using a protected spring
& Percentage of households whose main source of drinking water is a protected spring
\\\hline
\texttt{uspr} & Households using an unprotected spring
& Percentage of households whose main source of drinking water is an unprotected spring \\\hline
\texttt{pwel} & Households using a protected well
& Percentage of households whose main source of drinking water is a protected well \\\hline
\texttt{uwel} & Households using an unprotected well water
&Percentage of households whose main source of drinking water is an unprotected well
\\\hline
\texttt{surw} & Households using surface water
& Percentage of households whose main source of drinking water is surface water
\\\hline
\texttt{tles} & Households with water 30 minutes or less away round trip
&Percentage of households with water 30 minutes or less away round trip
\\\hline 
\texttt{tmor } & Households with water more than 30 minutes away round trip
& Percentage of households with water more than 30 minutes away round trip
\\\hline
\texttt{rain } &Households using rainwater
& Percentage of households whose main source of drinking water is rainwater
\\\hline
\texttt{phom}&Households using water piped into dwelling
&Percentage of households whose main source of drinking water is water piped into the dwelling
\\\hline

& \\\bottomrule
  \end{tabular}
  \caption{Summary of water accessibility variables (national averages at household level)}
  \label{tab:summary}
\end{table}

\hl{1. We could also add the relevant data extracted from DHS showing what countries were also collected.
2. Show numbers because numbers tells great story.}
\paragraph{•}
\hl{ 3. Describe the 11 water accessibility variables, where it came from \citep{price2019difference}

\begin{table}[h!]
  \centering
  \begin{tabular}{c c m{7cm}}\toprule
    \bf Variable & \bf Description & \bf Definition \\\midrule
    \texttt{imws} & Households using an improved water source
 & Percentage of households whose main source of drinking water is an improved source
 \\\hline

 & \\\bottomrule
  \end{tabular}
  \caption{Summary of explanatory variables (national averages at household level)}
  \label{tab:summary}
\end{table}

\section*{Methods}
%\hl{Ultimately, there will be a preamble to summarize the contents of this section}
\paragraph{Preamble}
We analyzed data from the Demographic Health Survey (DHS) and the Multiple Indicator Cluster Survey (MICS) (1990-2020) to highlight water accessibility predictors in 78 countries throughout the world.

\subsection*{Cluster analysis}
\hl{You have all you need to write this section. 
You want to describe the algorithms for clustering, the indices for measuring cluster quality, and process you follow to choose the best method/cluster number.}

\hl{ 1. Description of the algorithms implemented: single, average, complete, ward.D, and ward.D2 }

\hl{ 2. Explanation of using NBClust function in R for ward.D2, ward.D, Complete, and K-means; these four methods were elected due their organization of the initial clustering trees.}

\hl{ 3. List of the optimal number of cluster values and the indices number of values for different alogrithm methods for the clustering methods. Description of how to determine the potential best clustering methodology. }

\hl { 4. Clustering optimization data share and explain: Ward.D2 chosen with 4 clustering tree methodology. Share the finished clustering graph for visualization.}

\subsection*{Classification model to explain water accessibility predictors}
\hl{This is the second phase of the work, which you should hopefully start in January}

\section*{Results and Discussion}
\subsection*{Summary of variables}
\hl{Here, you show and discuss the data distributions (e.g. histograms, correlation plots, etc)}

\subsection*{Water accessibility typologies}
\hl{Show the typology/cluster results and discuss.}

\subsection*{Model results}

  % \section*{Discussion}

\section*{Conclusion}

\printbibliography

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2021-01-14 docs/manuscript/manuscript.tex
updated manuscript

--- 2021-01-14 docs/manuscript/Global-Water-Access-v5-wo-comments.tex
@@
 %=== GRAPHICS PATH ===========
-%\graphicspath{{./140408-Images/}}
+\graphicspath{{./images/}}
 % Marginpar width
@@
 \usepackage[style=authoryear,backend=biber, natbib,maxcitenames=2,doi=false,isbn=false,url=false,eprint=false]{biblatex}
-\addbibresource{../References/Water-Accessibility.bib}
+\addbibresource{../Water-Accessibility.bib}
 
@@
 
-\paragraph{}
-\citet{onda2012global}discusses the global access to safe water by accounting for water quality and the resulting impact on MDG progress.
-In 2012, \citet{onda2012global} altered the current Joint monitoring programme (JMP) estimation by including for microbial water quality and sanitary risk utilizing national water quality data.
-\citep{onda2012global} used principal components analysis to analyze the national environment and development indicators and then created models.
-Overall, our analysis can highlight the potential fatigue illness caused by people transporting the water if it is done manually.
-How the modes of transportation can affect one's health.
-\paragraph{}
-\citet{sima2013water} highlights the relation of the topic where it discovered that over 76 percent of the city's water is utilized by less than ten percent of household piped water.
-The study is focused in Kisumu Kenya and Sima interviewed 260 informal water business operators.
-This article revealed a new insight where the majority of the population in Kisumu relied on purchased water from kiosks (1.5 million $m^{3}$ per day) and used hand-drawn water-carts (0.75 million $m^{3}$ per day)\citep{sima2013water}.
-Also, we discovered that the water trucking industry utilized the most energy consumption and was high cost: water delivery trucks have the highest per cubic meter energy demand (35 $MJ/m^{3}$)\citep{sima2013water}.
-The article utilized SPSS Version 19.0 to analyze the data.
-\paragraph{}
-\citet{ho2014challenge} highlights a head-to-head comparison of such indicators with other possible distance and time metrics among rural 1,103 households in Nampula province, Mozambique.
-The researchers found out that Euclidean line distance to be a good representative for route distance ($R^{2} = 0.98$), while self-reported travel time is a poor representative ($R^{2} = 0.12$).
-One of key insights was that, 15-minute decrease in one-way walk time to water source is related to a 41\% reduction in diarrhea, improved children's nutrition, and an 11\% reduction in under-five child mortality rate \citep{ho2014challenge}.
-This research can be a helpful tool to understand how long people are traveling in a rural area in Africa where many of our data focus.
-Also, understanding how people are traveling to receive the water can better understand the condition of the rural people and implement better policies that help them to thrive.
-\paragraph{}
-\citet{bartram2014globala} highlights a background information about the Millennium Development Goals.
-How it has established global targets for drinking water and sanitation access (Bartram et al.
-2014).
-The article indirectly gives much information about progress towards focused targets, directed by international monitoring, has reduced the global disease burden and improved the quality of life \citep{bartram2014globala}.
-It utilizes a variety of sources such as JMP, DHS, MICS surveys to analyze the data of many different locations in the world, therefore, it is a broad scope of study of the progress of accessible and drinking water and sanitation globally.
-Our research can study further what modes of transportation of clean drinking water would affect the progress of improving drinking water and sanitation in one's community.
-Also, understanding how long it takes to attain the water can improve the quality of the services provided for the people.
-\paragraph{}
-\citet{onda2014country} discusses their new development of typology of country clusters pertain to water and sanitation sector based on congruence through multiple related indicators.
-\citet{onda2014country} used a hierarchical clustering method and a gap statistic analysis to cluster 156 countries which has 6.75 billion people into a relational clusters.
-\citet{onda2014country} suggests previous geography or income-based country clustering should be improved by using water and sanitation related indicators.
-Our research can improvise upon their finding and focus on identifying which water accessibility variables may improve water accessibility for people who are seeking clean drinking water.
-\paragraph{}
-\citet{WHOdrinkingwater} article gives an ambitious plan to reach universal drinking water by the year of 2030 that is affordable and safe followed by the Sustainable Development Goal six goals.
-The WHO/UNICEF Joint Monitoring Program for Water Supply and Sanitation (JMP) are monitoring progress on drinking water and sanitation since 1990 globally.
-The article outlines different plans with consistent traits to determine what is deemed to be safe and affordable drinking water and good sanitation.
-Our study can highlight which modes of transportation of safe and affordable drinking water can develop why one's community may excel in the improvement of safe and affordable drinking water and sanitation.
-Drinking, sanitation, and hand washing ladder explains the qualitative determination can indirectly help what is a good source of transporting safe and drinking water.
-\paragraph{}
-\citet{fuller2016tracking} highlights the global accessibility to safe drinking water and sanitation and concludes that it has been improving well during the Millennium Development Goal Period.
-This topic will benefit our research to understand what countries are doing for the mode of transportation of drinking water and sanitation which may have contributed to the improvements.
-\citet{fuller2016tracking} chose some countries and studied the overall countries progress which in his methodology revealed sigmoidal trends.
-Our study can reveal if certain methods of transportation of drinking water and sanitation may have influenced the effectiveness of the improved drinking water met by MDG standards.
-\paragraph{}
-\citet{jia2016highresolution} highlights that there is inequality within safe sanitation.
-There is a great difference in wealth levels in many low-income countries.
-Therefore, this will hinder the improvement of MDG progress.
-\citet{jia2016highresolution}discusses future interventions to be engaged with different wealth categories.
-Our study can assist in understanding improved sanitation increase by making inferences about how they receive water may give the community focusing on sanitation.
-Overall, income inequality affects who will have access to proper sanitation.
-\paragraph{}
-\citet{cassivi2018access} highlights about how there should be more information to the effect of including a 30-minute collection time to monitor access to drinking water.
-The lack of access to on-premises water sources can cause the use of greater distant alternative sources.
-Therefore causing the quantity of retrieved water to reduce as the water plateau phenomenon.\citet{cassivi2018access} hypothesizes that the households who must travel further are using unimproved sources.
-For our research, we can discover what modes of transportation people are traveling to retrieve the water and in what quantity depending on their means of transportation.
-Therefore, learning how to decrease the travel time of people to retrieve the water because according to the article, it can be a great burden and does not increase the quality of one's life to retrieve water daily and far away from their homes.
-\paragraph{}
-\citet{tortajada2018achieving} highlights that sharing the immense improvement at the 2015 convention negatively affected many people because they actually have not acquired clean drinking water.
-While 2.6 billion people gained access to drinking water, they did not have constructive feedback on the quality of the water.
-Also, our research can fill in the blank on how did these people actually retrieve the water?
-Our research can measure the performance of how people retrieve water which may not be in their premises.
-\paragraph{}
-\citet{amit2019measuring} highlights their data analysis similar to what we will be doing and found out that for non-piped households, collection costs are 22\% of the coping costs, while collection costs for piped households are less than 2\% of the coping costs.
-The researchers also realize that there is economic inequality because the majority of the piped households are in households that are considered wealthy households and vice versa.
-Our research can pinpoint how households are traveling to collect water to improve potential policy-making for water accessibility.
-\paragraph{}
-\citet{price2019difference} highlights the temporal dynamics of drinking water access and quality in urban slums which is home to about 1 billion people globally.
-\citet{price2019difference} discusses the temporal changes in drinking water access and quality in urban slums may influence on health risks.
-\citet{price2019difference} believes that monitoring temporal dynamics should be considered over three interlinked time scales:
+\citet{onda2012global} discusses the global access to safe water by accounting for water quality and the resulting impact
+on MDG progress. They altered the current Joint monitoring programme (JMP) estimation by
+including for microbial water quality and sanitary risk utilizing national water quality data. \citet{onda2012global}
+used principal components analysis to analyze the national environment and development indicators and then created
+models. Overall, our analysis can highlight the potential fatigue illness caused by people transporting the water if it
+is done manually. How the modes of transportation can affect one's health.
+
+\citet{sima2013water} highlights the relation of the topic where it discovered that over 76 percent of the city's water
+is utilized by less than ten percent of household piped water.  The study is focused in Kisumu Kenya and Sima
+interviewed 260 informal water business operators. This article revealed a new insight where the majority of the
+population in Kisumu relied on purchased water from kiosks (1.5 million $m^{3}$ per day) and used hand-drawn water-carts
+(0.75 million $m^{3}$ per day)\citep{sima2013water}. Also, we discovered that the water trucking industry utilized the
+most energy consumption and was high cost: water delivery trucks have the highest per cubic meter energy demand (35
+$MJ/m^{3}$)\citep{sima2013water}. The article utilized SPSS Version 19.0 to analyze the data.
+
+\citet{ho2014challenge} highlights a head-to-head comparison of such indicators with other possible distance and time
+metrics among rural 1,103 households in Nampula province, Mozambique. The researchers found out that Euclidean line
+distance to be a good representative for route distance ($R^{2} = 0.98$), while self-reported travel time is a poor
+representative ($R^{2} = 0.12$). One of key insights was that, 15-minute decrease in one-way walk time to water source
+is related to a 41\% reduction in diarrhea, improved children's nutrition, and an 11\% reduction in under-five child
+mortality rate \citep{ho2014challenge}. This research can be a helpful tool to understand how long people are traveling
+in a rural area in Africa where many of our data focus. Also, understanding how people are traveling to receive the
+water can better understand the condition of the rural people and implement better policies that help them to thrive.
+
+\citet{bartram2014globala} highlights a background information about the Millennium Development Goals. How it has
+established global targets for drinking water and sanitation access (Bartram et al. 2014). The article indirectly gives
+much information about progress towards focused targets, directed by international monitoring, has reduced the global
+disease burden and improved the quality of life \citep{bartram2014globala}. It utilizes a variety of sources such as
+JMP, DHS, MICS surveys to analyze the data of many different locations in the world, therefore, it is a broad scope of
+study of the progress of accessible and drinking water and sanitation globally. Our research can study further what
+modes of transportation of clean drinking water would affect the progress of improving drinking water and sanitation in
+one's community. Also, understanding how long it takes to attain the water can improve the quality of the services
+provided for the people.
+
+\citet{onda2014country} discusses their new development of typology of country clusters pertain to water and sanitation
+sector based on congruence through multiple related indicators. \citet{onda2014country} used a hierarchical clustering
+method and a gap statistic analysis to cluster 156 countries which has 6.75 billion people into a relational
+clusters. \citet{onda2014country} suggests previous geography or income-based country clustering should be improved by
+using water and sanitation related indicators. Our research can improvise upon their finding and focus on identifying
+which water accessibility variables may improve water accessibility for people who are seeking clean drinking water.
+
+\citet{WHOdrinkingwater} article gives an ambitious plan to reach universal drinking water by the year of 2030 that is
+affordable and safe followed by the Sustainable Development Goal six goals. The WHO/UNICEF Joint Monitoring Program for
+Water Supply and Sanitation (JMP) are monitoring progress on drinking water and sanitation since 1990 globally. The
+article outlines different plans with consistent traits to determine what is deemed to be safe and affordable drinking
+water and good sanitation. Our study can highlight which modes of transportation of safe and affordable drinking water
+can develop why one's community may excel in the improvement of safe and affordable drinking water and
+sanitation. Drinking, sanitation, and hand washing ladder explains the qualitative determination can indirectly help
+what is a good source of transporting safe and drinking water.
+
+\citet{fuller2016tracking} highlights the global accessibility to safe drinking water and sanitation and concludes that
+it has been improving well during the Millennium Development Goal Period. This topic will benefit our research to
+understand what countries are doing for the mode of transportation of drinking water and sanitation which may have
+contributed to the improvements. \citet{fuller2016tracking} chose some countries and studied the overall countries
+progress which in his methodology revealed sigmoidal trends. Our study can reveal if certain methods of transportation
+of drinking water and sanitation may have influenced the effectiveness of the improved drinking water met by MDG
+standards.
+
+\citet{jia2016highresolution} highlights that there is inequality within safe sanitation. There is a great difference in
+wealth levels in many low-income countries. Therefore, this will hinder the improvement of MDG
+progress. \citet{jia2016highresolution}discusses future interventions to be engaged with different wealth
+categories. Our study can assist in understanding improved sanitation increase by making inferences about how they
+receive water may give the community focusing on sanitation. Overall, income inequality affects who will have access to
+proper sanitation.
+
+\citet{cassivi2018access} highlights about how there should be more information to the effect of including a 30-minute
+collection time to monitor access to drinking water. The lack of access to on-premises water sources can cause the use
+of greater distant alternative sources. Therefore causing the quantity of retrieved water to reduce as the water plateau
+phenomenon.\citet{cassivi2018access} hypothesizes that the households who must travel further are using unimproved
+sources. For our research, we can discover what modes of transportation people are traveling to retrieve the water and
+in what quantity depending on their means of transportation. Therefore, learning how to decrease the travel time of
+people to retrieve the water because according to the article, it can be a great burden and does not increase the
+quality of one's life to retrieve water daily and far away from their homes.
+
+\citet{tortajada2018achieving} highlights that sharing the immense improvement at the 2015 convention negatively
+affected many people because they actually have not acquired clean drinking water. While 2.6 billion people gained
+access to drinking water, they did not have constructive feedback on the quality of the water. Also, our research can
+fill in the blank on how did these people actually retrieve the water? Our research can measure the performance of how
+people retrieve water which may not be in their premises.
+
+\citet{amit2019measuring} highlights their data analysis similar to what we will be doing and found out that for
+non-piped households, collection costs are 22\% of the coping costs, while collection costs for piped households are
+less than 2\% of the coping costs. The researchers also realize that there is economic inequality because the majority
+of the piped households are in households that are considered wealthy households and vice versa. Our research can
+pinpoint how households are traveling to collect water to improve potential policy-making for water accessibility.
+
+\citet{price2019difference} highlights the temporal dynamics of drinking water access and quality in urban slums which
+is home to about 1 billion people globally. \citet{price2019difference} discusses the temporal changes in drinking water
+access and quality in urban slums may influence on health risks. \citet{price2019difference} believes that monitoring
+temporal dynamics should be considered over three interlinked time scales:
+
 \begin{center}
@@
 
-\hl{1. We could also add the relevant data extracted from DHS showing what countries were also collected.
-2. Show numbers because numbers tells great story.}
-\paragraph{•}
-\hl{ 3. Describe the 11 water accessibility variables, where it came from \citep{price2019difference}
-
 \begin{table}[h!]
@@
 \section*{Methods}
-%\hl{Ultimately, there will be a preamble to summarize the contents of this section}
+
 \paragraph{Preamble}
@@
 \subsection*{Cluster analysis}
-\hl{You have all you need to write this section. 
-You want to describe the algorithms for clustering, the indices for measuring cluster quality, and process you follow to choose the best method/cluster number.}
-
-\hl{ 1. Description of the algorithms implemented: single, average, complete, ward.D, and ward.D2 }
-
-\hl{ 2. Explanation of using NBClust function in R for ward.D2, ward.D, Complete, and K-means; these four methods were elected due their organization of the initial clustering trees.}
-
-\hl{ 3.
-List of the optimal number of cluster values and the indices number of values for different alogrithm methods for the clustering methods.
-Description of how to determine the potential best clustering methodology.
-}
-
-\hl { 4. Clustering optimization data share and explain: Ward.D2 chosen with 4 clustering tree methodology. Share the finished clustering graph for visualization.}
 
@@
 
+\begin{figure}[h!]
+  \centering
+  %\includegraphics[width=.7\textwidth]{spider-plot}
+  \caption{Spider plots of water accessibility variables by typology}
+  \label{fig:spider}
+\end{figure}
 \subsection*{Model results}

### 2021-01-15 docs/manuscript/manuscript.tex
Added Data for exploratory variables.

@@
 Our research can make inferences about if the customers receive clean water elsewhere, such as water tankers, kiosks, vendors, etc.
-Overall, the article found that the percent of households with IWS was a few percentage points greater than of the reported utilities.
+Overall, the article found that the percent of households with (IWS) was a few percentage points greater than of the reported utilities.
 \paragraph{}
@@
     \texttt{ptap} & Households using a public tap/standpipe
- & Percentage of households whose main source of drinking water is a public tap/standpipe \\\hline
- \texttt{bore} & Households using a tube well/borehole & Percentage of households whose main source of drinking water is a tubewell/borehole
+ & Percentage of households whose main source of drinking water is a public tap/standpipe.
+ \\\hline
+ \texttt{bore} & Households using a tube well/borehole & Percentage of households whose main source of drinking water is a tube-well/borehole.
 \\\hline
 \texttt{pspr} & Households using a protected spring
-& Percentage of households whose main source of drinking water is a protected spring
+& Percentage of households whose main source of drinking water is a protected spring.
 \\\hline
 \texttt{uspr} & Households using an unprotected spring
-& Percentage of households whose main source of drinking water is an unprotected spring \\\hline
+& Percentage of households whose main source of drinking water is an unprotected spring.
+\\\hline
 \texttt{pwel} & Households using a protected well
-& Percentage of households whose main source of drinking water is a protected well \\\hline
+& Percentage of households whose main source of drinking water is a protected well.
+\\\hline
 \texttt{uwel} & Households using an unprotected well water
-&Percentage of households whose main source of drinking water is an unprotected well
+&Percentage of households whose main source of drinking water is an unprotected well.
 \\\hline
 \texttt{surw} & Households using surface water
-& Percentage of households whose main source of drinking water is surface water
+& Percentage of households whose main source of drinking water is surface water.
 \\\hline
 \texttt{tles} & Households with water 30 minutes or less away round trip
-&Percentage of households with water 30 minutes or less away round trip
+&Percentage of households with water 30 minutes or less away round trip.
 \\\hline 
-\texttt{tmor } & Households with water more than 30 minutes away round trip
-& Percentage of households with water more than 30 minutes away round trip
+\texttt{tmor} & Households with water more than 30 minutes away round trip & Percentage of households with water more than 30 minutes away round trip.
 \\\hline
-\texttt{rain } &Households using rainwater
-& Percentage of households whose main source of drinking water is rainwater
+\texttt{rain} & Households using rainwater
+& Percentage of households whose main source of drinking water is rainwater.
 \\\hline
-\texttt{phom}&Households using water piped into dwelling
-&Percentage of households whose main source of drinking water is water piped into the dwelling
+\texttt{phom} & Households using water piped into dwelling
+&Percentage of households whose main source of drinking water is water piped into the dwelling.
 \\\hline
@@
     \texttt{imws} & Households using an improved water source
- & Percentage of households whose main source of drinking water is an improved source
+ & Percentage of households whose main source of drinking water is an improved source.
  \\\hline
-
+  \texttt{pipy} & Households using water piped into yard/plot
+& Percentage of households whose main source of drinking water is water piped into the yard or plot.
+\\\hline
+ \texttt{truc} & Households using tanker truck & Percentage of households whose main source of drinking water is tanker truck.
+\\\hline
+ \texttt{ctan} & Households using cart with tank & Percentage of households whose main source of drinking water is cart with tank.
+\\\hline
+ \texttt{bott} & Households using bottled water & Percentage of households whose main source of drinking water is bottled water/demi john.
+\\\hline
+ \texttt{uiws} & Households using an unimproved water source & Percentage of households whose main source of drinking water is an unimproved source.
+\\\hline
+ \texttt{othw} & Households using other water source & Percentage of households whose main source of drinking water is other.
+\\\hline
+ \texttt{iwsp} & Households with improved water source on the premises & Percentage of the de jure population living in households with an improved water source on the premises.
+\\\hline
+ \texttt{basw} & Households with basic water service
+ & Percentage of households with basic water service, defined as an improved water source with either water on the premises or round-trip collection time is 30 minutes or less.
+\\\hline
+ \texttt{liws} & Households with limited water service
+ & Percentage of households with limited water service, defined as an improved water source with round-trip collection time greater than 30 minutes.
+\\\hline
+ \texttt{prem} & Households with water on the premises
+ & Percentage of households with water on the premises.
+\\\hline
+ \texttt{bicy} & Households possessing a bicycle
+ & Percentage of households possessing a bicycle.
+\\\hline
+ \texttt{acar} & Households possessing an animal drawn cart
+ & Percentage of households possessing an animal drawn cart.
+\\\hline
+ \texttt{mcyc} & Households possessing a motorcycle
+ & Percentage of households possessing a motorcycle.
+\\\hline
+ \texttt{pcar} & Households possessing a private car
+ & Percentage of households possessing a private car.
+\\\hline
+ \texttt{boat} & Households possessing a boat with a motor
+ & Percentage of households possessing a boat with a motor.
+\\\hline
+ \texttt{wigc} & Wealth index Gini coefficient
+ & The Gini coefficient indicates the level of concentration of wealth in the country.
+\\\hline
  & \\\bottomrule

### 2021-01-15 docs/manuscript/manuscript.tex
Beginning to write the Data Procedure

@@
 temporal dynamics should be considered over three interlinked time scales:
-
 \begin{center}
@@
 Our project can pinpoint additional water accessibility variables that can be used to focus to recognize the time variations as well.
-\paragraph{}
 \citet{rawas2020comparing} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru.
@@
 Overall, the article found that the percent of households with (IWS) was a few percentage points greater than of the reported utilities.
-\paragraph{}
 \citet{deshpande2020mapping} discusses their findings of improved overall water availability and piped water accessibility for the globe (40\% to 50.3\%).
@@
 Even with the water pipe access increasing, they can't conclude if the people had clean water as they needed right away.
-Overall, the access to safe drinking water and sanitation improved globally between 2000 and 2017, but dis-proportionality undermines reaching the SDG goals.
-\paragraph{}
+Overall, the access to safe drinking water and sanitation improved globally between 2000 and 2017, but dis-proportionality undermines reaching the Sustainable Development Goals.
 \citet{goal} explains specific targets that have goals and indicators to track global water accessibility.
@@
 %Other examples: \citep{amit2019measuring,ho2014challenge}
-\subsection{Gaps and areas for improvement}
+\subsection*{Gaps and areas for improvement}
 We are interested in analyzing the impact of transportation modes (usage/ownership) on water accessibility.
@@
 \eject
-
-\printbibliography
 
@@
   etc). Also discuss the number of countries, etc. I've started a table as an example. You can add other columns if needed.}
+\subsection{Data Preparation}
+
+Data structure this study were collected from the Demographic Health Survey (DHS) program.
+We used the STATCompiler from the DHS Program to develop customized table based on foucsing on water, transportation, and wealth indicators across more than 90 countries and across all time frame.
+
+Both the waterpoint census and household surveys were undertaken by trained local enumerators and delivered in local languages, predominantly Swahili and Digo.
+Prior to data collection, research permits and approvals were obtained from the Government of Kenya's National Council of Science and Technology and the Central University Research Ethics Committee at the University of Oxford.
 

### 2021-01-15 docs/manuscript/manuscript.tex
Rough Draft of the Data Section.

--- 2021-01-15 docs/manuscript/manuscript.tex
@@
 \end{center}
+
 \citet{price2019difference} recommends future research to explore the short-term water access and quality monitoring.
@@
 Our project can pinpoint additional water accessibility variables that can be used to focus to recognize the time variations as well.
-\paragraph{}
+
 \citet{rawas2020comparing} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru.
@@
 Overall, the article found that the percent of households with (IWS) was a few percentage points greater than of the reported utilities.
-\paragraph{}
+
 \citet{deshpande2020mapping} discusses their findings of improved overall water availability and piped water accessibility for the globe (40\% to 50.3\%).
@@
 Even with the water pipe access increasing, they can't conclude if the people had clean water as they needed right away.
-Overall, the access to safe drinking water and sanitation improved globally between 2000 and 2017, but dis-proportionality undermines reaching the SDG goals.
-\paragraph{}
+Overall, the access to safe drinking water and sanitation improved globally between 2000 and 2017, but dis-proportionality undermines reaching the Sustainable Development Goals.
+
 \citet{goal} explains specific targets that have goals and indicators to track global water accessibility.
@@
 %Other examples: \citep{amit2019measuring,ho2014challenge}
-\subsection{Gaps and areas for improvement}
+\subsection*{Gaps and areas for improvement}
 We are interested in analyzing the impact of transportation modes (usage/ownership) on water accessibility.
+We also desire to see an implementation of this statistical research piloted in one of the countries to see if the water accessibility variables predictors can improve the water accessibility for its country.
 
 \eject
-
-\printbibliography
 
@@
   etc). Also discuss the number of countries, etc. I've started a table as an example. You can add other columns if needed.}
+\subsection*{Data Preparation}
+Data structure this study were collected from the Demographic Health Survey (DHS) program.
+We used the STATCompiler from the DHS Program to develop customized table based on focusing on water, transportation, and wealth indicators across 78 countries and from 1990 to 2019.
 
+We prepossess the data by differentiating the different surveys collected by the DHS program between DHS and Malaria Indicators Survey (MIS) through excel.
+After, we used R to develop and simplify the table with the most recent surveys for the most up to date analysis for the 78 countries.
+There was two sets of data of population and households.
+Therefore, we decided to remove the population data from the table to simplify because population and households data were identical.
+
+Next, there were many missing values from the data.
+Therefore, we had to do a simple missing value analysis.
+We determined to avoid using water, transportation, or wealth variables which had greater than 50\% missing values.
+Ultimately, simplified data frame was 76 countries by 28 variables.
+After, two different tables are generated where one focused on the water accessibility variables and a explanatory variables.
+We selected 11 water accessibility variables based on the sources available such as location, time, and methodology \citep{price2019difference}.
+The water accessibility variable was organized
+in the following table to allow for quick visualization for table 1. The other variables were used as an explanatory variables for determining the water accessibility predictors in Table 2. 
 \begin{table}[h!]
@@
 
-\paragraph{Preamble}
+\subsection*{Preamble}
 We analyzed data from the Demographic Health Survey (DHS) and the Multiple Indicator Cluster Survey (MICS) (1990-2020) to highlight water accessibility predictors in 78 countries throughout the world.
@@
 
+\section*{Limitation}
+%he problem of missing data is relatively common in almost all research and can have a significant effect on the conclusions that can be drawn from the data.
+
 \printbibliography

### 2021-01-19 docs/manuscript/manuscript.tex
Outline of the clustering method.

@@
 \subsection*{Cluster analysis}
+% 1. Clustering focusing on 11 water accessibility variables
+% 2. Different clustering algorithm for initial analysis
+% 3. Determination and analysis of best cut and hierarchical clustering method
+% 4. Election of the best clustering
+% 5. Analysis of Ward.d2 & 4 cut tree analysis
+
+We elected 11 variables which describes water accessibility of 78 countries around the world \citep{price2019difference}.
+Next, we decided to implement clustering which is an unsupervised machine learning method to identify and group water accessibility variables in large data set to understand and make inference about particular water accessibility variables.
+Clustering ultimately can continue to assist on discovering the water accessibility predictor variables.
+First, we
+
+%Kwale County is located on the south coast of Kenya, between the city of Mombasa to the north, and Tanzania to the south.
+In 2013, Kwale had a population of approximately 730,000 people, four‐fifths of whom lived in rural areas [Commission on Revenue Allocation, 2013].
+Based on a poverty gap index measure, the region is one of the most impoverished in Kenya, ranked 41st out of 47 counties [Commission on Revenue Allocation, 2013].
+According to the census in 2009, protected wells and boreholes were the main sources of water for 21.9% of the county's households [Kenya National Bureau of Statistics, 2012].
 

### 2021-01-19 docs/manuscript/manuscript.tex
Added Cassivi's 2021 journal article literature review.

@@
 They have many ambitious plans to achieve by 2020 and 2030, and our research can find a correlation to why achieving this may be possible.
+
+\citet{cassivi2021evaluating} focuses a case study in Southern Malawi, where data collected in March and April of 2019 at the conclusion of the rainy season.
+The sample was consisted of 375 households without access to water on their premises.
+The article highlighted using Euclidean Distance measurement is more reliable than the self-reported time.
+Also queuing time measurement can be hard to be measured because of their ambiguity.
+Her Research gives greater insight into how do we define water accessibility and how increase of water accessibility can impact people's lives beneficially.
 % citep (for citations fully enclosed in parentheses)
@@
 Next, we decided to implement clustering which is an unsupervised machine learning method to identify and group water accessibility variables in large data set to understand and make inference about particular water accessibility variables.
+All of the clustering analysis was assisted by using R program.
 Clustering ultimately can continue to assist on discovering the water accessibility predictor variables.
-First, we
 
-%Kwale County is located on the south coast of Kenya, between the city of Mombasa to the north, and Tanzania to the south.
-In 2013, Kwale had a population of approximately 730,000 people, four‐fifths of whom lived in rural areas [Commission on Revenue Allocation, 2013].
-Based on a poverty gap index measure, the region is one of the most impoverished in Kenya, ranked 41st out of 47 counties [Commission on Revenue Allocation, 2013].
-According to the census in 2009, protected wells and boreholes were the main sources of water for 21.9% of the county's households [Kenya National Bureau of Statistics, 2012].
+First, we used different clustering algorithm to determine best clustering results.
+We tested and categorized differently by using single, average, complete, ward.d, and ward.d2 methods to develop clustering dedrogram.
+Then, we analyzed by determine the best hierarchical relationship of each of the method by observing the tree diagram itself; if it has more of even distribution of each trees are the optimal clustering methods.
 

### 2021-01-19 docs/manuscript/manuscript.tex
{jagals2006does} highlights ... increased literature review.

@@
 
+\citet{jagals2006does} highlights in his case study in rural area in South Africa discovered providing small communities, using untreated river water as their water source, with good quality water with a piped distribution system and increase of accessibility by installing communal taps did not fall within parameters of safe water.
+Their research gives a potential understanding even making the drinking water accessible might not improve the quality of the water itself.
+
 \citet{onda2012global} discusses the global access to safe water by accounting for water quality and the resulting impact
@@
 Her Research gives greater insight into how do we define water accessibility and how increase of water accessibility can impact people's lives beneficially.
+
 % citep (for citations fully enclosed in parentheses)

### 2021-01-19 docs/manuscript/manuscript.tex
2nd paragraph for Clustering Analysis draft.

@@
  short-term & medium term & long-term \\
-  & &  \\  
-  & &     
+
 \end{tabular}
@@
 
-We elected 11 variables which describes water accessibility of 78 countries around the world \citep{price2019difference}.
-Next, we decided to implement clustering which is an unsupervised machine learning method to identify and group water accessibility variables in large data set to understand and make inference about particular water accessibility variables.
-All of the clustering analysis was assisted by using R program.
-Clustering ultimately can continue to assist on discovering the water accessibility predictor variables.
+We elected 11 variables that describe water accessibility of 78 countries around the world \citep{price2019difference}.
+Next, we decided to implement clustering which is an unsupervised machine learning method to identify and group water accessibility variables in large data set to understand and make inferences about particular water accessibility variables.
+All of the clustering analysis was assisted by using the R program.
+Clustering ultimately can continue to assist in discovering the water accessibility predictor variables.
 
-First, we used different clustering algorithm to determine best clustering results.
-We tested and categorized differently by using single, average, complete, ward.d, and ward.d2 methods to develop clustering dedrogram.
-Then, we analyzed by determine the best hierarchical relationship of each of the method by observing the tree diagram itself; if it has more of even distribution of each trees are the optimal clustering methods.
+First, we used a different clustering algorithm to determine the best clustering results.
+We tested and categorized differently by using single, average, complete, ward.d, and ward.d2 methods to develop clustering dendrogram.
+Then, we analyzed by determining the best hierarchical relationship of each of the methods by observing the tree diagram itself; if it has more of even distribution of each tree are the optimal clustering methods.
+We determined that ward.d2 had the best visual and even distribution, but to check this further, we used a function in R called 'NBClust' to determine the best algorithm as well as the best cut or groupings for the clustering methods.
 

### 2021-01-19 docs/manuscript/manuscript.tex
Third paragraph for clustering analysis.

@@
 
+We compared Ward.D2, Ward.D, complete, and k-means algorithm methods to determine the best cuts, and also to confirm that Ward.D2 was the best clustering algorithm method to develop the deprogram tree.
+The 'NBClust' function was set to determine the best clustering methodology by choosing minimum clustering from 3 to max cuts of 7 and utilized all the indices.
+The majority rule was implemented by the 'NBClust' function for Ward.d2, Ward.D, and the complete algorithm determined that the best clustering cut was four.
+However, k-means determined to be 3, but we determined that Ward.D2 is the best methodology thus far.
+
+Next, we compared the best indicies values for different algorthim solving methodlogy to determine the best cuts to comfirm that Ward.D2 is the best methodlogy.
+
 \subsection*{Classification model to explain water accessibility predictors}

### 2021-01-19 docs/manuscript/manuscript.tex
fourth paragraph rough draft complete in clustering analysis.

@@
 
-Next, we compared the best indicies values for different algorthim solving methodlogy to determine the best cuts to comfirm that Ward.D2 is the best methodlogy.
+Next, we compared the best indices values for different algorithm solving methodology to determine the best cuts to confirm that Ward.D2 is the best methodology.
+Once again, we determined that Ward.D2 was the best algorithm and learned that best cut is four trees by doing analysis within the different indices values of different algorithm theories.
+We compared the values by choosing KL, Ratkowsky, and Marriot values to compare what is the best cut for different clustering methods.
+
+\begin{table}[h!]
+  \centering
+  \begin{tabular}{c c m{7cm}}\toprule
+    \bf Variable & \bf Description & \bf Definition \\\midrule
+    \texttt{imws} 
+ & \\\bottomrule
+  \end{tabular}
+  \caption{Summary of explanatory variables (national averages at household level)}
+  \label{tab:summary}
+\end{table}
 

### 2021-01-19 docs/manuscript/manuscript.tex
Table 3. Summary of optimal clustering index values (Different Clustering Methods)

@@
 Next, we compared the best indices values for different algorithm solving methodology to determine the best cuts to confirm that Ward.D2 is the best methodology.
-Once again, we determined that Ward.D2 was the best algorithm and learned that best cut is four trees by doing analysis within the different indices values of different algorithm theories.
+Once again, we determined that Ward.D2 was the best algorithm and learned that the best cut is four trees by analyzing within the different indices values of different algorithm theories.
 We compared the values by choosing KL, Ratkowsky, and Marriot values to compare what is the best cut for different clustering methods.
@@
   \centering
-  \begin{tabular}{c c m{7cm}}\toprule
-    \bf Variable & \bf Description & \bf Definition \\\midrule
-    \texttt{imws} 
- & \\\bottomrule
+  \begin{tabular}{c c c}\toprule
+    \bf Methods & \bf KL (clusters) & \bf Value Index \\\midrule
+    \texttt{Ward.D2} & 4 & 2.4982 \\\hline
+    \texttt{Ward.D} & 4 & 3.3397 \\\hline
+    \texttt{Complete} & 4 & 3.9093 \\\hline
+     \texttt{K-Means} & 12 & 8.6836  \\\bottomrule
   \end{tabular}
-  \caption{Summary of explanatory variables (national averages at household level)}
+   \centering
+  \begin{tabular}{c c c}\toprule
+    \bf Methods & \bf Ratkowsky (clusters) & \bf Value Index \\\midrule
+    \texttt{Ward.D2} & 4 & 0.2981 \\\hline
+    \texttt{Ward.D} & 4 & 0.2981 \\\hline
+    \texttt{Complete} & 3 & 0.3193 \\\hline
+     \texttt{K-Means} & 3 & 0.2948 \\\bottomrule
+  \end{tabular}
+   \centering
+  \begin{tabular}{c c c c c}\toprule
+    \bf Methods & \bf Marriot(clusters) & \bf Value Index & \bf Gap & \bf Majority Rule  \\\midrule
+    \texttt{Ward.D2} & 4 & 1.54E+40 &  8 & 4 \\\hline
+    \texttt{Ward.D} & 4 & 1.79E+40 & 8 & 4 \\\hline
+    \texttt{Complete} & 4 & 1.77E+40  & 7 & 4 \\\hline
+     \texttt{K-Means} & 4 & 1.54E+40  & \texttt{N/A} & 3  \\\bottomrule
+  \end{tabular}
+  \caption{Summary of optimal clustering index values (Different Clustering Methods)}
   \label{tab:summary}

### 2021-01-19 docs/manuscript/manuscript.tex
Explanation of table 3.

@@
 First, we used a different clustering algorithm to determine the best clustering results.
-We tested and categorized differently by using single, average, complete, ward.d, and ward.d2 methods to develop clustering dendrogram.
+We tested and categorized differently by using single, average, complete, ward.D, and ward.D2 methods to develop clustering dendrogram.
 Then, we analyzed by determining the best hierarchical relationship of each of the methods by observing the tree diagram itself; if it has more of even distribution of each tree are the optimal clustering methods.
@@
 
+Table 3 is a rough summary of why Ward.D2 is the best clustering methodology if we go by four cuts.
+For the KL cluster algorithm, we determined that the best Value index is 2.4982 which is the lowest for the index, and the KL algorithm determines that the lowest value index results from the best cut of four.
+Also, the Ratkowsky cluster theorem determined that Ward.D2 was the best method for four cuts because the value index was the lowest matching Ward.D at .2981.
+However, Ward.D is the incomplete version of Ward.D2 where they forgot to square one of the intermediate value in the theorem.
+Therefore, Ward.D2 is the best clustering method for Ward.D2.
+Similarly, for the Marriot clustering theorem, Ward.D2 is the best once again due to the Ward.D having the greatest value index which determines that it is the best clustering method.
+Similar to an explanation from Ratkowsky, Ward.D2 is chosen to be the best method for 4 cuts.
+
 \subsection*{Classification model to explain water accessibility predictors}

### 2021-01-19 docs/manuscript/manuscript.tex
Last paragraph for Cluster Methodology section.

@@
 \subsection*{Cluster analysis}
-% 1. Clustering focusing on 11 water accessibility variables
-% 2. Different clustering algorithm for initial analysis
-% 3. Determination and analysis of best cut and hierarchical clustering method
-% 4. Election of the best clustering
-% 5. Analysis of Ward.d2 & 4 cut tree analysis
-
 We elected 11 variables that describe water accessibility of 78 countries around the world \citep{price2019difference}.
@@
 
+Lastly, we developed a horizontal dendrogram using R focusing on Ward.D2 methodology for clustering.
+We also cut the dendrogram in 4 trees as an 'NBClust' analysis.
+The clustering trees can be further be analyzed and the results and analysis will be discussed in the results and discussion section.
+
 \subsection*{Classification model to explain water accessibility predictors}

### 2021-01-19 docs/manuscript/manuscript.tex
Updates from Overleaf
spider-plot image inserted.
The first paragraph of Results and Discussions.

@@
 
+Across 78 countries in this study, the data provides 28 numerical variables and four descriptive variables.
+Majority of the data is from 2010 - 2020 (Figure 1).
+Therefore, we can analyze the global condition of water accessibility by utilizing the most up to date surveys.
+Most of the variables are related to water accessibility by their locations, time, and water resources.
+This also allows us to find water accessibility predictor variables and new clustering graphs.
+
+% Across all countries in this study, results show that up to 40% of the national population needs more than 30 minutes to fetch water irrespective of the type of water source used by the household (Table 2).
+In over half of the countries examined, more than one household out of four lives at over 30 minutes (round-trip) of a water source.
+About half of the countries show an average collection time higher than 30 minutes.
+The national average time to collect water is lowest in Madagascar, taking 14 minutes, and reaches 44 minutes in Uganda.
+Results from dis aggregation by type of source show that the average time to collect improved sources of water is lower or equivalent than for unimproved sources.
+These findings suggest that the households who must travel further are also often using unimproved sources.
+Standard deviations show an important variation in collection time within each country, which could demonstrate high inequalities in terms of access to water.
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{year}
+  \caption{Distribution of years of water accessibility variables}
+  \label{fig:year}
+\end{figure}
+\subsection*{Model results}
+
 \subsection*{Water accessibility typologies}
@@
   \centering
-  %\includegraphics[width=.7\textwidth]{spider-plot}
+  \includegraphics[width=.7\textwidth]{spider-plot}
   \caption{Spider plots of water accessibility variables by typology}

### 2021-01-20 docs/manuscript/manuscript.tex
Updates from Overleaf

--- 2021-01-19 docs/manuscript/manuscript.tex
@@
 
+Across 78 countries in this study, the data provides 28 numerical variables and four descriptive variables.
+Majority of the data is from 2010 - 2020 (Figure 1).
+Therefore, we can analyze the global condition of water accessibility by utilizing the most up to date surveys.
+Most of the variables are related to water accessibility by their locations, time, and water resources.
+This also allows us to find water accessibility predictor variables and new clustering graphs.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{year}
+  \caption{Distribution of years of water accessibility variables}
+  \label{fig:year}
+\end{figure}
+\subsection*{Model results}
+First, we developed a correlation chart to quickly analyze if there is any great correlation between all the 28 numerical variables that may pertain to water accessibility.
+
+begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{year}
+  \caption{Distribution of years of water accessibility variables}
+  \label{fig:year}
+\end{figure}
+
 \subsection*{Water accessibility typologies}

### 2021-01-20 docs/manuscript/manuscript.tex
correlation-basic figure 2 inserted.

@@
   \centering
-  \begin{tabular}{c c m{7cm}}\toprule
+  \begin{tabular}{c c m{5cm}}\toprule
     \bf Variable & \bf Description & \bf Definition \\\midrule
@@
 First, we developed a correlation chart to quickly analyze if there is any great correlation between all the 28 numerical variables that may pertain to water accessibility.
+We recognized the strong and weak relationship by the size and color of the correlation plot is shown in Figure 2, which indicates households with limited water service (liws) has strongest and positive relationship with Households possessing a bicycle (bicy).
+Meanwhile, households using an improved water source (imws) has the strongest and negative relationship with households using an unimproved water source (uiws).
+This initial finding results indicated that the initial variables are logical and can give us insight into further understanding and relationship between these variables.
 
-begin{figure}[h!]
+\begin{figure}[h!]
   \centering
-  \includegraphics[width=.7\textwidth]{year}
-  \caption{Distribution of years of water accessibility variables}
+  \includegraphics[width=.7\textwidth]{correlation-basic}
+  \caption{Correlation graph of numerical variables}
   \label{fig:year}

### 2021-01-20 docs/manuscript/manuscript.tex
Updates from Overleaf
adv-cor-plot inserted to discuss the correlation.
adv-cor-plot explnation paragraph.

@@
 \end{figure}
+As we continued to study further upon the correlations of the variables, we used more advanced correlation graphs to determined their p value and to see relationship between different categorical variables is shown in Figure 3, which again indicates greater detailed comparison between the variables.
+Households with improved water source on the premises (iwsp) has strong relationship with  households with basic water service (basw) of .84 and p value of less than 0.001.
+Also, we learned and discovered that Households using an unimproved water source (uiws) have a positive relationship with Households using an unprotected well water (uwel), Households using an unprotected spring (uspr), and Households using surface water (surw) with p value of less than 0.001 as well.
+These relationships justify the necessity to research further if the particular water accessibility variables can increase accessibility of particular water sources in order to reach the universal access to water by 2030, as proposed in the SDGs.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{adv-cor-plot}
+  \caption{Correlation graph of numerical variables}
+  \label{fig:year}
+\end{figure}
 

### 2021-01-20 docs/manuscript/manuscript.tex
Updates from Overleaf

@@
 \end{figure}
+Additionally, we wanted to explore if the modes of transportation variables have any relationship with potential water variables.
+Therefore, we investigated by developing more correlation plots relating transportation variables with water accessibility variables which is shown in Figure .
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{transportation-time}
+  \caption{Correlation graph of numerical variables}
+  \label{fig:year}
+\end{figure}
 

### 2021-01-20 docs/manuscript/manuscript.tex
Advanced Correlation Graph of Modes of Transportation and Time

@@
   \includegraphics[width=.7\textwidth]{correlation-basic}
-  \caption{Correlation graph of numerical variables}
+  \caption{Correlation Graph of Numerical Variables}
   \label{fig:year}
@@
   \includegraphics[width=.7\textwidth]{adv-cor-plot}
-  \caption{Correlation graph of numerical variables}
+  \caption{Advanced Correlation Graph of Numerical Variables}
   \label{fig:year}
@@
 Additionally, we wanted to explore if the modes of transportation variables have any relationship with potential water variables.
-Therefore, we investigated by developing more correlation plots relating transportation variables with water accessibility variables which is shown in Figure .
+Therefore, we investigated by developing more correlation plots relating transportation variables with water accessibility variables which is shown in Figure 4,5,6 .
 
@@
   \includegraphics[width=.7\textwidth]{transportation-time}
-  \caption{Correlation graph of numerical variables}
+  \caption{Advanced Correlation Graph of Modes of Transportation and Time}
   \label{fig:year}
@@
   \includegraphics[width=.7\textwidth]{spider-plot}
-  \caption{Spider plots of water accessibility variables by typology}
+  \caption{Spider Plot of Water Accessibility Variables by Typology}
   \label{fig:spider}

### 2021-01-20 docs/manuscript/manuscript.tex
Insertion of transportation and water resources advanced correlation plot.

@@
 Additionally, we wanted to explore if the modes of transportation variables have any relationship with potential water variables.
-Therefore, we investigated by developing more correlation plots relating transportation variables with water accessibility variables which is shown in Figure 4,5,6 .
+Therefore, we investigated by developing more correlation plots relating transportation variables with water accessibility variables which is shown in Figure 4,5,6.
 
@@
   \caption{Advanced Correlation Graph of Modes of Transportation and Time}
+  \label{fig:year}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{transportation-water-1}
+  \caption{Advanced Correlation Graph of Modes of Transportation and Water Resources}
+  \label{fig:year}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{transportation-water-2}
+  \caption{Advanced Correlation Graph of Modes of Transportation and Additional Water Resources}
   \label{fig:year}

### 2021-01-20 docs/manuscript/manuscript.tex
Explanation of figure 4.

@@
 Across 78 countries in this study, the data provides 28 numerical variables and four descriptive variables.
-Majority of the data is from 2010 - 2020 (Figure 1).
+The majority of the data is from 2010 - 2020 (Figure 1).
 Therefore, we can analyze the global condition of water accessibility by utilizing the most up to date surveys.
@@
 First, we developed a correlation chart to quickly analyze if there is any great correlation between all the 28 numerical variables that may pertain to water accessibility.
-We recognized the strong and weak relationship by the size and color of the correlation plot is shown in Figure 2, which indicates households with limited water service (liws) has strongest and positive relationship with Households possessing a bicycle (bicy).
+We recognized the strong and weak relationship by the size and color of the correlation plot is shown in Figure 2, which indicates households with limited water service (liws) has the strongest and positive relationship with Households possessing a bicycle (bicy).
 Meanwhile, households using an improved water source (imws) has the strongest and negative relationship with households using an unimproved water source (uiws).
-This initial finding results indicated that the initial variables are logical and can give us insight into further understanding and relationship between these variables.
+This initial finding results indicated that the initial variables are logical and can give us insight into further understanding and the relationship between these variables.
 
@@
 \end{figure}
-As we continued to study further upon the correlations of the variables, we used more advanced correlation graphs to determined their p value and to see relationship between different categorical variables is shown in Figure 3, which again indicates greater detailed comparison between the variables.
-Households with improved water source on the premises (iwsp) has strong relationship with  households with basic water service (basw) of .84 and p value of less than 0.001.
-Also, we learned and discovered that Households using an unimproved water source (uiws) have a positive relationship with Households using an unprotected well water (uwel), Households using an unprotected spring (uspr), and Households using surface water (surw) with p value of less than 0.001 as well.
-These relationships justify the necessity to research further if the particular water accessibility variables can increase accessibility of particular water sources in order to reach the universal access to water by 2030, as proposed in the SDGs.
+As we continued to study further upon the correlations of the variables, we used more advanced correlation graphs to determined their p-value, and to see a relationship between different categorical variables is shown in Figure 3, which again indicates a greater detailed comparison between the variables.
+Households with an improved water source on the premises (iwsp) have a strong relationship with households with basic water service (basw) of .84 and a p-value of less than 0.001.
+Also, we learned and discovered that Households using an unimproved water source (uiws) have a positive relationship with Households using unprotected well water (uwel), Households using an unprotected spring (uspr), and Households using surface water (surw) with a p-value of less than 0.001 as well.
+These relationships justify the necessity to research further if the particular water accessibility variables can increase the accessibility of particular water sources to reach universal access to water by 2030, as proposed in the SDGs.
 
@@
 Additionally, we wanted to explore if the modes of transportation variables have any relationship with potential water variables.
-Therefore, we investigated by developing more correlation plots relating transportation variables with water accessibility variables which is shown in Figure 4,5,6.
+Therefore, we investigated by developing more correlation plots relating transportation variables with water accessibility variables which are shown in Figure 4,5,6.
+There is a significant positive relationship between households possessing a private car (pcar) and households with water on the premises (watp) as shown in Figure 4.
+There is also is a significant negative relationship between households possessing a private car (pcar) with households with water 30 minutes or less away round trip (tles) and households with water more than 30 minutes away round trip (tmor) with a p-value of less than 0.001.
+This information is essential to investigate further why this is the global trend.
 

### 2021-01-20 docs/manuscript/manuscript.tex
Explanation of Figure 5 and Figure 6.

@@
 \end{figure}
+Figure 5 shows that Households possessing a motorcycle (mcyc) has a positive correlation with Households using a tube well/borehole (bore) and Households using rainwater (rain) with a p-value of less than 0.01.
+Similar to the previous analysis, there is a relationship between a mode of transportation to retrieve potential water sources.
+Figure 6 also shows that households possessing an animal-drawn cart (acar) have a positive and strong relationship with households using unprotected well water (uwel) with a p-value of less than 0.001.
+The results demonstrate that different modes of transportation factor into the accessibility of retrieving water sources on a global scale, therefore, there should be consideration of these variables to increase access to improved water sources.
+
+% The results demonstrate that when a threshold of 30 minutes is used, 1% to 27% of the population of these 17 countries would be considered as not having access to improved water sources.
 

### 2021-01-21 docs/manuscript/manuscript.tex
Methods - last paragraph.

@@
 The results demonstrate that different modes of transportation factor into the accessibility of retrieving water sources on a global scale, therefore, there should be consideration of these variables to increase access to improved water sources.
-
-% The results demonstrate that when a threshold of 30 minutes is used, 1% to 27% of the population of these 17 countries would be considered as not having access to improved water sources.
-
 \begin{figure}[h!]
@@
 
+There is a problem of people not available to retrieve water quickly and there is a collection burden which is necessary to increase the water quality to improve many people's health and quality of life \citep{cassivi2018access}.
+These initial findings can allow us to research into what are the major factors of water accessibility for the 76 countries by developing clustering and determining the water accessibility variable predictors.
+A future study could use the water accessibility variable predictors to implement policies and experiment if it improved water accessibility for the community.
+This would be a necessary implementation to invest to improve their access to water to reach universal access by 2030.
+
 \subsection*{Water accessibility typologies}
@@
 \end{figure}
+
 \subsection*{Model results}

### 2021-01-21 docs/manuscript/manuscript.tex
Updates from Overleaf
Edit of labels for figures.
Insertion of cluster-analysis figures.

@@
 Similar to an explanation from Ratkowsky, Ward.D2 is chosen to be the best method for 4 cuts.
+Figure 7 shows the visual representation of the explanation by the table 3.
 
@@
 Across 78 countries in this study, the data provides 28 numerical variables and four descriptive variables.
-The majority of the data is from 2010 - 2020 (Figure 1).
+The majority of the data is from 2010 - 2020 as shown in Figure 1.
 Therefore, we can analyze the global condition of water accessibility by utilizing the most up to date surveys.
@@
   \caption{Correlation Graph of Numerical Variables}
-  \label{fig:year}
+  \label{fig:cor-basic}
 \end{figure}
@@
   \caption{Advanced Correlation Graph of Numerical Variables}
-  \label{fig:year}
+  \label{fig:cor-plot}
 \end{figure}
@@
   \caption{Advanced Correlation Graph of Modes of Transportation and Time}
-  \label{fig:year}
+  \label{fig:transporation-time}
 \end{figure}
@@
   \caption{Advanced Correlation Graph of Modes of Transportation and Water Resources}
-  \label{fig:year}
+  \label{fig:transporation-water}
 \end{figure}
@@
   \caption{Advanced Correlation Graph of Modes of Transportation and Additional Water Resources}
-  \label{fig:year}
+  \label{fig:transporation-water}
 \end{figure}
@@
 
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{optimal-cuts}
+  \caption{Optimal Number of Clusters for Ward.D2}
+  \label{fig:optimal}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{centroid}
+  \caption{Centroid of the Four Cluster Groupings}
+  \label{fig:centroid}
+\end{figure}
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{typology}
+  \caption{Cluster Typology Average of Water Accessibility Variables }
+  \label{fig:typology}
+\end{figure}
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{dendrogram}
+  \caption{Cluster Dendrogram with Four Cuts}
+  \label{fig:dendrogram}
+\end{figure}
 \begin{figure}[h!]

### 2021-01-21 docs/manuscript/manuscript.tex
Beginning to write the words for the water access typologies section.

@@
 \hl{Show the typology/cluster results and discuss.}
+NBClust statistic analysis developed and determined four distinct groups.
+First cluster group composed of countries near the equator or in the Sub-Sahara Regions.
+The second analysis of the second group represents countries considered developed or recently industrialized.
+
+%Results A gap statistic analysis resulted first in two distinct groups,with the first being composed of countries typically considered developed or newly industrialized.
+The second analysis of the second group resulted in four country clusters.
+The greater subdivision would have resulted in statistically similar clusters, while greater aggregation would have resulted in highly internally heterogeneous clusters.
+The dendrogram resulting from this cluster analysis is shown in Fig.
+1, and the corresponding gap curve yielding four clusters is shown in Fig.
+2.
+Table 3 shows the resulting five water and sanitation country clusters.
+Fig.
+3 shows the cluster country membership on a world map.
+The cluster indicator variable means are summarized in Table 4.
+The Dunn and Davies-BouldinIndices for our clusters, as well as the World Bank, WHO, and MDGgroupings are shown in Table 5.
+Our clusters have both the largest Dunn index and smallest Davies-Bouldin index, indicating that they represent a superior grouping along with the chosen relevant indicators.
 

### 2021-01-21 bin/R/DHS-exploratory-analysis-v10.R
transfer-item-again

@@ -1,0 +1,533 @@
+## Collection of the library for the code ###
+library(readxl)
+library(RColorBrewer)
+library(corrplot)
+#install.packages("psych")
+#install.packages("PerformanceAnalytics")
+library(PerformanceAnalytics)
+library(psych)
+library(mvtnorm) # col.norm
+library(tilting) # col.norm
+library(ggplot2)
+library("np") #npreg: you may need to install it
+#install.packages(c("FactoMineR", "factoextra"))
+library("FactoMineR")
+library(tidyverse) #data Manipulation
+library(cluster) #Clustering algorithms
+library(factoextra)#Clustering Algorithms & Visualization
+library(gridExtra)
+#install.packages("dendextend")
+#install.packages("ggplot2")
+#install.packages("rlang")
+library(dendextend) # For fancy deprograms
+library(ggplot2) # ggplot
+library(rlang) # clustering cutting
+#install.packages('NbClust')
+library(NbClust)
+#installed.packages('fmsb')
+library(fmsb) #For Spider Plots
+### PRE-PROCESSING  ###
+df <- read_excel("Data/Cumulative_Data.xlsx",sheet=1)
+head(df)
+### Cleaning Column Years ###
+for (i in seq(1, nrow(df))) {
+    if (grepl("-", df[i,'Year'])) {
+        yy1 = substr(df[i,"Year"], start=1,stop=2)
+        yy2 = substr(df[i,"Year"], start=6,stop=7)
+        yy3 = paste(yy1,yy2,sep = "")
+        #yy3 = as.integer(yy3)
+        df[i,"Year"] = yy3
+    }
+    }
+### Finding the Max Years ###
+df$Year = as.numeric(df$Year)
+df1 <- matrix(NA, nrow = 0, ncol = length(colnames(df))) #create new empty DF
+colnames(df1) = colnames(df) # assign column names to empty data frame
+for (country in unique(df$Country)) {
+    maxyear = max(df[df$Country==country,'Year'])
+    df1 = rbind( df1, df[(df$Country==country) & (df$Year==maxyear),])
+}        
+### Removing Population Variables (Removal of duplication) ###
+dfh = df1[,!grepl("^P_",names(df1))]
+### NA and zeros Analysis ###
+na_count = colSums(is.na(dfh)) #Counting all your NA in each data frames
+na_count
+### Removing column for household data frame if NA is greater than 50% ###
[…]

### 2021-01-21 docs/manuscript/manuscript.tex
Updates from Overleaf

@@
   \centering
-  \begin{tabular}{c c m{5cm}}\toprule
+  \begin{tabular}{p{1in} p{3in} p{2in}}\toprule
     \bf Variable & \bf Description & \bf Definition \\\midrule
@@
  & Percentage of households whose main source of drinking water is a public tap/standpipe.
- \\\hline
+ \\\midrule
  \texttt{bore} & Households using a tube well/borehole & Percentage of households whose main source of drinking water is a tube-well/borehole.
-\\\hline
+\\\midrule
 \texttt{pspr} & Households using a protected spring
 & Percentage of households whose main source of drinking water is a protected spring.
-\\\hline
+\\\midrule
 \texttt{uspr} & Households using an unprotected spring
 & Percentage of households whose main source of drinking water is an unprotected spring.
-\\\hline
+\\\midrule
 \texttt{pwel} & Households using a protected well
 & Percentage of households whose main source of drinking water is a protected well.
-\\\hline
+\\\midrule
 \texttt{uwel} & Households using an unprotected well water
 &Percentage of households whose main source of drinking water is an unprotected well.
-\\\hline
+\\\midrule
 \texttt{surw} & Households using surface water
 & Percentage of households whose main source of drinking water is surface water.
-\\\hline
+\\\midrule
 \texttt{tles} & Households with water 30 minutes or less away round trip
 &Percentage of households with water 30 minutes or less away round trip.
-\\\hline 
+\\\midrule 
 \texttt{tmor} & Households with water more than 30 minutes away round trip & Percentage of households with water more than 30 minutes away round trip.
-\\\hline
+\\\midrule
 \texttt{rain} & Households using rainwater
 & Percentage of households whose main source of drinking water is rainwater.
-\\\hline
+\\\midrule
 \texttt{phom} & Households using water piped into dwelling
 &Percentage of households whose main source of drinking water is water piped into the dwelling.
-\\\hline
+\\\midrule
 
@@
   \end{tabular}
-  \caption{Summary of explanatory variables (national averages at household level)}
+  \caption{Summary of explanatory variables (national averages measured as percentage at household level)}
   \label{tab:summary}
 \end{table}
+ \subsection*{Summary of variables}
+\hl{Here, you show and discuss the data distributions (e.g. histograms, correlation plots, etc)}
+
+Across 78 countries in this study, the data provides 28 numerical variables and four descriptive variables.
+The majority of the data is from 2010 - 2020 as shown in \autoref{fig:year}.
+Therefore, we can analyze the global condition of water accessibility by utilizing the most up to date surveys.
+Most of the variables are related to water accessibility by their locations, time, and water resources.
+This also allows us to find water accessibility predictor variables and new clustering graphs.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{year}
+  \caption{Distribution of years of water accessibility variables}
+  \label{fig:year}
+\end{figure}
 
@@
 \section*{Results and Discussion}
-\subsection*{Summary of variables}
-\hl{Here, you show and discuss the data distributions (e.g. histograms, correlation plots, etc)}
 
-Across 78 countries in this study, the data provides 28 numerical variables and four descriptive variables.
-The majority of the data is from 2010 - 2020 as shown in Figure 1.
-Therefore, we can analyze the global condition of water accessibility by utilizing the most up to date surveys.
-Most of the variables are related to water accessibility by their locations, time, and water resources.
-This also allows us to find water accessibility predictor variables and new clustering graphs.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=.7\textwidth]{year}
-  \caption{Distribution of years of water accessibility variables}
-  \label{fig:year}
-\end{figure}
 \subsection*{Model results}

### 2021-01-21 docs/manuscript/manuscript.tex
MICS -> MIS.

@@
 \subsection*{Preamble}
-We analyzed data from the Demographic Health Survey (DHS) and the Multiple Indicator Cluster Survey (MICS) (1990-2020) to highlight water accessibility predictors in 78 countries throughout the world.
+We analyzed data from the Demographic Health Survey (DHS) and the Malaria Indicator Survey (MIS) (1990-2020) to highlight water accessibility predictors in 78 countries throughout the world.
 

### 2021-01-22 docs/manuscript/manuscript.tex

@@
   \caption{Summary of water accessibility variables (national averages at household level)}
-  \label{tab:summary}
+  \label{tab:summary-wa}
 \end{table}
@@
   \caption{Summary of explanatory variables (national averages measured as percentage at household level)}
-  \label{tab:summary}
+  \label{tab:summary-var}
 \end{table}
@@
   \caption{Summary of optimal clustering index values (Different Clustering Methods)}
-  \label{tab:summary}
+  \label{tab:summary-optimal}
 \end{table}
@@
   \caption{Advanced Correlation Graph of Modes of Transportation and Water Resources}
-  \label{fig:transporation-water}
+  \label{fig:transporation-water-one}
 \end{figure}
@@
   \caption{Advanced Correlation Graph of Modes of Transportation and Additional Water Resources}
-  \label{fig:transporation-water}
+  \label{fig:transporation-water-two}
 \end{figure}
@@
 First cluster group composed of countries near the equator or in the Sub-Sahara Regions.
-The second analysis of the second group represents countries considered developed or recently industrialized.
+The analysis of the second group represents countries considered developed or recently industrialized.
+Third cluster group represents most of the countries in the central Africa and may have similar factors as the first cluster group.
+The fourth cluster group represents most of countries in the middle east or predominately non-African countries.
+There is a great internally heterogeneous clusters for the 76 distinct countries.
+The cluster analysis' dendrogram is shown in \autoref{fig:dendrogram}, and the corresponding NBClust curve yielding the optimal four clusters is shown in \autoref{fig:optimal-clusters, fig:optimal-cuts}.
 
-%Results A gap statistic analysis resulted first in two distinct groups,with the first being composed of countries typically considered developed or newly industrialized.
-The second analysis of the second group resulted in four country clusters.
-The greater subdivision would have resulted in statistically similar clusters, while greater aggregation would have resulted in highly internally heterogeneous clusters.
-The dendrogram resulting from this cluster analysis is shown in Fig.
-1, and the corresponding gap curve yielding four clusters is shown in Fig.
-2.
-Table 3 shows the resulting five water and sanitation country clusters.
+%   Table 3 shows the resulting five water and sanitation country clusters.
 Fig.
@@
 Our clusters have both the largest Dunn index and smallest Davies-Bouldin index, indicating that they represent a superior grouping along with the chosen relevant indicators.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{optimal-clusters-ward-d2}
+  \caption{Optimal Number of Clusters for Ward.D2}
+  \label{fig:optimal-clusters}
+\end{figure}
 

### 2021-01-22 docs/manuscript/manuscript.tex
First Paragraph of Water Accessibility typology.

@@
 There is a great internally heterogeneous clusters for the 76 distinct countries.
-The cluster analysis' dendrogram is shown in \autoref{fig:dendrogram}, and the corresponding NBClust curve yielding the optimal four clusters is shown in \autoref{fig:optimal-clusters, fig:optimal-cuts}.
+The cluster analysis' dendrogram is shown in \autoref{fig:dendrogram}, and the corresponding NBClust curve yielding the optimal four clusters is shown in \autoref{fig:optimal-clusters}, \autoref{fig:optimal}.
+\autoref{}
 

### 2021-01-22 docs/manuscript/manuscript.tex
Addition of figure hyper link labels.

@@
 The water accessibility variable was organized
-in the following table to allow for quick visualization for table 1. The other variables were used as an explanatory variables for determining the water accessibility predictors in Table 2. 
+in the following table to allow for quick visualization for \autoref{tab:summary-wa}.
+The other variables were used as an explanatory variables for determining the water accessibility predictors in \autoref{tab:summary-var}.
 \begin{table}[h!]
@@
  & Percentage of households whose main source of drinking water is an improved source.
- \\\hline
+ \\\midrule
   \texttt{pipy} & Households using water piped into yard/plot
 & Percentage of households whose main source of drinking water is water piped into the yard or plot.
-\\\hline
+\\\midrule
  \texttt{truc} & Households using tanker truck & Percentage of households whose main source of drinking water is tanker truck.
-\\\hline
+\\\midrule
  \texttt{ctan} & Households using cart with tank & Percentage of households whose main source of drinking water is cart with tank.
-\\\hline
+\\\midrule
  \texttt{bott} & Households using bottled water & Percentage of households whose main source of drinking water is bottled water/demi john.
-\\\hline
+\\\midrule
  \texttt{uiws} & Households using an unimproved water source & Percentage of households whose main source of drinking water is an unimproved source.
-\\\hline
+\\\midrule
  \texttt{othw} & Households using other water source & Percentage of households whose main source of drinking water is other.
-\\\hline
+\\\midrule
  \texttt{iwsp} & Households with improved water source on the premises & Percentage of the de jure population living in households with an improved water source on the premises.
-\\\hline
+\\\midrule
  \texttt{basw} & Households with basic water service
  & Percentage of households with basic water service, defined as an improved water source with either water on the premises or round-trip collection time is 30 minutes or less.
-\\\hline
+\\\midrule
  \texttt{liws} & Households with limited water service
  & Percentage of households with limited water service, defined as an improved water source with round-trip collection time greater than 30 minutes.
-\\\hline
+\\\midrule
  \texttt{prem} & Households with water on the premises
  & Percentage of households with water on the premises.
-\\\hline
+\\\midrule
  \texttt{bicy} & Households possessing a bicycle
  & Percentage of households possessing a bicycle.
-\\\hline
+\\\midrule
  \texttt{acar} & Households possessing an animal drawn cart
  & Percentage of households possessing an animal drawn cart.
-\\\hline
+\\\midrule
  \texttt{mcyc} & Households possessing a motorcycle
  & Percentage of households possessing a motorcycle.
-\\\hline
+\\\midrule
  \texttt{pcar} & Households possessing a private car
  & Percentage of households possessing a private car.
-\\\hline
+\\\midrule
  \texttt{boat} & Households possessing a boat with a motor
  & Percentage of households possessing a boat with a motor.
-\\\hline
+\\\midrule
  \texttt{wigc} & Wealth index Gini coefficient
  & The Gini coefficient indicates the level of concentration of wealth in the country.
-\\\hline
+\\\midrule
  & \\\bottomrule
@@
 Then, we analyzed by determining the best hierarchical relationship of each of the methods by observing the tree diagram itself; if it has more of even distribution of each tree are the optimal clustering methods.
-We determined that ward.d2 had the best visual and even distribution, but to check this further, we used a function in R called 'NBClust' to determine the best algorithm as well as the best cut or groupings for the clustering methods.
+We determined that ward.d2 had the best visual and even distribution, but to check this further, we used a function in R called 'NBClust' to determine the best algorithm as well as the best cut or groupings for the clustering methods as shown in \autoref{fig:optimal-clusters}.
 
@@
 Once again, we determined that Ward.D2 was the best algorithm and learned that the best cut is four trees by analyzing within the different indices values of different algorithm theories.
-We compared the values by choosing KL, Ratkowsky, and Marriot values to compare what is the best cut for different clustering methods.
+We compared the values by choosing KL, Ratkowsky, and Marriot values to compare what is the best cut for different clustering methods as shown in \autoref{fig:optimal}.
 
@@
   \centering
-  \begin{tabular}{c c c}\toprule
+  \begin{tabular}{p{1in} p{1in} p{1in}}\toprule
     \bf Methods & \bf KL (clusters) & \bf Value Index \\\midrule
-    \texttt{Ward.D2} & 4 & 2.4982 \\\hline
-    \texttt{Ward.D} & 4 & 3.3397 \\\hline
-    \texttt{Complete} & 4 & 3.9093 \\\hline
+    \texttt{Ward.D2} & 4 & 2.4982 \\\midrule
+    \texttt{Ward.D} & 4 & 3.3397 \\\midrule
+    \texttt{Complete} & 4 & 3.9093 \\\midrule
      \texttt{K-Means} & 12 & 8.6836  \\\bottomrule
@@
    \centering
-  \begin{tabular}{c c c}\toprule
+  \begin{tabular}{p{1in} p{1in} p{1in}}\toprule
     \bf Methods & \bf Ratkowsky (clusters) & \bf Value Index \\\midrule
-    \texttt{Ward.D2} & 4 & 0.2981 \\\hline
-    \texttt{Ward.D} & 4 & 0.2981 \\\hline
-    \texttt{Complete} & 3 & 0.3193 \\\hline
+    \texttt{Ward.D2} & 4 & 0.2981 \\\midrule
+    \texttt{Ward.D} & 4 & 0.2981 \\\midrule
+    \texttt{Complete} & 3 & 0.3193 \\\midrule
      \texttt{K-Means} & 3 & 0.2948 \\\bottomrule
@@
    \centering
-  \begin{tabular}{c c c c c}\toprule
+  \begin{tabular}{p{1in} p{1in} p{1in} p{1in} p{1in}}\toprule
     \bf Methods & \bf Marriot(clusters) & \bf Value Index & \bf Gap & \bf Majority Rule  \\\midrule
-    \texttt{Ward.D2} & 4 & 1.54E+40 &  8 & 4 \\\hline
-    \texttt{Ward.D} & 4 & 1.79E+40 & 8 & 4 \\\hline
-    \texttt{Complete} & 4 & 1.77E+40  & 7 & 4 \\\hline
+    \texttt{Ward.D2} & 4 & 1.54E+40 &  8 & 4 \\\midrule
+    \texttt{Ward.D} & 4 & 1.79E+40 & 8 & 4 \\\midrule
+    \texttt{Complete} & 4 & 1.77E+40  & 7 & 4 \\\midrule
      \texttt{K-Means} & 4 & 1.54E+40  & \texttt{N/A} & 3  \\\bottomrule
@@
 
-Table 3 is a rough summary of why Ward.D2 is the best clustering methodology if we go by four cuts.
+\autoref{tab:summary-optimal} is a rough summary of why Ward.D2 is the best clustering methodology if we go by four cuts.
 For the KL cluster algorithm, we determined that the best Value index is 2.4982 which is the lowest for the index, and the KL algorithm determines that the lowest value index results from the best cut of four.
@@
 Similar to an explanation from Ratkowsky, Ward.D2 is chosen to be the best method for 4 cuts.
-Figure 7 shows the visual representation of the explanation by the table 3.
+\autoref{fig:optimal} shows the visual representation of the explanation by the \autoref{tab:summary-optimal}.
 
 Lastly, we developed a horizontal dendrogram using R focusing on Ward.D2 methodology for clustering.
-We also cut the dendrogram in 4 trees as an 'NBClust' analysis.
+We also cut the dendrogram in 4 trees as an 'NBClust' analysis as shown in \autoref{fig:dendrogram}.
 The clustering trees can be further be analyzed and the results and analysis will be discussed in the results and discussion section.
@@
 First, we developed a correlation chart to quickly analyze if there is any great correlation between all the 28 numerical variables that may pertain to water accessibility.
-We recognized the strong and weak relationship by the size and color of the correlation plot is shown in Figure 2, which indicates households with limited water service (liws) has the strongest and positive relationship with Households possessing a bicycle (bicy).
+We recognized the strong and weak relationship by the size and color of the correlation plot is shown in \autoref{fig:cor-basic}, which indicates households with limited water service (liws) has the strongest and positive relationship with Households possessing a bicycle (bicy).
 Meanwhile, households using an improved water source (imws) has the strongest and negative relationship with households using an unimproved water source (uiws).
@@
 \end{figure}
-As we continued to study further upon the correlations of the variables, we used more advanced correlation graphs to determined their p-value, and to see a relationship between different categorical variables is shown in Figure 3, which again indicates a greater detailed comparison between the variables.
+As we continued to study further upon the correlations of the variables, we used more advanced correlation graphs to determined their p-value, and to see a relationship between different categorical variables is shown in \autoref{fig:cor-plot}, which again indicates a greater detailed comparison between the variables.
 Households with an improved water source on the premises (iwsp) have a strong relationship with households with basic water service (basw) of .84 and a p-value of less than 0.001.
@@
 Additionally, we wanted to explore if the modes of transportation variables have any relationship with potential water variables.
-Therefore, we investigated by developing more correlation plots relating transportation variables with water accessibility variables which are shown in Figure 4,5,6.
-There is a significant positive relationship between households possessing a private car (pcar) and households with water on the premises (watp) as shown in Figure 4.
+Therefore, we investigated by developing more correlation plots relating transportation variables with water accessibility variables which are shown in \autoref{fig:transporation-time}, \autoref{fig:transporation-water-one}, \autoref{fig:transporation-water-two}.
+There is a significant positive relationship between households possessing a private car (pcar) and households with water on the premises (watp) as shown in \autoref{fig:transporation-time}.
 There is also is a significant negative relationship between households possessing a private car (pcar) with households with water 30 minutes or less away round trip (tles) and households with water more than 30 minutes away round trip (tmor) with a p-value of less than 0.001.
@@
 \end{figure}
-Figure 5 shows that Households possessing a motorcycle (mcyc) has a positive correlation with Households using a tube well/borehole (bore) and Households using rainwater (rain) with a p-value of less than 0.01.
+\autoref{fig:transporation-water-one} shows that Households possessing a motorcycle (mcyc) has a positive correlation with Households using a tube well/borehole (bore) and Households using rainwater (rain) with a p-value of less than 0.01.
 Similar to the previous analysis, there is a relationship between a mode of transportation to retrieve potential water sources.
-Figure 6 also shows that households possessing an animal-drawn cart (acar) have a positive and strong relationship with households using unprotected well water (uwel) with a p-value of less than 0.001.
+\autoref{fig:transporation-water-two} also shows that households possessing an animal-drawn cart (acar) have a positive and strong relationship with households using unprotected well water (uwel) with a p-value of less than 0.001.
 The results demonstrate that different modes of transportation factor into the accessibility of retrieving water sources on a global scale, therefore, there should be consideration of these variables to increase access to improved water sources.

### 2021-01-22 docs/manuscript/manuscript.tex
Edit of labels for the table.

@@
     \bf Variable & \bf Description & \bf Definition \\\midrule
-    \texttt{imws} & Households using an improved water source
- & Percentage of households whose main source of drinking water is an improved source.
+    \texttt{imws} & Improved water source
+ & Drinking water is an improved source.
  \\\midrule
-  \texttt{pipy} & Households using water piped into yard/plot
-& Percentage of households whose main source of drinking water is water piped into the yard or plot.
+  \texttt{pipy} & Water piped into yard/plot
+& Drinking water is water piped into the yard or plot.
 \\\midrule
- \texttt{truc} & Households using tanker truck & Percentage of households whose main source of drinking water is tanker truck.
+ \texttt{truc} & Tanker truck & Drinking water is tanker truck.
 \\\midrule
- \texttt{ctan} & Households using cart with tank & Percentage of households whose main source of drinking water is cart with tank.
+ \texttt{ctan} & Cart with tank & Drinking water is cart with tank.
 \\\midrule
- \texttt{bott} & Households using bottled water & Percentage of households whose main source of drinking water is bottled water/demi john.
+ \texttt{bott} & Bottled water & Drinking water is bottled water/demi john.
 \\\midrule
- \texttt{uiws} & Households using an unimproved water source & Percentage of households whose main source of drinking water is an unimproved source.
+ \texttt{uiws} & Unimproved water source & Drinking water is an unimproved source.
 \\\midrule
- \texttt{othw} & Households using other water source & Percentage of households whose main source of drinking water is other.
+ \texttt{othw} & Other water source & Drinking water is other.
 \\\midrule
- \texttt{iwsp} & Households with improved water source on the premises & Percentage of the de jure population living in households with an improved water source on the premises.
+ \texttt{iwsp} & Improved water source on the premises & The de jure population living in households with an improved water source on the premises.
 \\\midrule
- \texttt{basw} & Households with basic water service
- & Percentage of households with basic water service, defined as an improved water source with either water on the premises or round-trip collection time is 30 minutes or less.
+ \texttt{basw} & Basic water service
+ & Defined as an improved water source with either water on the premises or round-trip collection time is 30 minutes or less.
 \\\midrule
@@
   \end{tabular}
-  \caption{Summary of explanatory variables (national averages measured as percentage at household level)}
+  \caption{Summary of explanatory variables (National Household usage Averages measured as percentage for Main Source)}
   \label{tab:summary-var}

### 2021-01-25 docs/manuscript/manuscript.tex
manu updates

@@
 \section{Introduction}
-%%(Roughly 3 paragraphs)
-In 2017, 29\% of the global population (2.2 billion people) did not use a safely managed drinking - water service which is defined as water located on-premise, available, and free from contamination \citep{WHOdrinkingwater}.
+In 2017, 29\% of the global population (2.2 billion people) did not use a safely managed drinking-water service, which is defined as water located on-premises, available, and free from contamination \citep{WHOdrinkingwater}.
 If you meet about 100 people, you can imagine that 29 of the people you meet do not have access to improved drinking water.
-Another recognition is that many people spend a lot of time collecting water where their access to drinking water
-supplies located on-premises are not common, about 26.3 percent of the world's population
-\citep{cassivi2018access}. Where this raises concerns about inequality problems related to the task and time variation 
+Where this raises concerns about inequality problems related to the task and time variation
 to collect water can be inferred in the urban and rural areas \citep{cassivi2018access}. Therefore, there needs to be a
@@
 
-Globally, access to improved drinking water sources and sanitation is increasing. From 1990 to 2012, utilization of
+Globally, access to improved drinking water sources and sanitation is increasing. From 1990 to 2012, the utilization of
 improved drinking water sources has increased globally from 76 percent to 89 percent, and utilization of improved
 sanitation has increased from 45 percent to 64 percent \citep{fuller2016tracking}. According to the MDG categorization,
-they consider improved drinking water sources as following: public tap, borehole, protected spring, rainwater
+they consider improved drinking water sources as the following: public tap, borehole, protected spring, rainwater
 collection, bottled water source, and piped water \citep{bartram2014globala}. However, these variables do not truly
-explain if the households have speedy access to retrieve the water. Therefore, we will investigate further on what
+explain if the households have speedy access to retrieve the water. In this paper, we investigate further on what
 variables should be included to enhance the accessibility of improved drinking water.
@@
 \section{Relevant work}
-Organized by chronology and the Categorical.
 
-\citet{jagals2006does} highlights in his case study in rural area in South Africa discovered providing small communities, using untreated river water as their water source, with good quality water with a piped distribution system and increase of accessibility by installing communal taps did not fall within parameters of safe water.
-Their research gives a potential understanding even making the drinking water accessible might not improve the quality of the water itself.
+\citet{jagals2006does} highlights in a case study in rural area in South Africa discovered providing small
+communities, using untreated river water as their water source, with good quality water with a piped distribution system
+and increase of accessibility by installing communal taps did not fall within parameters of safe water. Their research
+gives a potential understanding even making the drinking water accessible might not improve the quality of the water
+itself.
 
-\citet{onda2012global} discusses the global access to safe water by accounting for water quality and the resulting impact
-on MDG progress. They altered the current Joint monitoring programme (JMP) estimation by
-including for microbial water quality and sanitary risk utilizing national water quality data. \citet{onda2012global}
-used principal components analysis to analyze the national environment and development indicators and then created
-models. Overall, our analysis can highlight the potential fatigue illness caused by people transporting the water if it
-is done manually. How the modes of transportation can affect one's health.
+\citet{onda2012global} discusses the global access to safe water by accounting for water quality and the resulting
+impact on MDG progress. They altered the current Joint Monitoring Programme (JMP) estimation by accounting for microbial
+water quality and sanitary risk utilizing national water quality data. 
+%\citet{onda2012global} used principal components
+%analysis to analyze the national environment and development indicators and then created models. 
+Overall, their analysis highlights the potential fatigue illness caused by people transporting the water if it is done manually. How the
+modes of transportation can affect one's health.
 
@@
 
-\citet{onda2014country} discusses their new development of typology of country clusters pertain to water and sanitation
-sector based on congruence through multiple related indicators. \citet{onda2014country} used a hierarchical clustering
-method and a gap statistic analysis to cluster 156 countries which has 6.75 billion people into a relational
-clusters. \citet{onda2014country} suggests previous geography or income-based country clustering should be improved by
-using water and sanitation related indicators. Our research can improvise upon their finding and focus on identifying
-which water accessibility variables may improve water accessibility for people who are seeking clean drinking water.
+\citet{onda2014country} proposed a country typology based on water and sanitation indicators.
+They applied hierarchical clustering to 156 countries, accounting for 6.75 billion people.
+This work improves on previous geography or income-based country clustering.
+%Their research can improvise upon their finding and focus on identifying which water accessibility variables may improve water accessibility for people who are seeking clean drinking water.
 
@@
 affordable and safe followed by the Sustainable Development Goal six goals. The WHO/UNICEF Joint Monitoring Program for
-Water Supply and Sanitation (JMP) are monitoring progress on drinking water and sanitation since 1990 globally. The
+Water Supply and Sanitation (JMP) have been globallhy monitoring progress on drinking water and sanitation since 1990. The
 article outlines different plans with consistent traits to determine what is deemed to be safe and affordable drinking
@@
 access and quality in urban slums may influence on health risks. \citet{price2019difference} believes that monitoring
-temporal dynamics should be considered over three interlinked time scales:
-\begin{center}
-\begin{tabular}{ c c c }
- short-term & medium term & long-term \\
-
-\end{tabular}
-\end{center}
+temporal dynamics should be considered over three interlinked time scales: short, medium and long term.
 

### 2021-01-25 docs/manuscript/manuscript.tex

@@
 improved drinking water sources has increased globally from 76 percent to 89 percent, and utilization of improved
-sanitation has increased from 45 percent to 64 percent \citep{fuller2016tracking}. According to the MDG categorization,
+sanitation has increased from 45 percent to 64 percent \citep{fuller2016tracking}. According to the Millennium Development Goals (MDGs),
 they consider improved drinking water sources as the following: public tap, borehole, protected spring, rainwater
 collection, bottled water source, and piped water \citep{bartram2014globala}. However, these variables do not truly
-explain if the households have speedy access to retrieve the water. In this paper, we investigate further on what
-variables should be included to enhance the accessibility of improved drinking water.
+explain if the households have speedy access to retrieve the water.
 
@@
 Next, we present the data and methods...''}}
-
+In this paper, we investigate further on what
+variables should be included to enhance the accessibility of improved drinking water.
 In the next section we discuss relevant prior research on water accessibility quantification. We are interested in what type of modes of
-transportation are utilized to provide clean and drinking water to the community. Also, we desire to learn more about
+transportation are utilized to provide clean and drinking water to the community. Also, our objective is to investigate
 the time it takes to transport water to the community in a global modeling context. So that we can develop and quantify

### 2021-01-25 docs/manuscript/manuscript.tex
Updates from Overleaf

@@
 they consider improved drinking water sources as the following: public tap, borehole, protected spring, rainwater
-collection, bottled water source, and piped water \citep{bartram2014globala}. However, these variables do not truly
+collection, bottled water source, and piped water \citep{bartram_global_2014}. However, these variables do not truly
 explain if the households have speedy access to retrieve the water.
+Additionally, \citet{graham_analysis_2016} investigated that many people in sub-Saharan Africa (SSA) leave their homes to retrieve water, potentially placing them at risk for varieties of negative health outcomes.
+There are estimation of 3.36 million children and 13.54 million adult females were responsible for collection in households with fetching times greater than 30 minutes \citep{graham_analysis_2016}.
+
+%  Conclusion In the 24 SSA countries studied, an estimated 3.36 million children and 13.54 million adult females were responsible for water collection in households with collection times greater than 30 minutes.
+We suggest that accessibility to water, water collection by children, and gender ratios for water collection, especially when collection times are great, should be considered as key indicators for measuring progress in the water, sanitation and hygiene sector.
 

### 2021-01-25 docs/manuscript/manuscript.tex
Edit of introduction.

@@
 sanitation has increased from 45 percent to 64 percent \citep{fuller2016tracking}. According to the Millennium Development Goals (MDGs),
-they consider improved drinking water sources as the following: public tap, borehole, protected spring, rainwater
+they consider improved drinking water sources like the following: public tap, borehole, protected spring, rainwater
 collection, bottled water source, and piped water \citep{bartram_global_2014}. However, these variables do not truly
 explain if the households have speedy access to retrieve the water.
-Additionally, \citet{graham_analysis_2016} investigated that many people in sub-Saharan Africa (SSA) leave their homes to retrieve water, potentially placing them at risk for varieties of negative health outcomes.
-There are estimation of 3.36 million children and 13.54 million adult females were responsible for collection in households with fetching times greater than 30 minutes \citep{graham_analysis_2016}.
+Additionally, \citet{graham_analysis_2016} investigated that many people in sub-Saharan Africa (SSA) leave their homes to retrieve water, potentially placing them at risk for a variety of negative health outcomes.
+There are an estimation of 3.36 million children and 13.54 million adult females were responsible for collection in households with fetching times greater than 30 minutes \citep{graham_analysis_2016}.
+Therefore, there needs to be a way to alleviate people's burdens to retrieve water and potentially investigate water accessibility improving variables.
 

### 2021-01-25 docs/manuscript/manuscript.tex
Updates from Overleaf

@@
 Next, we present the data and methods...''}}
+\section{Objectives}
 In this paper, we investigate further on what
 variables should be included to enhance the accessibility of improved drinking water.
-In the next section we discuss relevant prior research on water accessibility quantification. We are interested in what type of modes of
+In the next section, we discuss relevant prior research on water accessibility quantification. We are interested in what type of modes of
 transportation are utilized to provide clean and drinking water to the community. Also, our objective is to investigate
@@
 and effective modes of transportation of accessible water. Therefore, the new models will be able to influence and
-encourage policymakers in specific countries to improve water accessibility for all. Next we present the data and methods of approaching the quantification of the water accessibility.
+encourage policymakers in specific countries to improve water accessibility for all. Next, we present the data and methods of approaching the quantification of the water accessibility.
 
-\section{Relevant work}
+\section{Contribution}
 
-\citet{jagals2006does} highlights in a case study in rural area in South Africa discovered providing small
-communities, using untreated river water as their water source, with good quality water with a piped distribution system
-and increase of accessibility by installing communal taps did not fall within parameters of safe water. Their research
+\citet{jagals2006does} highlights in a case study in a rural area in South Africa discovered providing small communities.
+The rural area used untreated river water as their water source.
+The Researchers installed good quality water with a piped distribution system and increase accessibility by installing communal taps.
+However, this implementation did not fall within the parameters of safe water.
+This didn't increase household quantities, which remained around 18 L per person per day.
+Their research
 gives a potential understanding even making the drinking water accessible might not improve the quality of the water

### 2021-01-25 docs/manuscript/manuscript.tex
Revision.

@@
 \citet{onda2012global} discusses the global access to safe water by accounting for water quality and the resulting
-impact on MDG progress. They altered the current Joint Monitoring Programme (JMP) estimation by accounting for microbial
+impact on (MDGs) progress. They altered the current Joint Monitoring Programme (JMP) estimation by accounting for microbial
 water quality and sanitary risk utilizing national water quality data. 
-%\citet{onda2012global} used principal components
-%analysis to analyze the national environment and development indicators and then created models. 
-Overall, their analysis highlights the potential fatigue illness caused by people transporting the water if it is done manually. How the
+Overall, their analysis highlights the potential fatigue can have causes that aren't due to underlying disease.
+The main example would be heavy exertion of carrying water.
+Improvement of modes of transportation can enhance water accessibility which can improve one's health overall.
 modes of transportation can affect one's health.
@@
 \citet{sima2013water} highlights the relation of the topic where it discovered that over 76 percent of the city's water
-is utilized by less than ten percent of household piped water.  The study is focused in Kisumu Kenya and Sima
+is utilized by less than ten percent of household piped water.  The study is focused on Kisumu Kenya and Sima
 interviewed 260 informal water business operators. This article revealed a new insight where the majority of the
@@
 most energy consumption and was high cost: water delivery trucks have the highest per cubic meter energy demand (35
-$MJ/m^{3}$)\citep{sima2013water}. The article utilized SPSS Version 19.0 to analyze the data.
+$MJ/m^{3}$)\citep{sima2013water}.
 
-\citet{ho2014challenge} highlights a head-to-head comparison of such indicators with other possible distance and time
+\citet{ho2014challenge} investigates a head-to-head comparison of such indicators with other possible distance and time
 metrics among rural 1,103 households in Nampula province, Mozambique. The researchers found out that Euclidean line
 distance to be a good representative for route distance ($R^{2} = 0.98$), while self-reported travel time is a poor
-representative ($R^{2} = 0.12$). One of key insights was that, 15-minute decrease in one-way walk time to water source
+representative ($R^{2} = 0.12$). One of the key insights was that a 15-minute decrease in one-way walk time to a water source
 is related to a 41\% reduction in diarrhea, improved children's nutrition, and an 11\% reduction in under-five child
-mortality rate \citep{ho2014challenge}. This research can be a helpful tool to understand how long people are traveling
-in a rural area in Africa where many of our data focus. Also, understanding how people are traveling to receive the
-water can better understand the condition of the rural people and implement better policies that help them to thrive.
+mortality rate \citep{ho2014challenge}.
 
-\citet{bartram2014globala} highlights a background information about the Millennium Development Goals. How it has
+\citet{bartram2014globala} highlights background information about the Millennium Development Goals. How it has
 established global targets for drinking water and sanitation access (Bartram et al. 2014). The article indirectly gives
@@
 modes of transportation of clean drinking water would affect the progress of improving drinking water and sanitation in
-one's community. Also, understanding how long it takes to attain the water can improve the quality of the services
-provided for the people.
+one's community.
 
-\citet{onda2014country} proposed a country typology based on water and sanitation indicators.
-They applied hierarchical clustering to 156 countries, accounting for 6.75 billion people.
+\citet{onda2014country} proposed a country typology based on nine water and sanitation indicators.
+They applied hierarchical clustering to 156 countries, accounting for 6.75 billion people (93\%).
 This work improves on previous geography or income-based country clustering.
-%Their research can improvise upon their finding and focus on identifying which water accessibility variables may improve water accessibility for people who are seeking clean drinking water.
 

### 2021-01-25 docs/manuscript/manuscript.tex
Revision of the literature review sections.

@@
 affordable and safe followed by the Sustainable Development Goal six goals. The WHO/UNICEF Joint Monitoring Program for
-Water Supply and Sanitation (JMP) have been globallhy monitoring progress on drinking water and sanitation since 1990. The
+Water Supply and Sanitation (JMP) have been globally monitoring progress on drinking water and sanitation since 1990. The
 article outlines different plans with consistent traits to determine what is deemed to be safe and affordable drinking
@@
 Also queuing time measurement can be hard to be measured because of their ambiguity.
-Her Research gives greater insight into how do we define water accessibility and how increase of water accessibility can impact people's lives beneficially.
+Their research gives greater insight into how do we define water accessibility and how increase of water accessibility can impact people's lives beneficially.
 
@@
 %Other examples: \citep{amit2019measuring,ho2014challenge}
-\subsection*{Gaps and areas for improvement}
 We are interested in analyzing the impact of transportation modes (usage/ownership) on water accessibility.
@@
 
-\section*{Data}
+\section{Data}
 \hl{In this section, describe the data sources and the variables, along with data preparation procedures, variable
@@
   etc). Also discuss the number of countries, etc. I've started a table as an example. You can add other columns if needed.}
-\subsection*{Data Preparation}
+\subsection{Data Preparation}
 Data structure this study were collected from the Demographic Health Survey (DHS) program.
@@
 
-\section*{Methods}
+\section{Methods}
 
-\subsection*{Preamble}
+\subsection{Preamble}
 We analyzed data from the Demographic Health Survey (DHS) and the Malaria Indicator Survey (MIS) (1990-2020) to highlight water accessibility predictors in 78 countries throughout the world.
 
-\subsection*{Cluster analysis}
+\subsection{Cluster analysis}
 We elected 11 variables that describe water accessibility of 78 countries around the world \citep{price2019difference}.
@@
 
-\subsection*{Classification model to explain water accessibility predictors}
+\subsection{Classification model to explain water accessibility predictors}
 \hl{This is the second phase of the work, which you should hopefully start in January}
 
-\section*{Results and Discussion}
+\section{Results and Discussion}
 
-\subsection*{Model results}
+\subsection{Model results}
 First, we developed a correlation chart to quickly analyze if there is any great correlation between all the 28 numerical variables that may pertain to water accessibility.
@@
 
-\subsection*{Water accessibility typologies}
+\subsection{Water accessibility typologies}
 \hl{Show the typology/cluster results and discuss.}
@@
 
-\subsection*{Model results}
+\subsection{Model results}
 
@@
 
-\section*{Conclusion}
+\section{Conclusion}
 
-\section*{Limitation}
+\section{Limitation}
 %he problem of missing data is relatively common in almost all research and can have a significant effect on the conclusions that can be drawn from the data.

### 2021-01-25 docs/manuscript/manuscript.tex

@@
 
-\citet{WHOdrinkingwater} article gives an ambitious plan to reach universal drinking water by the year of 2030 that is
+\citet{WHOdrinkingwater} article gives an ambitious plan to reach universal drinking water by the year 2030 that is
 affordable and safe followed by the Sustainable Development Goal six goals. The WHO/UNICEF Joint Monitoring Program for
-Water Supply and Sanitation (JMP) have been globally monitoring progress on drinking water and sanitation since 1990. The
-article outlines different plans with consistent traits to determine what is deemed to be safe and affordable drinking
-water and good sanitation. Our study can highlight which modes of transportation of safe and affordable drinking water
-can develop why one's community may excel in the improvement of safe and affordable drinking water and
-sanitation. Drinking, sanitation, and hand washing ladder explains the qualitative determination can indirectly help
-what is a good source of transporting safe and drinking water.
+Water Supply and Sanitation (JMP) have been globally monitoring progress on drinking water and sanitation since 1990.
+\citet{WHOdrinkingwater} outlines different plans with consistent traits to determine what is deemed to be safe and affordable drinking
+water and good sanitation.
 
-\citet{fuller2016tracking} highlights the global accessibility to safe drinking water and sanitation and concludes that
-it has been improving well during the Millennium Development Goal Period. This topic will benefit our research to
-understand what countries are doing for the mode of transportation of drinking water and sanitation which may have
-contributed to the improvements. \citet{fuller2016tracking} chose some countries and studied the overall countries
+\citet{fuller2016tracking} analyzes the global accessibility to safe drinking water and sanitation and concludes that
+it has been improving well during the Millennium Development Goal Period. \citet{fuller2016tracking} chose some countries and studied the overall countries
 progress which in his methodology revealed sigmoidal trends. Our study can reveal if certain methods of transportation
-of drinking water and sanitation may have influenced the effectiveness of the improved drinking water met by MDG
+of drinking water and sanitation may have influenced the effectiveness of the improved drinking water met by MDGs
 standards.
@@
     \bf Variable & \bf Description & \bf Definition \\\midrule
-    \texttt{ptap} & Households using a public tap/standpipe
- & Percentage of households whose main source of drinking water is a public tap/standpipe.
+    \texttt{ptap} &  Public tap/standpipe
+ & 
  \\\midrule
- \texttt{bore} & Households using a tube well/borehole & Percentage of households whose main source of drinking water is a tube-well/borehole.
+ \texttt{bore} & Tube well/borehole & 
 \\\midrule
-\texttt{pspr} & Households using a protected spring
-& Percentage of households whose main source of drinking water is a protected spring.
+\texttt{pspr} & Protected spring
+& 
 \\\midrule
-\texttt{uspr} & Households using an unprotected spring
-& Percentage of households whose main source of drinking water is an unprotected spring.
+\texttt{uspr} & Unprotected spring
+& 
 \\\midrule
-\texttt{pwel} & Households using a protected well
-& Percentage of households whose main source of drinking water is a protected well.
+\texttt{pwel} & Protected well
+& 
 \\\midrule
-\texttt{uwel} & Households using an unprotected well water
-&Percentage of households whose main source of drinking water is an unprotected well.
+\texttt{uwel} & Unprotected well water
+&
 \\\midrule
-\texttt{surw} & Households using surface water
-& Percentage of households whose main source of drinking water is surface water.
+\texttt{surw} & Surface water
+& 
 \\\midrule
-\texttt{tles} & Households with water 30 minutes or less away round trip
-&Percentage of households with water 30 minutes or less away round trip.
+\texttt{tles} & Water 30 minutes or less away round trip
+&
 \\\midrule 
-\texttt{tmor} & Households with water more than 30 minutes away round trip & Percentage of households with water more than 30 minutes away round trip.
+\texttt{tmor} & Water more than 30 minutes away round trip & 
 \\\midrule
-\texttt{rain} & Households using rainwater
-& Percentage of households whose main source of drinking water is rainwater.
+\texttt{rain} & Rainwater
+& 
 \\\midrule
-\texttt{phom} & Households using water piped into dwelling
-&Percentage of households whose main source of drinking water is water piped into the dwelling.
+\texttt{phom} & Piped into dwelling
+&
 \\\midrule
@@
   \end{tabular}
-  \caption{Summary of water accessibility variables (national averages at household level)}
+  \caption{Summary of water accessibility variables (National Household usage Averages measured as percentage for Main Drinking Source)}
   \label{tab:summary-wa}
@@
     \texttt{imws} & Improved water source
- & Drinking water is an improved source.
+ & 
  \\\midrule
   \texttt{pipy} & Water piped into yard/plot
-& Drinking water is water piped into the yard or plot.
+&
 \\\midrule
- \texttt{truc} & Tanker truck & Drinking water is tanker truck.
+ \texttt{truc} & Tanker truck & 
 \\\midrule
- \texttt{ctan} & Cart with tank & Drinking water is cart with tank.
+ \texttt{ctan} & Cart with tank & 
 \\\midrule
@@
 \\\midrule
- \texttt{uiws} & Unimproved water source & Drinking water is an unimproved source.
+ \texttt{uiws} & Unimproved water source & 
 \\\midrule
- \texttt{othw} & Other water source & Drinking water is other.
+ \texttt{othw} & Other water source & 
 \\\midrule
@@
 \\\midrule
- \texttt{liws} & Households with limited water service
- & Percentage of households with limited water service, defined as an improved water source with round-trip collection time greater than 30 minutes.
+ \texttt{liws} & Limited water service
+ & Limited water service, defined as an improved water source with round-trip collection time greater than 30 minutes.
 \\\midrule
- \texttt{prem} & Households with water on the premises
- & Percentage of households with water on the premises.
+ \texttt{prem} & Water on the premises
+ & 
 \\\midrule
- \texttt{bicy} & Households possessing a bicycle
- & Percentage of households possessing a bicycle.
+ \texttt{bicy} & Bicycle
+ &
 \\\midrule
- \texttt{acar} & Households possessing an animal drawn cart
- & Percentage of households possessing an animal drawn cart.
+ \texttt{acar} & Animal drawn cart
+ & 
 \\\midrule
- \texttt{mcyc} & Households possessing a motorcycle
- & Percentage of households possessing a motorcycle.
+ \texttt{mcyc} & Motorcycle
+ & 
 \\\midrule
- \texttt{pcar} & Households possessing a private car
- & Percentage of households possessing a private car.
+ \texttt{pcar} & Private car
+ & 
 \\\midrule
- \texttt{boat} & Households possessing a boat with a motor
- & Percentage of households possessing a boat with a motor.
+ \texttt{boat} & Boat with a motor
+ & 
 \\\midrule
@@
   \end{tabular}
-  \caption{Summary of explanatory variables (National Household usage Averages measured as percentage for Main Source)}
+  \caption{Summary of explanatory variables (National Household usage Averages measured as percentage for Main Drinking Source)}
   \label{tab:summary-var}
@@
 \subsection{Water accessibility typologies}
+
 \hl{Show the typology/cluster results and discuss.}
 NBClust statistic analysis developed and determined four distinct groups.
-First cluster group composed of countries near the equator or in the Sub-Sahara Regions.
+The first cluster group is composed of countries near the equator or in the Sub-Sahara Regions.
 The analysis of the second group represents countries considered developed or recently industrialized.
-Third cluster group represents most of the countries in the central Africa and may have similar factors as the first cluster group.
-The fourth cluster group represents most of countries in the middle east or predominately non-African countries.
-There is a great internally heterogeneous clusters for the 76 distinct countries.
+The third cluster group represents most of the countries in central Africa and may have similar factors as the first cluster group.
+The fourth cluster group represents most of the countries in the middle east or predominately non-African countries.
+There are great internally heterogeneous clusters for the 76 distinct countries.
 The cluster analysis' dendrogram is shown in \autoref{fig:dendrogram}, and the corresponding NBClust curve yielding the optimal four clusters is shown in \autoref{fig:optimal-clusters}, \autoref{fig:optimal}.
-\autoref{}
+\autoref{tab:water-cluster} shows the resulting four water accessibility country clusters.
+The cluster's water accessibility variables are summarized in \autoref{tab:summary-wa}.
+Through the optimal analysis of the clusters, which indicated the superior grouping of four cuts along with the relevant indicators.
 
-%   Table 3 shows the resulting five water and sanitation country clusters.
-Fig.
-3 shows the cluster country membership on a world map.
-The cluster indicator variable means are summarized in Table 4.
-The Dunn and Davies-BouldinIndices for our clusters, as well as the World Bank, WHO, and MDGgroupings are shown in Table 5.
-Our clusters have both the largest Dunn index and smallest Davies-Bouldin index, indicating that they represent a superior grouping along with the chosen relevant indicators.
+%    Fig. 3 shows the cluster country membership on a world map. 
+\begin{table}[h!]
+  \centering
+  \begin{tabular}{p{1in} p{3in} }\toprule
+    \bf Clusters & \bf Country Members \\\midrule
+    \texttt{A (n=22)} & Afghanistan, Bangladesh, Benin, Burkina Faso, Cameroon, Chad, Ivory Coast, Ghana, Guinea, India, Liberia, Malawi, Mali, Myanmar, Nepal, Niger, Nigeria, Pakistan, Togo, Uganda, Zambia, Zimbabwe  \\\midrule
+    \texttt{B (n=14)} & Albania, Armenia, Brazil, Colombia, Egypt, Honduras, Jordan, Kazakhstan, Morocco, Peru, South Africa, Turkey, Ukraine, Uzbekistan  \\\midrule
+    \texttt{C (n=20)} & Angola, Burundi, Central African Republic, Congo, Congo Democratic Republic, Eritrea, Eswatini, Ethiopia, Gambia, Haiti, Kenya, Lesotho, Madagascar, Mauritania, Mozambique, Papua New Guinea, Rwanda, Sao Tome and Principe, Sierra Leone, Tanzania  \\\midrule
+     \texttt{D (n=22)} & Azerbaijan, Bolivia, Cambodia, Comoros, Dominican Republic, Gabon, Guatemala, Guyana, Indonesia, Kyrgyz Republic, Maldives, Moldova, Namibia, Nicaragua, Paraguay, Philippines, Senegal, Tajikistan, Timor-Leste, Turkmenistan, Vietnam, Yemen  \\\bottomrule
+  \end{tabular}
+    \caption{Water accessibility country clusters.)}
+  \label{tab:water-cluster}
+\end{table}
 
@@
 \subsection{Model results}
+Cluster A consists of 22 countries with great cluster typology factor for
+
+%Cluster 1 consists of 33 developed and recently industrializedcountries from North America, Europe, and the wealthier Asiancountries.
+Cluster 1 countries are characterized by the highest GDPsper capita and GE scores, relatively low levels of inequality as mea-sured by the Gini index, the highest expected years of education,high urbanization, and near-universal water and sanitation access.Example members are Canada, Germany, and Japan.
 

### 2021-01-25 docs/manuscript/manuscript.tex
change of bib file name
Hyperlink of all the citations

@@
 \section{Introduction}
-In 2017, 29\% of the global population (2.2 billion people) did not use a safely managed drinking-water service, which is defined as water located on-premises, available, and free from contamination \citep{WHOdrinkingwater}.
+In 2017, 29\% of the global population (2.2 billion people) did not use a safely managed drinking-water service, which is defined as water located on-premises, available, and free from contamination \citep{whoDrinkingwater}.
 If you meet about 100 people, you can imagine that 29 of the people you meet do not have access to improved drinking water.
 Where this raises concerns about inequality problems related to the task and time variation
-to collect water can be inferred in the urban and rural areas \citep{cassivi2018access}. Therefore, there needs to be a
+to collect water can be inferred in the urban and rural areas \citep{cassiviAccessDrinkingWater2018}. Therefore, there needs to be a
 great motivation to bring improved drinking water sources to all people to improve their quality of life.
@@
 improved drinking water sources has increased globally from 76 percent to 89 percent, and utilization of improved
-sanitation has increased from 45 percent to 64 percent \citep{fuller2016tracking}. According to the Millennium Development Goals (MDGs),
+sanitation has increased from 45 percent to 64 percent \citep{fullerTrackingProgressGlobal2016}. According to the Millennium Development Goals (MDGs),
 they consider improved drinking water sources like the following: public tap, borehole, protected spring, rainwater
-collection, bottled water source, and piped water \citep{bartram_global_2014}. However, these variables do not truly
+collection, bottled water source, and piped water \citep{bartramGlobalMonitoringWater2014}. However, these variables do not truly
 explain if the households have speedy access to retrieve the water.
-Additionally, \citet{graham_analysis_2016} investigated that many people in sub-Saharan Africa (SSA) leave their homes to retrieve water, potentially placing them at risk for a variety of negative health outcomes.
-There are an estimation of 3.36 million children and 13.54 million adult females were responsible for collection in households with fetching times greater than 30 minutes \citep{graham_analysis_2016}.
+Additionally, \citet{grahamAnalysisWaterCollection2016} investigated that many people in sub-Saharan Africa (SSA) leave their homes to retrieve water, potentially placing them at risk for a variety of negative health outcomes.
+There are an estimation of 3.36 million children and 13.54 million adult females were responsible for collection in households with fetching times greater than 30 minutes \citep{grahamAnalysisWaterCollection2016}.
 Therefore, there needs to be a way to alleviate people's burdens to retrieve water and potentially investigate water accessibility improving variables.
@@
 
-\citet{jagals2006does} highlights in a case study in a rural area in South Africa discovered providing small communities.
+\citet{jagalsDoesImprovedAccess2006} highlights in a case study in a rural area in South Africa discovered providing small communities.
 The rural area used untreated river water as their water source.
@@
 
-\citet{onda2012global} discusses the global access to safe water by accounting for water quality and the resulting
+\citet{ondaGlobalAccessSafe2012} discusses the global access to safe water by accounting for water quality and the resulting
 impact on (MDGs) progress. They altered the current Joint Monitoring Programme (JMP) estimation by accounting for microbial
@@
 
-\citet{sima2013water} highlights the relation of the topic where it discovered that over 76 percent of the city's water
+\citet{simaWaterFlowsEnergy2013} highlights the relation of the topic where it discovered that over 76 percent of the city's water
 is utilized by less than ten percent of household piped water.  The study is focused on Kisumu Kenya and Sima
@@
 population in Kisumu relied on purchased water from kiosks (1.5 million $m^{3}$ per day) and used hand-drawn water-carts
-(0.75 million $m^{3}$ per day)\citep{sima2013water}. Also, we discovered that the water trucking industry utilized the
+(0.75 million $m^{3}$ per day)\citep{simaWaterFlowsEnergy2013}. Also, we discovered that the water trucking industry utilized the
 most energy consumption and was high cost: water delivery trucks have the highest per cubic meter energy demand (35
-$MJ/m^{3}$)\citep{sima2013water}.
+$MJ/m^{3}$)\citep{simaWaterFlowsEnergy2013}.
 
-\citet{ho2014challenge} investigates a head-to-head comparison of such indicators with other possible distance and time
+\citet{hoChallengeGlobalWater2014} investigates a head-to-head comparison of such indicators with other possible distance and time
 metrics among rural 1,103 households in Nampula province, Mozambique. The researchers found out that Euclidean line
@@
 is related to a 41\% reduction in diarrhea, improved children's nutrition, and an 11\% reduction in under-five child
-mortality rate \citep{ho2014challenge}.
+mortality rate \citep{hoChallengeGlobalWater2014}.
 
-\citet{bartram2014globala} highlights background information about the Millennium Development Goals. How it has
+\citet{bartramGlobalMonitoringWater2014} highlights background information about the Millennium Development Goals. How it has
 established global targets for drinking water and sanitation access (Bartram et al. 2014). The article indirectly gives
 much information about progress towards focused targets, directed by international monitoring, has reduced the global
-disease burden and improved the quality of life \citep{bartram2014globala}. It utilizes a variety of sources such as
+disease burden and improved the quality of life \citep{bartramGlobalMonitoringWater2014}. It utilizes a variety of sources such as
 JMP, DHS, MICS surveys to analyze the data of many different locations in the world, therefore, it is a broad scope of
@@
 
-\citet{onda2014country} proposed a country typology based on nine water and sanitation indicators.
+\citet{ondaCountryClusteringApplied2014} proposed a country typology based on nine water and sanitation indicators.
 They applied hierarchical clustering to 156 countries, accounting for 6.75 billion people (93\%).
@@
 
-\citet{WHOdrinkingwater} article gives an ambitious plan to reach universal drinking water by the year 2030 that is
+\citet{whoDrinkingwater} article gives an ambitious plan to reach universal drinking water by the year 2030 that is
 affordable and safe followed by the Sustainable Development Goal six goals. The WHO/UNICEF Joint Monitoring Program for
 Water Supply and Sanitation (JMP) have been globally monitoring progress on drinking water and sanitation since 1990.
-\citet{WHOdrinkingwater} outlines different plans with consistent traits to determine what is deemed to be safe and affordable drinking
+\citet{whoDrinkingwater} outlines different plans with consistent traits to determine what is deemed to be safe and affordable drinking
 water and good sanitation.
 
-\citet{fuller2016tracking} analyzes the global accessibility to safe drinking water and sanitation and concludes that
-it has been improving well during the Millennium Development Goal Period. \citet{fuller2016tracking} chose some countries and studied the overall countries
+\citet{fullerTrackingProgressGlobal2016} analyzes the global accessibility to safe drinking water and sanitation and concludes that
+it has been improving well during the Millennium Development Goal Period. \citet{fullerTrackingProgressGlobal2016} chose some countries and studied the overall countries
 progress which in his methodology revealed sigmoidal trends. Our study can reveal if certain methods of transportation
@@
 
-\citet{jia2016highresolution} highlights that there is inequality within safe sanitation. There is a great difference in
+\citet{jiaHighResolutionSpatialDistribution2016} highlights that there is inequality within safe sanitation. There is a great difference in
 wealth levels in many low-income countries. Therefore, this will hinder the improvement of MDG
-progress. \citet{jia2016highresolution}discusses future interventions to be engaged with different wealth
+progress. \citet{jiaHighResolutionSpatialDistribution2016}discusses future interventions to be engaged with different wealth
 categories. Our study can assist in understanding improved sanitation increase by making inferences about how they
@@
 
-\citet{cassivi2018access} highlights about how there should be more information to the effect of including a 30-minute
+\citet{cassiviAccessDrinkingWater2018} highlights about how there should be more information to the effect of including a 30-minute
 collection time to monitor access to drinking water. The lack of access to on-premises water sources can cause the use
 of greater distant alternative sources. Therefore causing the quantity of retrieved water to reduce as the water plateau
-phenomenon.\citet{cassivi2018access} hypothesizes that the households who must travel further are using unimproved
+phenomenon.\citet{cassiviAccessDrinkingWater2018} hypothesizes that the households who must travel further are using unimproved
 sources. For our research, we can discover what modes of transportation people are traveling to retrieve the water and
@@
 
-\citet{tortajada2018achieving} highlights that sharing the immense improvement at the 2015 convention negatively
+\citet{tortajadaAchievingUniversalAccess2018} highlights that sharing the immense improvement at the 2015 convention negatively
 affected many people because they actually have not acquired clean drinking water. While 2.6 billion people gained
@@
 
-\citet{amit2019measuring} highlights their data analysis similar to what we will be doing and found out that for
+\citet{amitMeasuringAffordabilityAccess2019} highlights their data analysis similar to what we will be doing and found out that for
 non-piped households, collection costs are 22\% of the coping costs, while collection costs for piped households are
@@
 
-\citet{price2019difference} highlights the temporal dynamics of drinking water access and quality in urban slums which
-is home to about 1 billion people globally. \citet{price2019difference} discusses the temporal changes in drinking water
-access and quality in urban slums may influence on health risks. \citet{price2019difference} believes that monitoring
-temporal dynamics should be considered over three interlinked time scales: short, medium and long term.
-
-\citet{price2019difference} recommends future research to explore the short-term water access and quality monitoring.
+\citet{priceDifferenceDayCan2019} highlights the temporal dynamics of drinking water access and quality in urban slums which
+is home to about 1 billion people globally. \citet{priceDifferenceDayCan2019} discusses the temporal changes in drinking water
+access and quality in urban slums may influence on health risks. \citet{priceDifferenceDayCan2019} believes that monitoring
+temporal dynamics should be considered over three interlinked time scales: short, medium and long term.\citet{priceDifferenceDayCan2019} recommends future research to explore the short-term water access and quality monitoring.
 Also we should learn to recognize the role of socio-cultural factors that may impact the temporal dynamics of safe water access.
@@
 
-\citet{rawas2020comparing} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru.
+\citet{rawasComparingUtilityreportedHours2020} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru.
 This relates and informs one mode of transportation of clean and accessible water.
@@
 
-\citet{deshpande2020mapping} discusses their findings of improved overall water availability and piped water accessibility for the globe (40\% to 50.3\%).
-\citet{deshpande2020mapping} also found that it was the lowest increase in sub-Saharan Africa, where the accessibility was mostly concentrated in urban regions.
-\citet{deshpande2020mapping} focuses mostly on understanding trends in diarrhea burden, where the needs are greatest for the improvement of access to safe drinking water and sanitation facilities.
+\citet{deshpandeMappingGeographicalInequalities2020} discusses their findings of improved overall water availability and piped water accessibility for the globe (40\% to 50.3\%).
+\citet{deshpandeMappingGeographicalInequalities2020} also found that it was the lowest increase in sub-Saharan Africa, where the accessibility was mostly concentrated in urban regions.
+\citet{deshpandeMappingGeographicalInequalities2020} focuses mostly on understanding trends in diarrhea burden, where the needs are greatest for the improvement of access to safe drinking water and sanitation facilities.
 Our study can benefit this research by implementing the best and effective modes of transportation of clean water to improve safe drinking water and sanitation.
@@
 
-\citet{goal} explains specific targets that have goals and indicators to track global water accessibility.
+\citet{GoalCleanWater} explains specific targets that have goals and indicators to track global water accessibility.
 The article quantifies and provides Graphs: safe drinking water, safe sanitation, and hygiene, safe sanitation and hygiene, ambient water quality, water use efficiency, levels of freshwater stress, integrated water management, trans-boundary water cooperation, protect and restore water-related ecosystems, water and sanitation support, local participation in sanitation management.
@@
 
-\citet{cassivi2021evaluating} focuses a case study in Southern Malawi, where data collected in March and April of 2019 at the conclusion of the rainy season.
+\citet{cassiviEvaluatingSelfreportedMeasures2021} focuses a case study in Southern Malawi, where data collected in March and April of 2019 at the conclusion of the rainy season.
 The sample was consisted of 375 households without access to water on their premises.
@@
 \texttt{phom} & Piped into dwelling
-&
-\\\midrule
 
-& \\\bottomrule
+ \\\bottomrule
   \end{tabular}
@@
  & The Gini coefficient indicates the level of concentration of wealth in the country.
-\\\midrule
- & \\\bottomrule
+  \\\bottomrule
   \end{tabular}

### 2021-01-25 docs/manuscript/manuscript.tex

@@
 
-There is a problem of people not available to retrieve water quickly and there is a collection burden which is necessary to increase the water quality to improve many people's health and quality of life \citep{cassivi2018access}.
+There is a problem of people not available to retrieve water quickly and there is a collection burden which is necessary to increase the water quality to improve many people's health and quality of life \citep{cassiviAccessDrinkingWater2018}.
 These initial findings can allow us to research into what are the major factors of water accessibility for the 76 countries by developing clustering and determining the water accessibility variable predictors.
@@
   \centering
-  \includegraphics[width=.7\textwidth]{typology}
+  \includegraphics[width=1\textwidth]{typology}
   \caption{Cluster Typology Average of Water Accessibility Variables }
@@
 \begin{figure}[h!]
-  \centering
-  \includegraphics[width=.7\textwidth]{dendrogram}
+\raggedright
+  \includegraphics[width=1.1\textwidth]{dendrogram}
   \caption{Cluster Dendrogram with Four Cuts}
@@
 \end{figure}
+
 \begin{figure}[h!]
   \centering
-  \includegraphics[width=.7\textwidth]{spider-plot}
+  \includegraphics[width= 1.25\textwidth]{spider-plot}
   \caption{Spider Plot of Water Accessibility Variables by Typology}
@@
 \subsection{Model results}
-Cluster A consists of 22 countries with great cluster typology factor for
+Cluster A consists of 22 countries with great cluster typology factor for Water 30 minutes or less away round trip (tles)
 
-%Cluster 1 consists of 33 developed and recently industrializedcountries from North America, Europe, and the wealthier Asiancountries.
-Cluster 1 countries are characterized by the highest GDPsper capita and GE scores, relatively low levels of inequality as mea-sured by the Gini index, the highest expected years of education,high urbanization, and near-universal water and sanitation access.Example members are Canada, Germany, and Japan.
+%
 

### 2021-01-25 docs/manuscript/manuscript.tex
model-results, first paragraph.

@@
 \raggedright
-  \includegraphics[width=1.1\textwidth]{dendrogram}
+  \includegraphics[width=1.25\textwidth]{dendrogram}
   \caption{Cluster Dendrogram with Four Cuts}
@@
   \centering
-  \includegraphics[width= 1.25\textwidth]{spider-plot}
+  \includegraphics[width= 1\textwidth]{spider-plot}
   \caption{Spider Plot of Water Accessibility Variables by Typology}
@@
 \subsection{Model results}
-Cluster A consists of 22 countries with great cluster typology factor for Water 30 minutes or less away round trip (tles)
+Cluster A consists of 22 countries with major cluster typology factor for Water 30 minutes or less away round trip (tles) and Tube well/borehole (bore) as shown in \autoref{fig:spider} and \autoref{fig:typology}.
+Cluster A has countries from Africa and Asia.
+Cluster A countries are characterized by the high measure of traveling 30 minutes or less to retrieve water and are known to use boreholes as their resource of water.
+Example country members are Afghanistan, Bangladesh, and Benin.
 
-%
+%Cluster 1 consists of 33 developed and recently industrialized countries from North America, Europe, and the wealthier Asian countries.
+Cluster 1 countries are characterized by the highest GDPs per capita and GE scores, relatively low levels of inequality as mea-sured by the Gini index, the highest expected years of education,high urbanization, and near-universal water and sanitation access.Example members are Canada, Germany, and Japan.
 

### 2021-01-26 docs/manuscript/manuscript.tex
writing results for cluster B
Updates from Overleaf
Description of Cluster C.

--- 2021-01-25 docs/manuscript/manuscript.tex
@@
 The cluster's water accessibility variables are summarized in \autoref{tab:summary-wa}.
+\autoref{fig:cluster-map} shows the cluster country membership of the water accessibility on a world map.
 Through the optimal analysis of the clusters, which indicated the superior grouping of four cuts along with the relevant indicators.
 
-%    Fig. 3 shows the cluster country membership on a world map. 
 \begin{table}[h!]
@@
 \end{table}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=1\textwidth]{cluster-map}
+  \caption{Water Accessibility Cluster Map}
+  \label{fig:cluster-map}
+\end{figure}
 
@@
 \subsection{Model results}
-Cluster A consists of 22 countries with great cluster typology factor for Water 30 minutes or less away round trip (tles)
+Cluster A consists of 22 countries with major cluster typology factor for Water 30 minutes or less away round trip and Tube well/borehole as shown in \autoref{fig:spider} and \autoref{fig:typology}.
+Cluster A has countries from Africa and Asia.
+Cluster A countries are characterized by the high measure of traveling 30 minutes or less to retrieve water and are known to use boreholes as their resource of water.
+Example country members are Afghanistan, Bangladesh, and Benin.
+
+Cluster B consists of 14 further advanced and industrialized countries from South America, Europe, Asia, and wealthier African nations.
+Cluster B countries are categorized by their cluster typology factor for Piped into dwelling as shown in \autoref{fig:spider} and \autoref{fig:typology}.
+Example country members are Albania, Armenia, and Brazil.
+
+Cluster C consists of 20 countries mostly in central Africa, some in west Africa and one country in the pacific islands.
+Cluster C countries are grouped by their cluster typology factor for largest for Water 30 minutes or less away round trip, second largest typology factor of Public tap/standpipe, and third largest typology factor of Water more than 30 minutes away round trip.
+Example country members are Angola, Burundi, Central African Republic.
 
@@
 
-  % \section*{Discussion}
+\section{Discussion}
 
@@
 \section{Limitation}
-%he problem of missing data is relatively common in almost all research and can have a significant effect on the conclusions that can be drawn from the data.
+%The problem of missing data is relatively common in almost all research and can have a significant effect on the conclusions that can be drawn from the data.
 

### 2021-01-26 docs/manuscript/manuscript.tex
Removal of parentheses for variable labels.

@@
 First, we developed a correlation chart to quickly analyze if there is any great correlation between all the 28 numerical variables that may pertain to water accessibility.
-We recognized the strong and weak relationship by the size and color of the correlation plot is shown in \autoref{fig:cor-basic}, which indicates households with limited water service (liws) has the strongest and positive relationship with Households possessing a bicycle (bicy).
-Meanwhile, households using an improved water source (imws) has the strongest and negative relationship with households using an unimproved water source (uiws).
+We recognized the strong and weak relationship by the size and color of the correlation plot is shown in \autoref{fig:cor-basic}, which indicates households with limited water service has the strongest and positive relationship with Households possessing a bicycle.
+Meanwhile, households using an improved water source has the strongest and negative relationship with households using an unimproved water source.
 This initial finding results indicated that the initial variables are logical and can give us insight into further understanding and the relationship between these variables.
@@
 As we continued to study further upon the correlations of the variables, we used more advanced correlation graphs to determined their p-value, and to see a relationship between different categorical variables is shown in \autoref{fig:cor-plot}, which again indicates a greater detailed comparison between the variables.
-Households with an improved water source on the premises (iwsp) have a strong relationship with households with basic water service (basw) of .84 and a p-value of less than 0.001.
-Also, we learned and discovered that Households using an unimproved water source (uiws) have a positive relationship with Households using unprotected well water (uwel), Households using an unprotected spring (uspr), and Households using surface water (surw) with a p-value of less than 0.001 as well.
+Households with an improved water source on the premises have a strong relationship with households with basic water service of .84 and a p-value of less than 0.001.
+Also, we learned and discovered that Households using an unimproved water source have a positive relationship with Households using unprotected well water, Households using an unprotected spring, and Households using surface water with a p-value of less than 0.001 as well.
 These relationships justify the necessity to research further if the particular water accessibility variables can increase the accessibility of particular water sources to reach universal access to water by 2030, as proposed in the SDGs.
@@
 Therefore, we investigated by developing more correlation plots relating transportation variables with water accessibility variables which are shown in \autoref{fig:transporation-time}, \autoref{fig:transporation-water-one}, \autoref{fig:transporation-water-two}.
-There is a significant positive relationship between households possessing a private car (pcar) and households with water on the premises (watp) as shown in \autoref{fig:transporation-time}.
-There is also is a significant negative relationship between households possessing a private car (pcar) with households with water 30 minutes or less away round trip (tles) and households with water more than 30 minutes away round trip (tmor) with a p-value of less than 0.001.
+There is a significant positive relationship between households possessing a private car and households with water on the premises as shown in \autoref{fig:transporation-time}.
+There is also is a significant negative relationship between households possessing a private car with households with water 30 minutes or less away round trip and households with water more than 30 minutes away round trip with a p-value of less than 0.001.
 This information is essential to investigate further why this is the global trend.
@@
 \end{figure}
-\autoref{fig:transporation-water-one} shows that Households possessing a motorcycle (mcyc) has a positive correlation with Households using a tube well/borehole (bore) and Households using rainwater (rain) with a p-value of less than 0.01.
+\autoref{fig:transporation-water-one} shows that Households possessing a motorcycle has a positive correlation with Households using a tube well/borehole and Households using rainwater with a p-value of less than 0.01.
 Similar to the previous analysis, there is a relationship between a mode of transportation to retrieve potential water sources.
-\autoref{fig:transporation-water-two} also shows that households possessing an animal-drawn cart (acar) have a positive and strong relationship with households using unprotected well water (uwel) with a p-value of less than 0.001.
+\autoref{fig:transporation-water-two} also shows that households possessing an animal-drawn cart have a positive and strong relationship with households using unprotected well water with a p-value of less than 0.001.
 The results demonstrate that different modes of transportation factor into the accessibility of retrieving water sources on a global scale, therefore, there should be consideration of these variables to increase access to improved water sources.

### 2021-01-26 docs/manuscript/manuscript.tex
Contribution first paragraph.

@@
 \section{Contribution}
-
+Lack of water accessibility is a major crisis for many vulnerable people around the world.
+We have determined to seek literature reviews from the past 15 years to have the most up to date information because information on water accessibility is a topic that is constantly changing throughout the world due to different international, government, and even local policies.
+The literature review intends to investigate different journal articles on water accessibility in a global scale.
+Therefore we can conduct a research to fill many major gaps  to analyze the water accessibility predictor variables which could increase and enhance water accessibility in global scale.
+We will always gain new insight and wisdom on information about water accessibility through the global scale as we investigate and determine literature reviews chronologically.
 \citet{jagalsDoesImprovedAccess2006} highlights in a case study in a rural area in South Africa discovered providing small communities.
@@
 \raggedright
-  \includegraphics[width=1.25\textwidth]{dendrogram}
+  \includegraphics[width=1.15\textwidth]{dendrogram}
   \caption{Cluster Dendrogram with Four Cuts}
@@
 Cluster C consists of 20 countries mostly in central Africa, some in west Africa and one country in the pacific islands.
-Cluster C countries are grouped by their cluster typology factor for largest for Water 30 minutes or less away round trip, second largest typology factor of Public tap/standpipe, and third largest typology factor of Water more than 30 minutes away round trip.
+Cluster C countries are grouped by their cluster typology factor for largest for Water 30 minutes or less away round trip, second largest typology factor of Public tap/standpipe, and third largest typology factor of Water more than 30 minutes away round trip as shown in \autoref{fig:spider} and \autoref{fig:typology}.
 Example country members are Angola, Burundi, Central African Republic.
 
-%
+Cluster D consists of 22 geographically dispersed countries.
+Cluster D countries are categorized by their cluster typology factor for rainwater compared to the other cluster typology.
+Additionally, the highest factor was Piped into dwelling, but is still lower than the Cluster B as shown in \autoref{fig:spider} and \autoref{fig:typology}.
+Example country members are Azerbaijan, Bolivia, and Cambodia.
 

### 2021-01-26 docs/manuscript/manuscript.tex

@@
 We will always gain new insight and wisdom on information about water accessibility through the global scale as we investigate and determine literature reviews chronologically.
+\subsection{2006-2016}
+Improving water accessibility have been a priority by many organizations around the world, realizing it is the basic human rights to have clean and accessible drinkable water in past 15 years.
+During 2006 to 2016, overall water accessibility has been improving by utilizing the indicators of the past \citep{fullerTrackingProgressGlobal2016}.
+However, some researchers raises concerns on lack of clarity with the global water accessibility monitoring \citep{ondaGlobalAccessSafe2012, jiaHighResolutionSpatialDistribution2016}.
+However, in an era of rapidly changing advanced technologies, international, and government policies, some of the older research might not be adequate for up to date information for  example of The eight Millennium Development Goals (MDGs) upgrading into The Sustainable Development Goals (SDGs).
+In light of the newest SDGs of providing clean and accessible drinking water for all by 2030, researchers have devoted their study to understand water accessibility in global and specific country levels.
+
+%  Several empirical studies have focused on Facebook usage in adolescent girls (Tiggermann & Slater, 2013; Meier & Gray, 2014) and in young adult women (Smith, Hames, & Joiner, 2013; Fardouly et al., 2015; Cohen, Newton-John & Slater, 2017), while a systematic review by Holland and Timmerman (2016) confirmed a relationship between social networking and body image for both women and men.
+Across these studies, there is consistent evidence that body image issues are influenced not by social media usage in general, but by engagement with the visual and interactive aspects of these platforms.
+
 \citet{jagalsDoesImprovedAccess2006} highlights in a case study in a rural area in South Africa discovered providing small communities.
@@
 
-\citet{simaWaterFlowsEnergy2013} highlights the relation of the topic where it discovered that over 76 percent of the city's water
+\citet{simaWaterFlowsEnergy2013} highlights topic where it discovered that over 76 percent of the Kenya's city's water
 is utilized by less than ten percent of household piped water.  The study is focused on Kisumu Kenya and Sima
@@
 \citet{bartramGlobalMonitoringWater2014} highlights background information about the Millennium Development Goals. How it has
-established global targets for drinking water and sanitation access (Bartram et al. 2014). The article indirectly gives
+established global targets for drinking water and sanitation access. The article directly gives
 much information about progress towards focused targets, directed by international monitoring, has reduced the global
@@
 
-\citet{ondaCountryClusteringApplied2014} proposed a country typology based on nine water and sanitation indicators.
+\citet{ondaCountryClusteringApplied2014} proposed a country typology based on nine water and sanitation indicators in a global scale.
 They applied hierarchical clustering to 156 countries, accounting for 6.75 billion people (93\%).
@@
 
+\subsection{2017-Present}
 \citet{cassiviAccessDrinkingWater2018} highlights about how there should be more information to the effect of including a 30-minute

### 2021-01-27 docs/manuscript/manuscript.tex
The first body section of contribution.

@@
 
-%  Conclusion In the 24 SSA countries studied, an estimated 3.36 million children and 13.54 million adult females were responsible for water collection in households with collection times greater than 30 minutes.
-We suggest that accessibility to water, water collection by children, and gender ratios for water collection, especially when collection times are great, should be considered as key indicators for measuring progress in the water, sanitation and hygiene sector.
-
 % {\reversemarginpar\marginnote{\scriptsize\rd The introduction should end with a paragraph summarizing the contents and layout of the remainder of the manuscript.
@@
 In light of the newest SDGs of providing clean and accessible drinking water for all by 2030, researchers have devoted their study to understand water accessibility in global and specific country levels.
-
-%  Several empirical studies have focused on Facebook usage in adolescent girls (Tiggermann & Slater, 2013; Meier & Gray, 2014) and in young adult women (Smith, Hames, & Joiner, 2013; Fardouly et al., 2015; Cohen, Newton-John & Slater, 2017), while a systematic review by Holland and Timmerman (2016) confirmed a relationship between social networking and body image for both women and men.
-Across these studies, there is consistent evidence that body image issues are influenced not by social media usage in general, but by engagement with the visual and interactive aspects of these platforms.
+Several research and studies have focused global water accessibility ,and its progress \citep{sorensonSafeAccessSafe2011,bainAccountingWaterQuality2012,wssccWaterSanitationHygiene2014,  ondaGlobalAccessSafe2012, bartramGlobalMonitoringWater2014, ondaCountryClusteringApplied2014, whoDrinkingwater,fullerTrackingProgressGlobal2016, weiyuyuGlobalPerspectiveDrinkingwater2016, grahamAnalysisWaterCollection2016,overboOnplotDrinkingWater2016}.
+While there has been studies focusing on local community in different parts of the country learning about their water accessibility and sanitation initiation \citep{jagalsDoesImprovedAccess2006, geereDomesticWaterCarrying2010, simaWaterFlowsEnergy2013, hoChallengeGlobalWater2014, sakisakaIntroductionCommunityWater2015, jiaHighResolutionSpatialDistribution2016}.
+Where majority raised concerns about lack of monitoring sustainable water accessibility measurements for many people, and even few argues for greater focus in many people's access to drinking water sources.
+Nonetheless, there is a lack of robust study on modes of transportation related to enhance water accessibility for all people by 2030.
 

### 2021-01-27 docs/manuscript/manuscript.tex
Simplified contribution section for chronology.

@@
 Nonetheless, there is a lack of robust study on modes of transportation related to enhance water accessibility for all people by 2030.
-
-\citet{jagalsDoesImprovedAccess2006} highlights in a case study in a rural area in South Africa discovered providing small communities.
-The rural area used untreated river water as their water source.
-The Researchers installed good quality water with a piped distribution system and increase accessibility by installing communal taps.
-However, this implementation did not fall within the parameters of safe water.
-This didn't increase household quantities, which remained around 18 L per person per day.
-Their research
-gives a potential understanding even making the drinking water accessible might not improve the quality of the water
-itself.
-
-\citet{ondaGlobalAccessSafe2012} discusses the global access to safe water by accounting for water quality and the resulting
-impact on (MDGs) progress. They altered the current Joint Monitoring Programme (JMP) estimation by accounting for microbial
-water quality and sanitary risk utilizing national water quality data. 
-Overall, their analysis highlights the potential fatigue can have causes that aren't due to underlying disease.
-The main example would be heavy exertion of carrying water.
-Improvement of modes of transportation can enhance water accessibility which can improve one's health overall.
-modes of transportation can affect one's health.
-
-\citet{simaWaterFlowsEnergy2013} highlights topic where it discovered that over 76 percent of the Kenya's city's water
-is utilized by less than ten percent of household piped water.  The study is focused on Kisumu Kenya and Sima
-interviewed 260 informal water business operators. This article revealed a new insight where the majority of the
-population in Kisumu relied on purchased water from kiosks (1.5 million $m^{3}$ per day) and used hand-drawn water-carts
-(0.75 million $m^{3}$ per day)\citep{simaWaterFlowsEnergy2013}. Also, we discovered that the water trucking industry utilized the
-most energy consumption and was high cost: water delivery trucks have the highest per cubic meter energy demand (35
-$MJ/m^{3}$)\citep{simaWaterFlowsEnergy2013}.
-
-\citet{hoChallengeGlobalWater2014} investigates a head-to-head comparison of such indicators with other possible distance and time
-metrics among rural 1,103 households in Nampula province, Mozambique. The researchers found out that Euclidean line
-distance to be a good representative for route distance ($R^{2} = 0.98$), while self-reported travel time is a poor
-representative ($R^{2} = 0.12$). One of the key insights was that a 15-minute decrease in one-way walk time to a water source
-is related to a 41\% reduction in diarrhea, improved children's nutrition, and an 11\% reduction in under-five child
-mortality rate \citep{hoChallengeGlobalWater2014}.
-
-\citet{bartramGlobalMonitoringWater2014} highlights background information about the Millennium Development Goals. How it has
-established global targets for drinking water and sanitation access. The article directly gives
-much information about progress towards focused targets, directed by international monitoring, has reduced the global
-disease burden and improved the quality of life \citep{bartramGlobalMonitoringWater2014}. It utilizes a variety of sources such as
-JMP, DHS, MICS surveys to analyze the data of many different locations in the world, therefore, it is a broad scope of
-study of the progress of accessible and drinking water and sanitation globally. Our research can study further what
-modes of transportation of clean drinking water would affect the progress of improving drinking water and sanitation in
-one's community.
-
-\citet{ondaCountryClusteringApplied2014} proposed a country typology based on nine water and sanitation indicators in a global scale.
-They applied hierarchical clustering to 156 countries, accounting for 6.75 billion people (93\%).
-This work improves on previous geography or income-based country clustering.
-
-\citet{whoDrinkingwater} article gives an ambitious plan to reach universal drinking water by the year 2030 that is
-affordable and safe followed by the Sustainable Development Goal six goals. The WHO/UNICEF Joint Monitoring Program for
-Water Supply and Sanitation (JMP) have been globally monitoring progress on drinking water and sanitation since 1990.
-\citet{whoDrinkingwater} outlines different plans with consistent traits to determine what is deemed to be safe and affordable drinking
-water and good sanitation.
-
-\citet{fullerTrackingProgressGlobal2016} analyzes the global accessibility to safe drinking water and sanitation and concludes that
-it has been improving well during the Millennium Development Goal Period. \citet{fullerTrackingProgressGlobal2016} chose some countries and studied the overall countries
-progress which in his methodology revealed sigmoidal trends. Our study can reveal if certain methods of transportation
-of drinking water and sanitation may have influenced the effectiveness of the improved drinking water met by MDGs
-standards.
-
-\citet{jiaHighResolutionSpatialDistribution2016} highlights that there is inequality within safe sanitation. There is a great difference in
-wealth levels in many low-income countries. Therefore, this will hinder the improvement of MDG
-progress. \citet{jiaHighResolutionSpatialDistribution2016}discusses future interventions to be engaged with different wealth
-categories. Our study can assist in understanding improved sanitation increase by making inferences about how they
-receive water may give the community focusing on sanitation. Overall, income inequality affects who will have access to
-proper sanitation.
-
 \subsection{2017-Present}

### 2021-01-27 docs/manuscript/manuscript.tex
Reorganization of contribution section.

@@
 We have determined to seek literature reviews from the past 15 years to have the most up to date information because information on water accessibility is a topic that is constantly changing throughout the world due to different international, government, and even local policies.
-The literature review intends to investigate different journal articles on water accessibility in a global scale.
-Therefore we can conduct a research to fill many major gaps  to analyze the water accessibility predictor variables which could increase and enhance water accessibility in global scale.
+The literature review intends to investigate different journal articles on water accessibility on a global scale.
+Therefore we can conduct research to fill many major gaps to analyze the water accessibility predictor variables which could increase and enhance water accessibility on a global scale.
 We will always gain new insight and wisdom on information about water accessibility through the global scale as we investigate and determine literature reviews chronologically.
 \subsection{2006-2016}
-Improving water accessibility have been a priority by many organizations around the world, realizing it is the basic human rights to have clean and accessible drinkable water in past 15 years.
-During 2006 to 2016, overall water accessibility has been improving by utilizing the indicators of the past \citep{fullerTrackingProgressGlobal2016}.
-However, some researchers raises concerns on lack of clarity with the global water accessibility monitoring \citep{ondaGlobalAccessSafe2012, jiaHighResolutionSpatialDistribution2016}.
+Improving water accessibility has been a priority by many organizations around the world, realizing it is the basic human right to have clean and accessible drinkable water in the past 15 years.
+From 2006 to 2016, overall water accessibility has been improving by utilizing the indicators of the past \citep{fullerTrackingProgressGlobal2016}.
+However, some researchers raise concerns on lack of clarity with the global water accessibility monitoring \citep{ondaGlobalAccessSafe2012, jiaHighResolutionSpatialDistribution2016}.
 However, in an era of rapidly changing advanced technologies, international, and government policies, some of the older research might not be adequate for up to date information for example of The eight Millennium Development Goals (MDGs) upgrading into The Sustainable Development Goals (SDGs).
-In light of the newest SDGs of providing clean and accessible drinking water for all by 2030, researchers have devoted their study to understand water accessibility in global and specific country levels.
+In light of the newest SDGs of providing clean and accessible drinking water for all by 2030, researchers have devoted their studies to understand water accessibility at global and specific country levels.
 Several research and studies have focused global water accessibility ,and its progress \citep{sorensonSafeAccessSafe2011,bainAccountingWaterQuality2012,wssccWaterSanitationHygiene2014,  ondaGlobalAccessSafe2012, bartramGlobalMonitoringWater2014, ondaCountryClusteringApplied2014, whoDrinkingwater,fullerTrackingProgressGlobal2016, weiyuyuGlobalPerspectiveDrinkingwater2016, grahamAnalysisWaterCollection2016,overboOnplotDrinkingWater2016}.
-While there has been studies focusing on local community in different parts of the country learning about their water accessibility and sanitation initiation \citep{jagalsDoesImprovedAccess2006, geereDomesticWaterCarrying2010, simaWaterFlowsEnergy2013, hoChallengeGlobalWater2014, sakisakaIntroductionCommunityWater2015, jiaHighResolutionSpatialDistribution2016}.
-Where majority raised concerns about lack of monitoring sustainable water accessibility measurements for many people, and even few argues for greater focus in many people's access to drinking water sources.
-Nonetheless, there is a lack of robust study on modes of transportation related to enhance water accessibility for all people by 2030.
+While there have been studies focusing on the local community in different parts of the country learning about their water accessibility and sanitation initiation \citep{jagalsDoesImprovedAccess2006, geereDomesticWaterCarrying2010, simaWaterFlowsEnergy2013, hoChallengeGlobalWater2014, sakisakaIntroductionCommunityWater2015, jiaHighResolutionSpatialDistribution2016}.
+Where majority raised concerns about the lack of monitoring sustainable water accessibility measurements for many people and even a few argue for greater focus on many people's access to drinking water sources.
+Nonetheless, there is a lack of robust study on modes of transportation-related to enhance water accessibility for all people by 2030.
 \subsection{2017-Present}
-\citet{cassiviAccessDrinkingWater2018} highlights about how there should be more information to the effect of including a 30-minute
-collection time to monitor access to drinking water. The lack of access to on-premises water sources can cause the use
-of greater distant alternative sources. Therefore causing the quantity of retrieved water to reduce as the water plateau
-phenomenon.\citet{cassiviAccessDrinkingWater2018} hypothesizes that the households who must travel further are using unimproved
-sources. For our research, we can discover what modes of transportation people are traveling to retrieve the water and
-in what quantity depending on their means of transportation. Therefore, learning how to decrease the travel time of
-people to retrieve the water because according to the article, it can be a great burden and does not increase the
-quality of one's life to retrieve water daily and far away from their homes.
+However, in the most recent years, there is a lack of information that needs to focus on people to receive accessible drinking water.
+The relation between water accessibility and the total time to retrieve drinking water is a concern and (SDGs) should emphasize this study \citep{cassiviAccessDrinkingWater2018}.
+In the most recent years, there needs a greater understanding of the water accessibility indicators to enhance drinking water accessibility on a global scale for many vulnerable people.
+While research on water accessibility on a global scale has diminished due to an increase of water for many people \citep{fullerTrackingProgressGlobal2016}, but \citet{deshpandeMappingGeographicalInequalities2020} raises concerns that poorer community around the world are still not receiving clean and accessible drinking water.
+Several research and studies focus on a global scale to understand water accessibility and sanitation \citep{cassiviAccessDrinkingWater2018, tortajadaAchievingUniversalAccess2018, priceDifferenceDayCan2019,deshpandeMappingGeographicalInequalities2020}.
+Some research highlighted the coping cost in small community in India \citep{amitMeasuringAffordabilityAccess2019} and researched on inequality with water accessibility and sanitation \citet{amitMeasuringAffordabilityAccess2019,priceDifferenceDayCan2019, rawasComparingUtilityreportedHours2020}.
+As many research journal articles focused on different aspects of water accessibility and sanitation in particular countries also brought greater insight and necessary to understand water accessibility variables \citep{rawasComparingUtilityreportedHours2020, cassiviEvaluatingSelfreportedMeasures2021}.
+\citet{GoalCleanWater} emphasized providing safe and accessible drinking water for all on a global scale, which defines that we need a greater understanding of water accessibility and what water accessibility variables be investigated further.
+Therefore, we will be conducting robust research to understand water accessibility variables and predictors to improve water accessibility to bring clean and safe drinking water for all by 2030 as (SDGs) visioned.
 
-\citet{tortajadaAchievingUniversalAccess2018} highlights that sharing the immense improvement at the 2015 convention negatively
-affected many people because they actually have not acquired clean drinking water. While 2.6 billion people gained
-access to drinking water, they did not have constructive feedback on the quality of the water. Also, our research can
-fill in the blank on how did these people actually retrieve the water? Our research can measure the performance of how
-people retrieve water which may not be in their premises.
-
-\citet{amitMeasuringAffordabilityAccess2019} highlights their data analysis similar to what we will be doing and found out that for
-non-piped households, collection costs are 22\% of the coping costs, while collection costs for piped households are
-less than 2\% of the coping costs. The researchers also realize that there is economic inequality because the majority
-of the piped households are in households that are considered wealthy households and vice versa. Our research can
-pinpoint how households are traveling to collect water to improve potential policy-making for water accessibility.
-
-\citet{priceDifferenceDayCan2019} highlights the temporal dynamics of drinking water access and quality in urban slums which
-is home to about 1 billion people globally. \citet{priceDifferenceDayCan2019} discusses the temporal changes in drinking water
-access and quality in urban slums may influence on health risks. \citet{priceDifferenceDayCan2019} believes that monitoring
-temporal dynamics should be considered over three interlinked time scales: short, medium and long term.\citet{priceDifferenceDayCan2019} recommends future research to explore the short-term water access and quality monitoring.
-Also we should learn to recognize the role of socio-cultural factors that may impact the temporal dynamics of safe water access.
-Our project can pinpoint additional water accessibility variables that can be used to focus to recognize the time variations as well.
-
-\citet{rawasComparingUtilityreportedHours2020} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru.
-This relates and informs one mode of transportation of clean and accessible water.
-Our research can make inferences about if the customers receive clean water elsewhere, such as water tankers, kiosks, vendors, etc.
-Overall, the article found that the percent of households with (IWS) was a few percentage points greater than of the reported utilities.
-
-\citet{deshpandeMappingGeographicalInequalities2020} discusses their findings of improved overall water availability and piped water accessibility for the globe (40\% to 50.3\%).
-\citet{deshpandeMappingGeographicalInequalities2020} also found that it was the lowest increase in sub-Saharan Africa, where the accessibility was mostly concentrated in urban regions.
-\citet{deshpandeMappingGeographicalInequalities2020} focuses mostly on understanding trends in diarrhea burden, where the needs are greatest for the improvement of access to safe drinking water and sanitation facilities.
-Our study can benefit this research by implementing the best and effective modes of transportation of clean water to improve safe drinking water and sanitation.
-Even with the water pipe access increasing, they can't conclude if the people had clean water as they needed right away.
-Overall, the access to safe drinking water and sanitation improved globally between 2000 and 2017, but dis-proportionality undermines reaching the Sustainable Development Goals.
-
-\citet{GoalCleanWater} explains specific targets that have goals and indicators to track global water accessibility.
-The article quantifies and provides Graphs: safe drinking water, safe sanitation, and hygiene, safe sanitation and hygiene, ambient water quality, water use efficiency, levels of freshwater stress, integrated water management, trans-boundary water cooperation, protect and restore water-related ecosystems, water and sanitation support, local participation in sanitation management.
-They have many ambitious plans to achieve by 2020 and 2030, and our research can find a correlation to why achieving this may be possible.
-
-\citet{cassiviEvaluatingSelfreportedMeasures2021} focuses a case study in Southern Malawi, where data collected in March and April of 2019 at the conclusion of the rainy season.
-The sample was consisted of 375 households without access to water on their premises.
-The article highlighted using Euclidean Distance measurement is more reliable than the self-reported time.
-Also queuing time measurement can be hard to be measured because of their ambiguity.
-Their research gives greater insight into how do we define water accessibility and how increase of water accessibility can impact people's lives beneficially.
+The journal articles were missing what modes of transportation may enhance water accessibility.
+Therefore we will be analyzing the impact of transportation modes (usage/ownership) on water accessibility.
+We will therefore research by statistical analysis in a global scale and pilot in one of the countries to determine if the water accessibility variables predictors can improve the water accessibility for its country.
 
@@
 % citet (for citations where the year only is in parentheses)
-%%\textit{Discuss the who/what/when/where? (You already have this... just need to reorganize or improve structure)}
-%Another example is \citet{fuller2016tracking}.
-%\citet{rawas2020comparing} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru.
-%\citet{cassivi2018access} discuss time access...
-%\hl{Perhaps insert a table summarizing the various dimensions of water accessibility that have been studied}
-%Other examples: \citep{amit2019measuring,ho2014challenge}
-We are interested in analyzing the impact of transportation modes (usage/ownership) on water accessibility.
-We also desire to see an implementation of this statistical research piloted in one of the countries to see if the water accessibility variables predictors can improve the water accessibility for its country.
+%\textit{Discuss the who/what/when/where? (You already have this... just need to reorganize or improve structure)}
 

### 2021-01-27 docs/manuscript/manuscript.tex
Added body paragraph for introduction for more information on the access we are studying.

@@
 %%\item What has been the rate of improvement over the years?
+A simple necessity of access is the existence of an actual water source.
+People can obtain water from nature in a variety of ways, but Sustainable Development Goals (SDGs) has determined the "A safely managed drinking water service" which are proper for human consumption \citep{wssccWaterSanitationHygiene2014,whoDrinkingwater,GoalCleanWater}.
+A safely managed drinking water service is defined as one located on-premises, available when needed and free from contamination \citep{GoalCleanWater}.
+This definition requires a basic drinking water source that is positioned on-premises, available when needed, and free of fecal and priority chemical contamination.
+While basic drinking water sources are described as piped water, boreholes or
+tube wells, protected dug wells, protected springs and rainwater provided collection time is no more than 30 minutes for
+a round trip including queuing \citep{wssccWaterSanitationHygiene2014}.
+Also (SDGs) establishes "access" as satisfactory water to meet household needs is surely available close to home \citep{GoalCleanWater}.
 
@@
 \subsection{2006-2016}
-Improving water accessibility has been a priority by many organizations around the world, realizing it is the basic human right to have clean and accessible drinkable water in the past 15 years.
+Improving water accessibility has been a priority by many organizations, realizing it is the basic human right to have clean and accessible drinkable water in the past 15 years.
 From 2006 to 2016, overall water accessibility has been improving by utilizing the indicators of the past \citep{fullerTrackingProgressGlobal2016}.
 However, some researchers raise concerns on lack of clarity with the global water accessibility monitoring \citep{ondaGlobalAccessSafe2012, jiaHighResolutionSpatialDistribution2016}.
-However, in an era of rapidly changing advanced technologies, international, and government policies, some of the older research might not be adequate for up to date information for example of The eight Millennium Development Goals (MDGs) upgrading into The Sustainable Development Goals (SDGs).
-In light of the newest SDGs of providing clean and accessible drinking water for all by 2030, researchers have devoted their studies to understand water accessibility at global and specific country levels.
+However, in an era of rapidly changing technologies, international and government policies, the older researches are not adequate.
+For an example of The eight Millennium Development Goals (MDGs) upgrading into The Sustainable Development Goals (SDGs).
+In light of the newest (SDGs) providing clean and accessible drinking water for all by 2030, researchers have devoted their studies to understand water accessibility at global and specific country levels.
 Several research and studies have focused global water accessibility ,and its progress \citep{sorensonSafeAccessSafe2011,bainAccountingWaterQuality2012,wssccWaterSanitationHygiene2014,  ondaGlobalAccessSafe2012, bartramGlobalMonitoringWater2014, ondaCountryClusteringApplied2014, whoDrinkingwater,fullerTrackingProgressGlobal2016, weiyuyuGlobalPerspectiveDrinkingwater2016, grahamAnalysisWaterCollection2016,overboOnplotDrinkingWater2016}.
 While there have been studies focusing on the local community in different parts of the country learning about their water accessibility and sanitation initiation \citep{jagalsDoesImprovedAccess2006, geereDomesticWaterCarrying2010, simaWaterFlowsEnergy2013, hoChallengeGlobalWater2014, sakisakaIntroductionCommunityWater2015, jiaHighResolutionSpatialDistribution2016}.
-Where majority raised concerns about the lack of monitoring sustainable water accessibility measurements for many people and even a few argue for greater focus on many people's access to drinking water sources.
-Nonetheless, there is a lack of robust study on modes of transportation-related to enhance water accessibility for all people by 2030.
+Where majority raised concerns about the lack of monitoring sustainable water accessibility measurements for many people and some argue for a focus on many people's access to drinking water sources.
+Nonetheless, there is a lack of robust study on transportation to enhance water accessibility for all people by 2030.
 \subsection{2017-Present}
@@
 We will therefore research by statistical analysis in a global scale and pilot in one of the countries to determine if the water accessibility variables predictors can improve the water accessibility for its country.
-
 % citep (for citations fully enclosed in parentheses)
@@
 %\textit{Discuss the who/what/when/where? (You already have this... just need to reorganize or improve structure)}
-
 \eject

### 2021-01-28 docs/manuscript/manuscript.tex
Edit of bib file names

@@
 \section{Introduction}
-In 2017, 29\% of the global population (2.2 billion people) did not use a safely managed drinking-water service, which is defined as water located on-premises, available, and free from contamination \citep{whoDrinkingwater}.
+In 2017, 29\% of the global population (2.2 billion people) did not use a safely managed drinking-water service, which is defined as water located on-premises, available, and free from contamination \citep{who2015drinkingwater}.
 If you meet about 100 people, you can imagine that 29 of the people you meet do not have access to improved drinking water.
 Where this raises concerns about inequality problems related to the task and time variation
-to collect water can be inferred in the urban and rural areas \citep{cassiviAccessDrinkingWater2018}. Therefore, there needs to be a
+to collect water can be inferred in the urban and rural areas \citep{cassivi2018access}. Therefore, there needs to be a
 great motivation to bring improved drinking water sources to all people to improve their quality of life.
@@
 A simple necessity of access is the existence of an actual water source.
-People can obtain water from nature in a variety of ways, but Sustainable Development Goals (SDGs) has determined the "A safely managed drinking water service" which are proper for human consumption \citep{wssccWaterSanitationHygiene2014,whoDrinkingwater,GoalCleanWater}.
-A safely managed drinking water service is defined as one located on-premises, available when needed and free from contamination \citep{GoalCleanWater}.
+People can obtain water from nature in a variety of ways, but Sustainable Development Goals (SDGs) has determined the "A safely managed drinking water service" which are proper for human consumption \citep{wssccWaterSanitationHygiene2014,who2015drinkingwater,goal}.
+A safely managed drinking water service is defined as one located on-premises, available when needed and free from contamination \citep{goal}.
 This definition requires a basic drinking water source that is positioned on-premises, available when needed, and free of fecal and priority chemical contamination.
@@
 tube wells, protected dug wells, protected springs and rainwater provided collection time is no more than 30 minutes for
-a round trip including queuing \citep{wssccWaterSanitationHygiene2014}.
-Also (SDGs) establishes "access" as satisfactory water to meet household needs is surely available close to home \citep{GoalCleanWater}.
+a round trip including queuing \citep{wssccWaterSanitationHygiene2014}. Also (SDGs) establishes "access" as satisfactory water to meet household needs is surely available close to home \citep{goal}.
 
@@
 improved drinking water sources has increased globally from 76 percent to 89 percent, and utilization of improved
-sanitation has increased from 45 percent to 64 percent \citep{fullerTrackingProgressGlobal2016}. According to the Millennium Development Goals (MDGs),
+sanitation has increased from 45 percent to 64 percent \citep{fuller2016tracking}. According to the Millennium Development Goals (MDGs),
 they consider improved drinking water sources like the following: public tap, borehole, protected spring, rainwater
-collection, bottled water source, and piped water \citep{bartramGlobalMonitoringWater2014}. However, these variables do not truly
+collection, bottled water source, and piped water \citep{bartram2014global}. However, these variables do not truly
 explain if the households have speedy access to retrieve the water.
@@
 Improving water accessibility has been a priority by many organizations, realizing it is the basic human right to have clean and accessible drinkable water in the past 15 years.
-From 2006 to 2016, overall water accessibility has been improving by utilizing the indicators of the past \citep{fullerTrackingProgressGlobal2016}.
-However, some researchers raise concerns on lack of clarity with the global water accessibility monitoring \citep{ondaGlobalAccessSafe2012, jiaHighResolutionSpatialDistribution2016}.
+From 2006 to 2016, overall water accessibility has been improving by utilizing the indicators of the past \citep{fuller2016tracking}.
+However, some researchers raise concerns on lack of clarity with the global water accessibility monitoring \citep{onda2012global, jia2016highresolution}.
 However, in an era of rapidly changing technologies, international and government policies, the older researches are not adequate.
@@
 In light of the newest (SDGs) providing clean and accessible drinking water for all by 2030, researchers have devoted their studies to understand water accessibility at global and specific country levels.
-Several research and studies have focused global water accessibility ,and its progress \citep{sorensonSafeAccessSafe2011,bainAccountingWaterQuality2012,wssccWaterSanitationHygiene2014,  ondaGlobalAccessSafe2012, bartramGlobalMonitoringWater2014, ondaCountryClusteringApplied2014, whoDrinkingwater,fullerTrackingProgressGlobal2016, weiyuyuGlobalPerspectiveDrinkingwater2016, grahamAnalysisWaterCollection2016,overboOnplotDrinkingWater2016}.
-While there have been studies focusing on the local community in different parts of the country learning about their water accessibility and sanitation initiation \citep{jagalsDoesImprovedAccess2006, geereDomesticWaterCarrying2010, simaWaterFlowsEnergy2013, hoChallengeGlobalWater2014, sakisakaIntroductionCommunityWater2015, jiaHighResolutionSpatialDistribution2016}.
+Several research and studies have focused global water accessibility ,and its progress \citep{sorensonSafeAccessSafe2011,bainAccountingWaterQuality2012,wssccWaterSanitationHygiene2014,  onda2012global, bartram2014global, onda2014country, who2015drinkingwater,fuller2016tracking, weiyuyuGlobalPerspectiveDrinkingwater2016, grahamAnalysisWaterCollection2016,overboOnplotDrinkingWater2016}.
+While there have been studies focusing on the local community in different parts of the country learning about their water accessibility and sanitation initiation \citep{jagals2006does, geereDomesticWaterCarrying2010, sima2013water, ho2014challenge, sakisakaIntroductionCommunityWater2015, jia2016highresolution}.
 Where majority raised concerns about the lack of monitoring sustainable water accessibility measurements for many people and some argue for a focus on many people's access to drinking water sources.
@@
 However, in the most recent years, there is a lack of information that needs to focus on people to receive accessible drinking water.
-The relation between water accessibility and the total time to retrieve drinking water is a concern and (SDGs) should emphasize this study \citep{cassiviAccessDrinkingWater2018}.
+The relation between water accessibility and the total time to retrieve drinking water is a concern and (SDGs) should emphasize this study \citep{cassivi2018access}.
 In the most recent years, there needs a greater understanding of the water accessibility indicators to enhance drinking water accessibility on a global scale for many vulnerable people.
@@
 Some research highlighted the coping cost in small community in India \citep{amitMeasuringAffordabilityAccess2019} and researched on inequality with water accessibility and sanitation \citet{amitMeasuringAffordabilityAccess2019,priceDifferenceDayCan2019, rawasComparingUtilityreportedHours2020}.
-As many research journal articles focused on different aspects of water accessibility and sanitation in particular countries also brought greater insight and necessary to understand water accessibility variables \citep{rawasComparingUtilityreportedHours2020, cassiviEvaluatingSelfreportedMeasures2021}.
+As many research journal articles focused on different aspects of water accessibility and sanitation in particular countries also brought greater insight and necessary to understand water accessibility variables \citep{rawasComparingUtilityreportedHours2020, cassiviEvaluatingSelfreportedMeasures2021, martinez-santos2017determinants}.
 \citet{GoalCleanWater} emphasized providing safe and accessible drinking water for all on a global scale, which defines that we need a greater understanding of water accessibility and what water accessibility variables be investigated further.

### 2021-01-28 docs/manuscript/manuscript.tex
Edit of bib names for citation to enable hyperlinks.

@@
 In the most recent years, there needs a greater understanding of the water accessibility indicators to enhance drinking water accessibility on a global scale for many vulnerable people.
-While research on water accessibility on a global scale has diminished due to an increase of water for many people \citep{fullerTrackingProgressGlobal2016}, but \citet{deshpandeMappingGeographicalInequalities2020} raises concerns that poorer community around the world are still not receiving clean and accessible drinking water.
-Several research and studies focus on a global scale to understand water accessibility and sanitation \citep{cassiviAccessDrinkingWater2018, tortajadaAchievingUniversalAccess2018, priceDifferenceDayCan2019,deshpandeMappingGeographicalInequalities2020}.
-Some research highlighted the coping cost in small community in India \citep{amitMeasuringAffordabilityAccess2019} and researched on inequality with water accessibility and sanitation \citet{amitMeasuringAffordabilityAccess2019,priceDifferenceDayCan2019, rawasComparingUtilityreportedHours2020}.
-As many research journal articles focused on different aspects of water accessibility and sanitation in particular countries also brought greater insight and necessary to understand water accessibility variables \citep{rawasComparingUtilityreportedHours2020, cassiviEvaluatingSelfreportedMeasures2021, martinez-santos2017determinants}.
-\citet{GoalCleanWater} emphasized providing safe and accessible drinking water for all on a global scale, which defines that we need a greater understanding of water accessibility and what water accessibility variables be investigated further.
+While research on water accessibility on a global scale has diminished due to an increase of water for many people \citep{fuller2016tracking}, but \citet{deshpande2020mapping} raises concerns that poorer community around the world are still not receiving clean and accessible drinking water.
+Several research and studies focus on a global scale to understand water accessibility and sanitation \citep{cassivi2018access, tortajada2018achieving, price2019difference,deshpande2020mapping}.
+Some research highlighted the coping cost in small community in India \citep{amit2019measuring} and researched on inequality with water accessibility and sanitation \citet{amit2019measuring,price2019difference, rawas2020comparing}.
+As many research journal articles focused on different aspects of water accessibility and sanitation in particular countries also brought greater insight and necessary to understand water accessibility variables \citep{rawas2020comparing, cassivi2021evaluating, martinez-santos2017determinants}.
+\citet{goal} emphasized providing safe and accessible drinking water for all on a global scale, which defines that we need a greater understanding of water accessibility and what water accessibility variables be investigated further.
 Therefore, we will be conducting robust research to understand water accessibility variables and predictors to improve water accessibility to bring clean and safe drinking water for all by 2030 as (SDGs) visioned.
@@
 
-There is a problem of people not available to retrieve water quickly and there is a collection burden which is necessary to increase the water quality to improve many people's health and quality of life \citep{cassiviAccessDrinkingWater2018}.
+There is a problem of people not available to retrieve water quickly and there is a collection burden which is necessary to increase the water quality to improve many people's health and quality of life \citep{cassivi2018access}.
 These initial findings can allow us to research into what are the major factors of water accessibility for the 76 countries by developing clustering and determining the water accessibility variable predictors.

### 2021-01-28 docs/manuscript/manuscript.tex

@@
 In light of the newest (SDGs) providing clean and accessible drinking water for all by 2030, researchers have devoted their studies to understand water accessibility at global and specific country levels.
-Several research and studies have focused global water accessibility ,and its progress \citep{sorensonSafeAccessSafe2011,bainAccountingWaterQuality2012,wssccWaterSanitationHygiene2014,  onda2012global, bartram2014global, onda2014country, who2015drinkingwater,fuller2016tracking, weiyuyuGlobalPerspectiveDrinkingwater2016, grahamAnalysisWaterCollection2016,overboOnplotDrinkingWater2016}.
+Several research and studies have focused global water accessibility ,and its progress \citep{sorensonSafeAccessSafe2011,bainAccountingWaterQuality2012,pickering2012freshwater,wssccWaterSanitationHygiene2014,  onda2012global, bartram2014global, onda2014country, who2015drinkingwater,fuller2016tracking, weiyuyuGlobalPerspectiveDrinkingwater2016, grahamAnalysisWaterCollection2016,overboOnplotDrinkingWater2016}.
 While there have been studies focusing on the local community in different parts of the country learning about their water accessibility and sanitation initiation \citep{jagals2006does, geereDomesticWaterCarrying2010, sima2013water, ho2014challenge, sakisakaIntroductionCommunityWater2015, jia2016highresolution}.

### 2021-01-29 docs/manuscript/manuscript.tex

@@
 
-\section{Contribution}
+\section{Background}
 Lack of water accessibility is a major crisis for many vulnerable people around the world.
@@
   \begin{tabular}{p{1in} p{3in} p{2in}}\toprule
-    \bf Variable & \bf Description & \bf Definition \\\midrule
+    \bf Variable & \bf Description & \bf Notes \\\midrule
     \texttt{ptap} &  Public tap/standpipe
@@
   \begin{tabular}{p{1in} p{3in} p{2in}}\toprule
-    \bf Variable & \bf Description & \bf Definition \\\midrule
+    \bf Variable & \bf Description & \bf Notes \\\midrule
     \texttt{imws} & Improved water source
@@
   \end{tabular}
-  \caption{Summary of explanatory variables (National Household usage Averages measured as percentage for Main Drinking Source)}
+  \caption{Summary of explanatory variables (National Household usage Averages measured as percentage for Main Drinking Source) }
   \label{tab:summary-var}
@@
 
-\subsection{Preamble}
 We analyzed data from the Demographic Health Survey (DHS) and the Malaria Indicator Survey (MIS) (1990-2020) to highlight water accessibility predictors in 78 countries throughout the world.
@@
 
-\subsection{Model results}
+\subsection{Exploratory Data Analysis}
 First, we developed a correlation chart to quickly analyze if there is any great correlation between all the 28 numerical variables that may pertain to water accessibility.
@@
 
-\hl{Show the typology/cluster results and discuss.}
 NBClust statistic analysis developed and determined four distinct groups.

### 2021-02-04 docs/manuscript/manuscript.tex

@@
 \section{Data}
-\hl{In this section, describe the data sources and the variables, along with data preparation procedures, variable
-  selection criteria, etc. You should have a table summarizing the water accessibility variables and their
-  definitions. Another table can summarize the explanatory variables (e.g. transportation-related, income-related,
-  etc). Also discuss the number of countries, etc. I've started a table as an example. You can add other columns if needed.}
+%\hl{In this section, describe the data sources and the variables, along with data preparation procedures, variable selection criteria, etc.
+You should have a table summarizing the water accessibility variables and their definitions.
+Another table can summarize the explanatory variables (e.g.
+transportation-related, income-related, etc).
+Also discuss the number of countries, etc.
+I've started a table as an example.
+You can add other columns if needed.}
 \subsection{Data Preparation}
@@
  \subsection*{Summary of variables}
-\hl{Here, you show and discuss the data distributions (e.g. histograms, correlation plots, etc)}
+%\hl{Here, you show and discuss the data distributions (e.g. histograms, correlation plots, etc)}
 
@@
 \subsection{Classification model to explain water accessibility predictors}
-\hl{This is the second phase of the work, which you should hopefully start in January}
+%\hl{This is the second phase of the work, which you should hopefully start in January}
 

### 2021-02-16 docs/manuscript/manuscript.tex

@@
 
-\subsection{Model results}
+\subsection{Typology results}
 Cluster A consists of 22 countries with major cluster typology factor for Water 30 minutes or less away round trip and Tube well/borehole as shown in \autoref{fig:spider} and \autoref{fig:typology}.

### 2021-02-22 docs/manuscript/manuscript.tex
Updates from Overleaf

@@
 \section{Objectives}
-In this paper, we investigate further on what
-variables should be included to enhance the accessibility of improved drinking water.
+The objective of this work is to  investigate associations between measures of water accessibility and access to modes of transit on the national scale.
+
 In the next section, we discuss relevant prior research on water accessibility quantification. We are interested in what type of modes of
@@
 We will always gain new insight and wisdom on information about water accessibility through the global scale as we investigate and determine literature reviews chronologically.
-\subsection{2006-2016}
+\subsection{2000-2015}
 Improving water accessibility has been a priority by many organizations, realizing it is the basic human right to have clean and accessible drinkable water in the past 15 years.
@@
 Nonetheless, there is a lack of robust study on transportation to enhance water accessibility for all people by 2030.
-\subsection{2017-Present}
+\subsection{2015-2021}
 However, in the most recent years, there is a lack of information that needs to focus on people to receive accessible drinking water.
@@
 
-\section{Data}
+\section{Methods}
 %\hl{In this section, describe the data sources and the variables, along with data preparation procedures, variable selection criteria, etc.
@@
 You can add other columns if needed.}
-\subsection{Data Preparation}
-Data structure this study were collected from the Demographic Health Survey (DHS) program.
-We used the STATCompiler from the DHS Program to develop customized table based on focusing on water, transportation, and wealth indicators across 78 countries and from 1990 to 2019.
+\subsection{Data Sources}
+Data for this study were collected from the Demographic Health Survey (DHS) program, which is .
+We used the STATCompiler tool available from the DHS Program to develop customized tables on indicators related to  water, transportation, and wealth from 78 countries between 1990-2019.
 
-We prepossess the data by differentiating the different surveys collected by the DHS program between DHS and Malaria Indicators Survey (MIS) through excel.
+We preprocessed the data by differentiating the different surveys collected by the DHS program between DHS and Malaria Indicators Survey (MIS) through excel.
 After, we used R to develop and simplify the table with the most recent surveys for the most up to date analysis for the 78 countries.
-There was two sets of data of population and households.
+There were two sets of data of population and households.
 Therefore, we decided to remove the population data from the table to simplify because population and households data were identical.
 
-Next, there were many missing values from the data.
-Therefore, we had to do a simple missing value analysis.
-We determined to avoid using water, transportation, or wealth variables which had greater than 50\% missing values.
+We only preserved variables where less than 50\% of values were missing.
 Ultimately, simplified data frame was 76 countries by 28 variables.
 After, two different tables are generated where one focused on the water accessibility variables and a explanatory variables.
-We selected 11 water accessibility variables based on the sources available such as location, time, and methodology \citep{price2019difference}.
-The water accessibility variable was organized
-in the following table to allow for quick visualization for \autoref{tab:summary-wa}.
-The other variables were used as an explanatory variables for determining the water accessibility predictors in \autoref{tab:summary-var}.
+From the remaining variables, we selected 11 which related to water accessibility.
+These included: types of water source (piped into dwelling, public tap/standpipe, tube well/borehole, protected well, protected spring, rainwater, unprotected well, unprotected spring, surface water), and time to water source (less than 30 minutes, more than 30 minutes) \citep{price2019difference} \autoref{tab:summary-wa}.
+Remaining variables related to wealth and transportation were used as an explanatory variables in the model \autoref{tab:summary-var}.
 \begin{table}[h!]
@@
 
- \\\bottomrule
-  \end{tabular}
-  \caption{Summary of water accessibility variables (National Household usage Averages measured as percentage for Main Drinking Source)}
-  \label{tab:summary-wa}
-\end{table}
-
-\begin{table}[h!]
-  \centering
-  \begin{tabular}{p{1in} p{3in} p{2in}}\toprule
-    \bf Variable & \bf Description & \bf Notes \\\midrule
+ \\\midrule
 \texttt{imws} & Improved water source
@@
 \\\midrule
- \texttt{ctan} & Cart with tank & 
+ \texttt{cart} & Cart with tank & 
 \\\midrule
@@
 \\\midrule
- \texttt{prem} & Water on the premises
+ \texttt{watp} & Water on the premises
  & 
-\\\midrule
+ \\\bottomrule
+  \end{tabular}
+  \caption{Summary of water accessibility variables (National Household usage Averages measured as percentage for Main Drinking Source)}
+  \label{tab:summary-wa}
+\end{table}
+
+\begin{table}[h!]
+  \centering
+  \begin{tabular}{p{1in} p{3in} p{2in}}\toprule
+    \bf Variable & \bf Description & \bf Notes \\\midrule
  \texttt{bicy} & Bicycle
@@
 \\\midrule
- \texttt{boat} & Boat with a motor
+ \texttt{cgdp} & GDP (current US \$) 
+ & [NY.GDP.MKTP.CD]
+\\\midrule
+ \texttt{tpop} & Total Population
+ &  [SP.POP.TOTL]
+\\\midrule
+ \texttt{upop} & Urban population 
+ & (\% of total population)  [SP.URB.TOTL.IN.ZS]
+\\\midrule
+ \texttt{popd} & Population density 
+ & (people per sq. km of land area)  [EN.POP.DNST]
+\\\midrule
+ \texttt{land} & Land area
+
+ & (sq. km) [AG.LND.TOTL.K2]
+\\\midrule
+ \texttt{lita} & Boat with a motor
+ & 
+\\\midrule
+ \texttt{lity} & Boat with a motor
+ & 
+\\\midrule
+ \texttt{mori} & Boat with a motor
  & 
@@
  & The Gini coefficient indicates the level of concentration of wealth in the country.
+
   \\\bottomrule
   \end{tabular}
-  \caption{Summary of explanatory variables (National Household usage Averages measured as percentage for Main Drinking Source) }
+  \caption{Summary of explanatory variables (National Household usage Averages measured as percentage for Main Drinking Source or otherwise) }
   \label{tab:summary-var}
 \end{table}
+
  \subsection*{Summary of variables}
@@
 
-\section{Methods}
-
-We analyzed data from the Demographic Health Survey (DHS) and the Malaria Indicator Survey (MIS) (1990-2020) to highlight water accessibility predictors in 78 countries throughout the world.
-
 \subsection{Cluster analysis}
-We elected 11 variables that describe water accessibility of 78 countries around the world \citep{price2019difference}.
-Next, we decided to implement clustering which is an unsupervised machine learning method to identify and group water accessibility variables in large data set to understand and make inferences about particular water accessibility variables.
-All of the clustering analysis was assisted by using the R program.
+We selected 11 variables that describe water accessibility of 78 countries around the world \citep{price2019difference}.
+Next, we used clustering, an unsupervised machine learning method to identify and group water accessibility variables in a large data set to understand and make inferences using R.
 Clustering ultimately can continue to assist in discovering the water accessibility predictor variables.
 
-First, we used a different clustering algorithm to determine the best clustering results.
+First, we used different clustering algorithms to determine the best clustering results.
 We tested and categorized differently by using single, average, complete, ward.D, and ward.D2 methods to develop clustering dendrogram.
-Then, we analyzed by determining the best hierarchical relationship of each of the methods by observing the tree diagram itself; if it has more of even distribution of each tree are the optimal clustering methods.
-We determined that ward.d2 had the best visual and even distribution, but to check this further, we used a function in R called 'NBClust' to determine the best algorithm as well as the best cut or groupings for the clustering methods as shown in \autoref{fig:optimal-clusters}.
+Then, we analyzed by determining the best hierarchical relationship of each of the methods by observing the tree diagram itself; more even distributions of each tree are interpreted as optimal clustering methods.
+We determined that ward.d2 had most even distribution, verified through the use of 'NBClust' to determine the best algorithm as well as the best cut or groupings for the clustering methods as shown in \autoref{fig:optimal-clusters}.
 
@@
 \subsection{Exploratory Data Analysis}
-First, we developed a correlation chart to quickly analyze if there is any great correlation between all the 28 numerical variables that may pertain to water accessibility.
-We recognized the strong and weak relationship by the size and color of the correlation plot is shown in \autoref{fig:cor-basic}, which indicates households with limited water service has the strongest and positive relationship with Households possessing a bicycle.
+A correlation chart was used to assess strong correlations between the 28 numerical variables that may pertain to water accessibility \autoref{fig:cor-basic}.
+Households with limited water service has the strongest and positive relationship with households possessing a bicycle.
 Meanwhile, households using an improved water source has the strongest and negative relationship with households using an unimproved water source.
@@
 \end{figure}
+
 Additionally, we wanted to explore if the modes of transportation variables have any relationship with potential water variables.
@@
 \end{figure}
+
 \autoref{fig:transporation-water-one} shows that Households possessing a motorcycle has a positive correlation with Households using a tube well/borehole and Households using rainwater with a p-value of less than 0.01.

### 2021-02-25 docs/manuscript/manuscript.tex

@@
 \subsection{Data Sources}
-Data for this study were collected from the Demographic Health Survey (DHS) program, which is .
-We used the STATCompiler tool available from the DHS Program to develop customized tables on indicators related to  water, transportation, and wealth from 78 countries between 1990-2019.
+Data for this study were collected from the Demographic Health Survey (DHS) program, which is a program that has collected, analyzed, and distributed accurate and representative data on population, health, HIV, and nutrition through more than 400 surveys in over 90 countries.
+DHS the program uses a wide range of monitoring and impact evaluation indicators in the areas of population, health, and nutrition.
+Also, Standard DHS Surveys have high sample sizes (between 5,000 and 30,000 households) and are administered about every 5 years, to enable comparisons over time.
+DHS program also uses Malaria Indicator Surveys (MIS) is a survey where they provide data on bed-net ownership and use inhibition of malaria during pregnancy, and timely and adequate treatment of fever in young children.
+In many cases, bio-marker experiments for malaria and anemia are also included.
+We used the STAT-Compiler tool available from the DHS Program to develop customized tables on indicators related to water, transportation, and wealth from 78 countries between 1990-2019.
 
-We preprocessed the data by differentiating the different surveys collected by the DHS program between DHS and Malaria Indicators Survey (MIS) through excel.
-After, we used R to develop and simplify the table with the most recent surveys for the most up to date analysis for the 78 countries.
-There were two sets of data of population and households.
-Therefore, we decided to remove the population data from the table to simplify because population and households data were identical.
+We preprocessed the data, and we recognized that there were three different surveys collected by the DHS program via excel.
+They are DHS, MIS, and AIDS Indicator Survey (AIS).
+After, we used R to develop an algorithm to find the most recent surveys and showed results for 78 countries.
+We also simplified the variables titles into four character acronyms to recognize the different 28 variables quickly.
+We used the household data only.
+The STAT-compiler collected the aggregated data by the 78 countries
 
-We only preserved variables where less than 50\% of values were missing.
-Ultimately, simplified data frame was 76 countries by 28 variables.
-After, two different tables are generated where one focused on the water accessibility variables and a explanatory variables.
-From the remaining variables, we selected 11 which related to water accessibility.
-These included: types of water source (piped into dwelling, public tap/standpipe, tube well/borehole, protected well, protected spring, rainwater, unprotected well, unprotected spring, surface water), and time to water source (less than 30 minutes, more than 30 minutes) \citep{price2019difference} \autoref{tab:summary-wa}.
+We preserved 28 variables from 78 where less than 50\% of values were missing and the population survey data were removed because they were similar to the household surveys.
+Ultimately, the simplified data frame was 78 countries by 28 variables.
+After the simplification, two separate tables are distinguished within the 28 variables.
+One focused on the water accessibility variables and the other focused on the explanatory variables \citep{price2019difference}.
+From the remaining variables, we selected 22 which related to water accessibility.
+These included: types of water source (improved water source,
+piped into dwelling, piped into yard/plot,
+ public tap/standpipe, tube well/borehole, protected well, protected spring, rainwater, unimproved water source, unprotected well, unprotected spring, cart with small tank/drum, tanker truck, bottled water, other water source, surface water), time to water source (less than 30 minutes, more than 30 minutes), and premise to water source (water on the premises, improved water source on the premises, basic water service, limited water service) \autoref{tab:summary-wa}.
 Remaining variables related to wealth and transportation were used as an explanatory variables in the model \autoref{tab:summary-var}.
+
 \begin{table}[h!]
@@
     \bf Variable & \bf Description & \bf Notes \\\midrule
+\texttt{imws} & Improved water source
+ & 
+ \\\midrule
+ \texttt{phom} & Piped into dwelling &
+ \\\midrule
+ \texttt{pipy} & Water piped into yard/plot
+&
+\\\midrule
     \texttt{ptap} &  Public tap/standpipe
@@
 \\\midrule
+\texttt{pwel} & Protected well
+& 
+\\\midrule
 \texttt{pspr} & Protected spring
+& 
+\\\midrule
+
+\texttt{rain} & Rainwater
+& 
+\\\midrule
+ \texttt{uiws} & Unimproved water source & 
+\\\midrule
+\texttt{uwel} & Unprotected well water
 &
@@
 \\\midrule
-\texttt{pwel} & Protected well
-& 
+
+ \texttt{truc} & Tanker truck & 
 \\\midrule
-\texttt{uwel} & Unprotected well water
-&
+ \texttt{ctan} & Cart with tank & 
+\\\midrule
+ \texttt{bott} & Bottled water & Drinking water is bottled water/demi john.
+\\\midrule
+ \texttt{othw} & Other water source & 
 \\\midrule
@@
 \\\midrule
-\texttt{rain} & Rainwater
+ \texttt{watp} & Water on the premises
  & 
-\\\midrule
-\texttt{phom} & Piped into dwelling
-
- \\\midrule
-\texttt{imws} & Improved water source
- & 
- \\\midrule
-\texttt{pipy} & Water piped into yard/plot
-&
-\\\midrule
- \texttt{truc} & Tanker truck & 
-\\\midrule
- \texttt{cart} & Cart with tank & 
-\\\midrule
- \texttt{bott} & Bottled water & Drinking water is bottled water/demi john.
-\\\midrule
- \texttt{uiws} & Unimproved water source & 
-\\\midrule
- \texttt{othw} & Other water source & 
  \\\midrule
@@
  & Limited water service, defined as an improved water source with round-trip collection time greater than 30 minutes.
-\\\midrule
- \texttt{watp} & Water on the premises
- & 
  \\\bottomrule
@@
  \texttt{land} & Land area 
-
  & (sq. km) [AG.LND.TOTL.K2]
 \\\midrule
- \texttt{lita} & Boat with a motor
- & 
+ \texttt{lita} & Literacy rate, adult total 
+ & (\% of people ages 15 and above) [SE.ADT.LITR.ZS]
 \\\midrule
- \texttt{lity} & Boat with a motor
- & 
+ \texttt{lity} & Literacy rate, youth 
+ & (ages 15-24), gender parity index (GPI) [SE.ADT.1524.LT.FM.ZS]
 \\\midrule
- \texttt{mori} & Boat with a motor
- & 
+ \texttt{mori} & Mortality rate, under-5 
+ & (per 1,000 live births) [SH.DYN.MORT]
 \\\midrule

### 2021-02-27 docs/manuscript/manuscript.tex

@@
     \bf Variable & \bf Description & \bf Notes \\\midrule
-\texttt{imws} & Improved water source
- & 
- \\\midrule
  \texttt{phom} & Piped into dwelling &
@@
 & 
-\\\midrule
- \texttt{uiws} & Unimproved water source & 
 \\\midrule
@@
  & 
- \\\midrule
- \texttt{iwsp} & Improved water source on the premises & The de jure population living in households with an improved water source on the premises.
-\\\midrule
- \texttt{basw} & Basic water service
- & Defined as an improved water source with either water on the premises or round-trip collection time is 30 minutes or less.
-\\\midrule
- \texttt{liws} & Limited water service
- & Limited water service, defined as an improved water source with round-trip collection time greater than 30 minutes.
  \\\bottomrule
@@
 
-Across 78 countries in this study, the data provides 28 numerical variables and four descriptive variables.
+Across 78 countries in this study, the data provides 28 quantitative variables and three descriptive variables such as the country names, survey type, and survey year.
 The majority of the data is from 2010 - 2020 as shown in \autoref{fig:year}.
-Therefore, we can analyze the global condition of water accessibility by utilizing the most up to date surveys.
+Therefore, we can analyze the global condition of water accessibility by utilizing the most up-to-date surveys.
 Most of the variables are related to water accessibility by their locations, time, and water resources.
@@
 \subsection{Cluster analysis}
-We selected 11 variables that describe water accessibility of 78 countries around the world \citep{price2019difference}.
+We selected 22 variables that describe water accessibility of 78 countries around the world \citep{price2019difference}.
 Next, we used clustering, an unsupervised machine learning method to identify and group water accessibility variables in a large data set to understand and make inferences using R.
@@
 First, we used different clustering algorithms to determine the best clustering results.
-We tested and categorized differently by using single, average, complete, ward.D, and ward.D2 methods to develop clustering dendrogram.
+We tested and clustered using different cluster methodology by utilizing single, average, complete, ward.D, and ward.D2 methods to develop initial clustering dendrogram.
 Then, we analyzed by determining the best hierarchical relationship of each of the methods by observing the tree diagram itself; more even distributions of each tree are interpreted as optimal clustering methods.
-We determined that ward.d2 had most even distribution, verified through the use of 'NBClust' to determine the best algorithm as well as the best cut or groupings for the clustering methods as shown in \autoref{fig:optimal-clusters}.
+We determined that ward.d2 had the evenest distribution, verified through the use of 'NBClust' to determine the best algorithm as well as the best cut or groupings for the clustering methods as shown in \autoref{fig:optimal-clusters}.
 
 We compared Ward.D2, Ward.D, complete, and k-means algorithm methods to determine the best cuts, and also to confirm that Ward.D2 was the best clustering algorithm method to develop the deprogram tree.
-The 'NBClust' function was set to determine the best clustering methodology by choosing minimum clustering from 3 to max cuts of 7 and utilized all the indices.
+The 'NBClust' function was set to determine the best clustering methodology by choosing minimum clustering from 3 to max cuts of 7 and utilized all the indices.NBClust function in R is the 30 algorithm indices for defining the number of clusters and proposes to the user the best clustering design from the different outcomes acquired by diversifying all combinations of many clusters, distance measures, and clustering methods.
+Therefore, NBClust determines the best clustering cut for the different methods by determining the mode of the cut numbers which then is called the majority rule.
 The majority rule was implemented by the 'NBClust' function for Ward.d2, Ward.D, and the complete algorithm determined that the best clustering cut was four.
-However, k-means determined to be 3, but we determined that Ward.D2 is the best methodology thus far.
+While, k-means method's majority rule is determined to be three cuts, but we determined that Ward.D2 is the best methodology due to the random nature of the k-mean clustering method.
 

### 2021-03-03 docs/manuscript/manuscript.tex
Updates from Overleaf

@@
 \subsection{Data Sources}
-Data for this study were collected from the Demographic Health Survey (DHS) program, which is a program that has collected, analyzed, and distributed accurate and representative data on population, health, HIV, and nutrition through more than 400 surveys in over 90 countries.
+Data for this study were collected from the Demographic Health Survey (DHS) program, which is a program that has collected, analyzed, and distributed accurate and representative data on population, health, HIV, and nutrition through more than 400 surveys in over 90 countries \citep{statcompiler}.
 DHS the program uses a wide range of monitoring and impact evaluation indicators in the areas of population, health, and nutrition.
@@
 In many cases, bio-marker experiments for malaria and anemia are also included.
-We used the STAT-Compiler tool available from the DHS Program to develop customized tables on indicators related to water, transportation, and wealth from 78 countries between 1990-2019.
+We used the STAT-Compiler tool available from the DHS Program to develop customized tables on indicators related to water, transportation, and wealth from 73 countries between 2000-2019 \citep{statcompiler}.
+
+We also used World Bank's database to gain additional explanatory variable data for this study.
+World Bank (WB) provides over 200,000 published documents to understand the development policies and programs globally.
+Their reports, working papers, documents, and primary data are free to the public.
+WB DataBank is an analysis and visualization instrument that contains examples of time series data on many topics.
+We generated tables to analyze them by R.
+We used World Development Indicators (WDI) within the DataBank is the primary World Bank collection of development indicators, organized from officially-recognized international sources.
+It exhibits the most current, accurate global development, and available data includes national, regional, and global estimates.
+We collected the 2018 surveys for the 73 countries for the following variables mentioned in \autoref{tab:summary-var}.
 
@@
 
-We preserved 28 variables from 78 where less than 50\% of values were missing and the population survey data were removed because they were similar to the household surveys.
+We preserved 28 variables from 73 countries where less than 50\% of values were missing and the population survey data were removed because they were similar to the household surveys.
 Ultimately, the simplified data frame was 78 countries by 28 variables.
@@
 One focused on the water accessibility variables and the other focused on the explanatory variables \citep{price2019difference}.
-From the remaining variables, we selected 22 which related to water accessibility.
-These included: types of water source (improved water source,
-piped into dwelling, piped into yard/plot,
- public tap/standpipe, tube well/borehole, protected well, protected spring, rainwater, unimproved water source, unprotected well, unprotected spring, cart with small tank/drum, tanker truck, bottled water, other water source, surface water), time to water source (less than 30 minutes, more than 30 minutes), and premise to water source (water on the premises, improved water source on the premises, basic water service, limited water service) \autoref{tab:summary-wa}.
+From the remaining variables, we selected 17 which related to water accessibility.
+These included: types of water source (piped into dwelling, piped into yard/plot,
+ public tap/standpipe, tube well/borehole, protected well, protected spring, rainwater, unprotected well, unprotected spring, cart with small tank/drum, tanker truck, bottled water, other water source, surface water), time to water source (less than 30 minutes, more than 30 minutes), and premise to water source (water on the premises) \autoref{tab:summary-wa}.
 Remaining variables related to wealth and transportation were used as an explanatory variables in the model \autoref{tab:summary-var}.
@@
     \bf Variable & \bf Description & \bf Notes \\\midrule
- \texttt{phom} & Piped into dwelling &
+ \texttt{phom} & Piped into dwelling & DHS Program Surveys
  \\\midrule
  \texttt{pipy} & Water piped into yard/plot
-&
+& DHS Program Surveys
 \\\midrule
     \texttt{ptap} &  Public tap/standpipe
- & 
+ & DHS Program Surveys
  \\\midrule
@@
 \texttt{pwel} & Protected well
-& 
+& DHS Program Surveys
 \\\midrule
 \texttt{pspr} & Protected spring
-& 
+& DHS Program Surveys
 \\\midrule
@@
 \texttt{rain} & Rainwater
-& 
+& DHS Program Surveys
 \\\midrule
 \texttt{uwel} & Unprotected well water
-&
+& DHS Program Surveys
 \\\midrule
 \texttt{uspr} & Unprotected spring
-& 
+& DHS Program Surveys
 \\\midrule
 
- \texttt{truc} & Tanker truck & 
+ \texttt{truc} & Tanker truck & DHS Program Surveys
 \\\midrule
- \texttt{ctan} & Cart with tank & 
+ \texttt{ctan} & Cart with tank & DHS Program Surveys
 \\\midrule
- \texttt{bott} & Bottled water & Drinking water is bottled water/demi john.
+ \texttt{bott} & Bottled water & Drinking water is bottled water/demi john (DHS Program Surveys).
 \\\midrule
- \texttt{othw} & Other water source & 
+ \texttt{othw} & Other water source & DHS Program Surveys
 \\\midrule
 \texttt{surw} & Surface water
-& 
+& DHS Program Surveys
 \\\midrule
 \texttt{tles} & Water 30 minutes or less away round trip
-&
+& DHS Program Surveys
 \\\midrule 
-\texttt{tmor} & Water more than 30 minutes away round trip & 
+\texttt{tmor} & Water more than 30 minutes away round trip & DHS Program Surveys
 \\\midrule
  \texttt{watp} & Water on the premises
- & 
+ & DHS Program Surveys
  \\\bottomrule
@@
  \texttt{bicy} & Bicycle
- &
+ & DHS Program Surveys
 \\\midrule
  \texttt{acar} & Animal drawn cart
- & 
+ & DHS Program Surveys
 \\\midrule
  \texttt{mcyc} & Motorcycle
- & 
+ & DHS Program Surveys
 \\\midrule
  \texttt{pcar} & Private car
- & 
+ & DHS Program Surveys
 \\\midrule
  \texttt{cgdp} & GDP (current US \$) 
- & [NY.GDP.MKTP.CD]
+ & World Bank Data [NY.GDP.MKTP.CD]
 \\\midrule
  \texttt{tpop} & Total Population
- &  [SP.POP.TOTL]
+ & World Bank Data   [SP.POP.TOTL]
 \\\midrule
  \texttt{upop} & Urban population 
- & (\% of total population)  [SP.URB.TOTL.IN.ZS]
+ & World Bank Data  (\% of total population)  [SP.URB.TOTL.IN.ZS]
 \\\midrule
  \texttt{popd} & Population density 
- & (people per sq. km of land area)  [EN.POP.DNST]
+ & World Bank Data  (people per sq. km of land area)  [EN.POP.DNST]
 \\\midrule
  \texttt{land} & Land area 
- & (sq. km) [AG.LND.TOTL.K2]
+ & World Bank Data  (sq. km) [AG.LND.TOTL.K2]
 \\\midrule
  \texttt{lita} & Literacy rate, adult total 
- & (\% of people ages 15 and above) [SE.ADT.LITR.ZS]
+ & World Bank Data  (\% of people ages 15 and above) [SE.ADT.LITR.ZS]
 \\\midrule
  \texttt{lity} & Literacy rate, youth 
- & (ages 15-24), gender parity index (GPI) [SE.ADT.1524.LT.FM.ZS]
+ & World Bank Data  (ages 15-24), gender parity index (GPI) [SE.ADT.1524.LT.FM.ZS]
 \\\midrule
  \texttt{mori} & Mortality rate, under-5 
- & (per 1,000 live births) [SH.DYN.MORT]
+ & World Bank Data  (per 1,000 live births) [SH.DYN.MORT]
 \\\midrule
  \texttt{wigc} & Wealth index Gini coefficient
- & The Gini coefficient indicates the level of concentration of wealth in the country.
+ & The Gini coefficient indicates the level of concentration of wealth in the country (World Bank Data).
 

### 2021-03-03 docs/manuscript/manuscript.tex

@@
 They are DHS, MIS, and AIDS Indicator Survey (AIS).
-After, we used R to develop an algorithm to find the most recent surveys and showed results for 78 countries.
+After, we used R to develop an algorithm to find the most recent surveys and showed results for 73 countries.
 We also simplified the variables titles into four character acronyms to recognize the different 28 variables quickly.
 We used the household data only.
-The STAT-compiler collected the aggregated data by the 78 countries
+The STAT-compiler collected the aggregated data by the 73 countries
 
 We preserved 28 variables from 73 countries where less than 50\% of values were missing and the population survey data were removed because they were similar to the household surveys.
-Ultimately, the simplified data frame was 78 countries by 28 variables.
+Ultimately, the simplified data frame was 73 countries by 28 variables.
 After the simplification, two separate tables are distinguished within the 28 variables.
@@
     \bf Variable & \bf Description & \bf Notes \\\midrule
- \texttt{phom} & Piped into dwelling & DHS Program Surveys
+ \texttt{phom} & Piped into dwelling & 
  \\\midrule
  \texttt{pipy} & Water piped into yard/plot
-& DHS Program Surveys
+& 
 \\\midrule
     \texttt{ptap} &  Public tap/standpipe
- & DHS Program Surveys
+ & 
  \\\midrule
@@
 \texttt{pwel} & Protected well
-& DHS Program Surveys
+& 
 \\\midrule
 \texttt{pspr} & Protected spring
-& DHS Program Surveys
+& 
 \\\midrule
@@
 \texttt{rain} & Rainwater
-& DHS Program Surveys
+& 
 \\\midrule
 \texttt{uwel} & Unprotected well water
-& DHS Program Surveys
+&
 \\\midrule
 \texttt{uspr} & Unprotected spring
-& DHS Program Surveys
+& 
 \\\midrule
 
- \texttt{truc} & Tanker truck & DHS Program Surveys
+ \texttt{truc} & Tanker truck & 
 \\\midrule
- \texttt{ctan} & Cart with tank & DHS Program Surveys
+ \texttt{ctan} & Cart with tank & 
 \\\midrule
- \texttt{bott} & Bottled water & Drinking water is bottled water/demi john (DHS Program Surveys).
+ \texttt{bott} & Bottled water & Drinking water is bottled water/demi john.
 \\\midrule
- \texttt{othw} & Other water source & DHS Program Surveys
+ \texttt{othw} & Other water source & 
 \\\midrule
 \texttt{surw} & Surface water
-& DHS Program Surveys
+& 
 \\\midrule
 \texttt{tles} & Water 30 minutes or less away round trip
-& DHS Program Surveys
+& 
 \\\midrule 
@@
  \texttt{watp} & Water on the premises
- & DHS Program Surveys
+ & 
  \\\bottomrule
@@
 
-Across 78 countries in this study, the data provides 28 quantitative variables and three descriptive variables such as the country names, survey type, and survey year.
+Across 73 countries in this study, the data provides 28 quantitative variables and three descriptive variables such as the country names, survey type, and survey year.
 The majority of the data is from 2010 - 2020 as shown in \autoref{fig:year}.
@@
 \subsection{Cluster analysis}
-We selected 22 variables that describe water accessibility of 78 countries around the world \citep{price2019difference}.
+We selected 17 variables that describe water accessibility of 73 countries around the world \citep{price2019difference}.
 Next, we used clustering, an unsupervised machine learning method to identify and group water accessibility variables in a large data set to understand and make inferences using R.

### 2021-03-04 docs/manuscript/manuscript.tex

@@
   \includegraphics[width=.7\textwidth]{adv-cor-plot}
-  \caption{Advanced Correlation Graph of Numerical Variables}
+  \caption{Advanced Correlation Graph of Water Variables}
   \label{fig:cor-plot}
@@
 There are great internally heterogeneous clusters for the 76 distinct countries.
-The cluster analysis' dendrogram is shown in \autoref{fig:dendrogram}, and the corresponding NBClust curve yielding the optimal four clusters is shown in \autoref{fig:optimal-clusters}, \autoref{fig:optimal}.
+The cluster analysis' dendrogram is shown in \autoref{fig:dendrogram}, and the corresponding NBClust curve yielding the optimal four clusters is shown in \autoref{fig:optimal-clusters}, \autoref{fig:optimal-clusters-maj}, \autoref{fig:optimal}.
 \autoref{tab:water-cluster} shows the resulting four water accessibility country clusters.
@@
   \label{fig:optimal-clusters}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{optimal-clusters-wardd2}
+  \caption{Optimal Number of Clusters for Ward.D2}
+  \label{fig:optimal-clusters-maj}
 \end{figure}

### 2021-03-04 docs/manuscript/manuscript.tex
Updates from Overleaf

@@
 \\\midrule 
-\texttt{tmor} & Water more than 30 minutes away round trip & DHS Program Surveys
+\texttt{tmor} & Water more than 30 minutes away round trip & 
 \\\midrule
@@
  \texttt{wigc} & Wealth index Gini coefficient
- & The Gini coefficient indicates the level of concentration of wealth in the country (World Bank Data).
+ & The Gini coefficient indicates the level of concentration of wealth in the country (DHS Program).
 
@@
   \label{fig:cor-plot}
+\end{figure}
+Next, we studied the explanatory variables that we collected and extracted from the World Bank Data.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.7\textwidth]{cor-wb}
+  \caption{Advanced Correlation Graph of World Bank Explanatory Variables}
+  \label{fig:cor-wb}
 \end{figure}

### 2021-03-09 docs/manuscript/manuscript.tex
Updates from Overleaf

@@
 Next, we studied the explanatory variables that we collected and extracted from the World Bank Data.
+Then we developed an advanced correlation graph to quickly analyze the relationships between the eight variables previously mentioned in \autoref{tab:summary-var}.
+Total adult literacy rate and youth literacy rate (ages 15-24) has the strongest correlation of .97 and p-value of less than 0.001 which are summarized with three asterisks in \autoref{fig:cor-wb}.
+GDP (current US Currency) and Total Population has the second strongest correlation of .95 and p-value of less than .001 in \autoref{fig:cor-wb}.
+These explanatory variables will assist in finding results in the classification of which explanatory variables may affect the improvement of the water accessibility variables.
 

### 2021-03-11 docs/manuscript/manuscript.tex

@@
 It exhibits the most current, accurate global development, and available data includes national, regional, and global estimates.
-We collected the 2018 surveys for the 73 countries for the following variables mentioned in \autoref{tab:summary-var}.
+We collected the most data available year (2018 and 2014) surveys for the 73 countries for the following variables mentioned in \autoref{tab:summary-var}.
 
@@
  & World Bank Data  (per 1,000 live births) [SH.DYN.MORT]
+\\\midrule
+ \texttt{prec} & Average precipitation in depth (mm per year)
+ & World Bank Data  Average precipitation in depth (mm per year)
+ [AG.LND.PRCP.MM]
 \\\midrule
@@
 
+\subsection{Factor Analysis}
+
 \subsection{Water accessibility typologies}

### 2021-03-15 docs/manuscript/manuscript.tex

@@
 \section{Introduction}
-In 2017, 29\% of the global population (2.2 billion people) did not use a safely managed drinking-water service, which is defined as water located on-premises, available, and free from contamination \citep{who2015drinkingwater}.
-If you meet about 100 people, you can imagine that 29 of the people you meet do not have access to improved drinking water.
+In 2017, 29\% of the global population (2.2 billion people) did not use a safely managed drinking water service, which is defined as water located on-premises, available, and free from contamination \citep{who2015drinkingwater}.
 Where this raises concerns about inequality problems related to the task and time variation
@@
 Next, we present the data and methods...''}}
-\section{Objectives}
+%\section{Objectives}
 The objective of this work is to  investigate associations between measures of water accessibility and access to modes of transit on the national scale. 
-
 In the next section, we discuss relevant prior research on water accessibility quantification. We are interested in what type of modes of
@@
 Lack of water accessibility is a major crisis for many vulnerable people around the world.
-We have determined to seek literature reviews from the past 15 years to have the most up to date information because information on water accessibility is a topic that is constantly changing throughout the world due to different international, government, and even local policies.
+Information on water accessibility is a topic that is constantly changing throughout the world due to different international, government, and even local policies.
 The literature review intends to investigate different journal articles on water accessibility on a global scale.
@@
 We will always gain new insight and wisdom on information about water accessibility through the global scale as we investigate and determine literature reviews chronologically.
+
 \subsection{2000-2015}
@@
 However, some researchers raise concerns on lack of clarity with the global water accessibility monitoring \citep{onda2012global, jia2016highresolution}.
-However, in an era of rapidly changing technologies, international and government policies, the older researches are not adequate.
+However, in an era of rapidly changing technologies, international and government policies, more relevant indicators are required.
 For an example of The eight Millennium Development Goals (MDGs) upgrading into The Sustainable Development Goals (SDGs).
@@
 Nonetheless, there is a lack of robust study on transportation to enhance water accessibility for all people by 2030.
-\subsection{2015-2021}
+
+\subsection{2015-current}
 However, in the most recent years, there is a lack of information that needs to focus on people to receive accessible drinking water.
@@
 A correlation chart was used to assess strong correlations between the 28 numerical variables that may pertain to water accessibility \autoref{fig:cor-basic}.
-Households with limited water service has the strongest and positive relationship with households possessing a bicycle.
-Meanwhile, households using an improved water source has the strongest and negative relationship with households using an unimproved water source.
+Countries with a high percentage of households with limited water service has the strongest and positive relationship with households possessing a bicycle.
+Meanwhile, countries with a high percentage of households using an improved water source has the strongest and negative relationship with countries with a high percentage of households using an unimproved water source.
 This initial finding results indicated that the initial variables are logical and can give us insight into further understanding and the relationship between these variables.
@@
 As we continued to study further upon the correlations of the variables, we used more advanced correlation graphs to determined their p-value, and to see a relationship between different categorical variables is shown in \autoref{fig:cor-plot}, which again indicates a greater detailed comparison between the variables.
-Households with an improved water source on the premises have a strong relationship with households with basic water service of .84 and a p-value of less than 0.001.
-Also, we learned and discovered that Households using an unimproved water source have a positive relationship with Households using unprotected well water, Households using an unprotected spring, and Households using surface water with a p-value of less than 0.001 as well.
+Countries with a high percentage of households with an improved water source on the premises have a strong relationship with countries with a high percentage of households with basic water service of .84 and a p-value of less than 0.001.
+Also, we learned and discovered that countries with a high percentage of households using an unimproved water source have a positive relationship with countries with a high percentage of households using unprotected well water, countries with a high percentage of households using an unprotected spring, and countries with a high percentage of households using surface water with a p-value of less than 0.001 as well.
+The reason for the strong correlations is because the surveys has derivation from each of the responses.
 These relationships justify the necessity to research further if the particular water accessibility variables can increase the accessibility of particular water sources to reach universal access to water by 2030, as proposed in the SDGs.
@@
 There is a significant positive relationship between households possessing a private car and households with water on the premises as shown in \autoref{fig:transporation-time}.
-There is also is a significant negative relationship between households possessing a private car with households with water 30 minutes or less away round trip and households with water more than 30 minutes away round trip with a p-value of less than 0.001.
+There is also is a significant negative relationship between countries with a high percentage of households possessing a private car with countries with a high percentage of households with water 30 minutes or less away round trip and countries with a high percentage of households with water more than 30 minutes away round trip with a p-value of less than 0.001.
+This gives an inference that owning a private car could mean that the households are wealthy, and therefore they may have piped or close water sources nearby.
 This information is essential to investigate further why this is the global trend.
@@
 
-\autoref{fig:transporation-water-one} shows that Households possessing a motorcycle has a positive correlation with Households using a tube well/borehole and Households using rainwater with a p-value of less than 0.01.
+\autoref{fig:transporation-water-one} shows that countries with a high percentage of households possessing a motorcycle has a positive correlation with countries with a high percentage of households using a tube well/borehole and countries with a high percentage of households using rainwater with a p-value of less than 0.01.
 Similar to the previous analysis, there is a relationship between a mode of transportation to retrieve potential water sources.
-\autoref{fig:transporation-water-two} also shows that households possessing an animal-drawn cart have a positive and strong relationship with households using unprotected well water with a p-value of less than 0.001.
-The results demonstrate that different modes of transportation factor into the accessibility of retrieving water sources on a global scale, therefore, there should be consideration of these variables to increase access to improved water sources.
+\autoref{fig:transporation-water-two} also shows that countries with a high percentage of households possessing an animal-drawn cart have a positive and strong relationship with countries with a high percentage of households using unprotected well water with a p-value of less than 0.001.
+The analysis potentially demonstrates that different modes of transportation factor into the accessibility of retrieving water sources on a global scale, therefore, there should consideration of explanatory variables to enhance access to improved water sources.
 \begin{figure}[h!]
@@
 \subsection{Factor Analysis}
+Before, we investigate the typologies or understand the clusters further, it is necessary to implement factor analysis to reduce the 17 water accessibility variables which is mentioned in \autoref{tab:summary-wa}.
+Factor analysis is a statistical method to explain variability among observed, correlated variables to decrease number of unobserved variables called factors.
 
@@
 
-NBClust statistic analysis developed and determined four distinct groups.
+NBClust statistic analysis developed and determined four distinct groups of the four factors that were derived from water accessibility variables mentioned in \autoref{tab:summary-wa}.
 The first cluster group is composed of countries near the equator or in the Sub-Sahara Regions.
@@
 
-\section{Limitation}
+\section{Limitations}
 %The problem of missing data is relatively common in almost all research and can have a significant effect on the conclusions that can be drawn from the data.
